### 2003-03-08 trunk/vcd/paper/draft.Rnw
Draft for vcd paper added. dsc2003.* is now depracated and will probably be removed. Draft is now an Sweave file and therefore the graphics were removed from the cvs.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@217 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[10pt]{article}
\usepackage{color, natbib, graphicx}
\usepackage{dsc2003}

\setlength{\parskip}{0.5ex plus0.1ex minus0.1ex}
\setlength{\parindent}{0em}

\bibpunct{(}{)}{;}{a}{}{,}

\begin{document}

\SweaveOpts{engine=R,eps=FALSE}

<<preliminaries,echo=FALSE,results=hide>>=
library(vcd)
data(Bundesliga)
goals <- xtabs(~ AwayGoals + HomeGoals, data = Bundesliga, subset = Year == 1995)
data(HairEyeColor)
haireye <- margin.table(HairEyeColor, 1:2)
data(UCBAdmissions)
@

\title{Visualizing the Independence Problem Using Extended Association and
Mosaic Plots}
\author{\hfill David Meyer$^\dag$ \hfill Achim Zeileis$^\dag$ \hfill Kurt Hornik$^\ddag$ \hfill \hfill \\[0.2cm]
\dag {\it \small Institut f\"ur Statistik \& Wahrscheinlichkeitstheorie,
 Technische Universit\"at Wien}\\
\ddag {\it \small Institut f\"ur Statistik, Wirtschaftsuniversit\"at Wien}
}

\maketitle

\begin{abstract}
Two visualization techniques for the independence problem in 2-way
contingency tables---assocation and mosaic plots---are extended in two directions:
\begin{enumerate}
\item The visualization is enhanced by displaying the significance of an appropriate
test for independence and by using better color schemes.
\item The implementation in the \textsf{R} system is improved using a more modular
design and allowing for more flexible specification of plotting parameters.
\end{enumerate}
Furthermore, some enhancements for displays of various types of independence in
general multi-way tables are suggested.\footnote{Please note that all of this is
work in progress and neither theoretically nor computationally fully sound.}
\end{abstract}

\section{Introduction} \label{sec:introduction}

Statistical models built for the analysis of multivariate data
quickly become complex with increasing dimensionality.
One idea of visualization techniques is to
use the human visual system to detect structures in the data that
possibly are not obvious from solely numeric output (e.g., test
statistics).
The \textsf{R} package \texttt{vcd}---inspired by the Book ``Visualizing Categorical
Data'' \citep{vcd:Friendly:2000}---includes methods for the (mostly
graphical) exploration of categorical data, such as:
\begin{itemize}
 \item fitting and graphing of discrete distributions,
 \item plots and tests for independence and symmetry problems,
 \item visualization techniques for log-linear models.
\end{itemize}
Here, we focus on the visualization of the independence
problem, in particular in 2-way tables.

\section{Tests for independence in 2-way tables} \label{sec:tests}

We consider a 2-way contingency table with cell frequencies $(n_{ij})$ for
$i = 1, \dots, I$ and $j = 1, \dots, J$ and row and column sums
$n_{i+} = \sum_i n_{ij}$ and $n_{+j} = \sum_j n_{ij}$ respectively. For
convenience the number of observations is denoted $n = n_{++}$.

Given an underlying distribution with theoretical cell probabilities
$\pi_{ij}$ the null hypothesis of independence of the two categorical
variables can be formulated as
\begin{equation} \label{eq:H0}
H_0: \; \pi_{ij} \quad = \quad \pi_{i+} \pi_{+j}.
\end{equation}

The expected cell frequencies in this model are
$\hat n_{ij} = n_{i+}n_{+j}/n$. The best known and most used way
to measure the discrepancy
between observed and expected values are the Pearson residuals
\begin{equation} \label{eq:Pearson}
r_{ij} = \frac{n_{ij} - \hat n_{ij}}{\sqrt{\hat n_{ij}}}.
\end{equation}
Therefore, a rather intuitive idea is to reject the null hypothesis when
there are residuals which are too extreme, i.e., not close enough to
zero. The most convenient way to aggregate the $I \times J$ residuals
to one test statistic is their squared sum
\begin{equation} \label{eq:ChiSquare}
X^2 = \sum_{i, j}r_{ij}^2,
\end{equation}
because it is known to have a limiting $\chi^2$ distribution with $(I-1)(J-1)$
degrees of freedom under the null hypothesis.
This is the well-known $\chi^2$ test for independence in 2-way tables.
%% reference needed

But this is not the only plausible way of aggregation of the Pearson
residuals. There are many thinkable functionals $\lambda(\cdot)$ which
lead to reasonable test statistics $\lambda(r_{ij})$, the sum of squares
is just one of them. Another functional which is very helpful for identifying
the cells which cause the potential dependence is the maximum of the absolute
values
\begin{equation} \label{eq:MaxAbs}
M = \max_{i, j} |r_{ij}|.
\end{equation}
Given a critical value $c_\alpha$ for this test statistic all residuals whose
absolute value exceeds $c_\alpha$ violate the hypothesis of indendence at level
$\alpha$ \cite[][ch. 7]{vcd:Mazanec+Strasser:2000}. Thus, the culprits causing
the dependence can be easily identified. The critical values $c_\alpha$ can be
derived from the conditional permutation distribution of the table $(n_{ij})$.
Of course, the $\chi^2$ test \ref{eq:ChiSquare} can also be carried out as
a conditional test instead of using the unconditional asymptotic distribution.

Other classical tests for the hypothesis of independence include
Fisher's exact test, asymptotic tests of the odds ratio
and the Mantel-Haenzel test.

\section{Visualization techniques for 2-way tables} \label{sec:visualization}

The two best known visualization techniques for independence in 2-way tables are
association plots and mosaic plots. Both are suitable to bring out departure of an
observed table $(n_{ij})$ from the expected table $(\hat n_{ij})$ in a graphical way.
The former focuses on the visualization of the Pearson residuals $r_{ij}$ (under
independence) while the latter primarily displays the observed frequencies
$n_{ij}$.

\subsection{Association plots}

Association plots \citep{vcd:Cohen:1980} visualize the table of Pearson residuals:
each cell is represented by a rectangle that has (signed) height
proportional to the corresponding Pearson-residual $r_{ij}$ and width proportional to the
square root of the expected counts $\sqrt{\hat n_{ij}}$. Thus the area is proportional to
the raw residuals $n_{ij} - \hat n_{ij}$. The sign of the residual is redundantly
coded by orientation and color of the corresponding rectangle.

Figure \ref{fig:goals-assoc-classic} shows the association plot for the table of
home and away goals for all 306 games in the 1995/6 season of the German soccer league
Bundesliga \citep{vcd:Knorr-Held:1999}. In particular, it can be seen that there are
fewer games ending 1:0 and more ending 1:1 than would be expected under independence.
But it is absolutely unclear whether this difference is significant or not.

\begin{figure}[htbp]
\begin{center}
<<goals-assoc-classic,fig=TRUE,echo=FALSE,height=5,width=5>>=
grid.assocplot(goals, rot = 0)
@
\caption{Association plot for the Bundesliga data.}
\label{fig:goals-assoc-classic}
\end{center}
\end{figure}

\subsection{Mosaic plots}

Mosaic plots can be seen as an extension of grouped
bar charts, where width and height of the bars show the relative
frequencies of the two variables: a mosaic plot simply
consists of a collection of tiles whose
sizes are proportional to the observed cell frequencies
(see Figure \ref{fig:goals-mosaic-classic}).

\begin{figure}[htbp]
\begin{center}
<<goals-mosaic-classic,fig=TRUE,echo=FALSE,height=5,width=5>>=
mosaicplot(t(goals), shade = TRUE, main = "")
@
\caption{Friendly mosaic plot for the Bundesliga data.}
\label{fig:goals-mosaic-classic}
\end{center}
\end{figure}

Sequential horizontal and vertical recursive splits are used to
visualize the frequencies of more than 2 variables, each new variable
conditional to the previously entered variables. A first extension by
\cite{vcd:Friendly:1994} uses a color coding of the tiles to visualize deviations
(residuals) from a given log-linear model fitted to the table, that is, from
the expected frequencies under arbitrary independence hypotheses. In this
extension, the sign of the residuals is coded by solid and dashed line
rectangles respectively. Furthermore, residuals that exceed an absolute
value of 2 are shaded light blue and red respectively, those that even
exceed an absolute value of 4 are shaded with full saturation (in an
HSV color scheme). The heuristic idea behind this shading is that
cells with a large residual might cause the maximum test (\ref{eq:MaxAbs})
to reject the hypothesis of independence. But in general it is unknown to
which significance level $\alpha$ the values $c_\alpha = 2$ and $4$
correspond.

Figure \ref{fig:goals-mosaic-classic} shows the Friendly mosaic plot for
the same Bundesliga data as in Figure \ref{fig:goals-assoc-classic}.
In this plot the residuals
are visualized by the color shading of the cells. Again, the cells for
the 1:0 and 1:1 results are highlighted, furthermore there seem to be
too many 6:3 results as would be plausible under independence. Whereas
the association plot in Figure \ref{fig:goals-assoc-classic} contains
no information about the result of a test for independence, the shading
in this plot conveys the impression that there might indeed be
significant departure from independence (which would indicate that there
is not only {\em no} home field advantage in the German Bundesliga but even
an advantage for the away team). However, it is not clear if there
is indeed evidence for this or not.

\section{Extensions} \label{sec:extensions}

In this section we suggest some improvements and extensions which can be
applied to both visualization techniques introduced in Section
\ref{sec:visualization}. Firstly, the visual enhancements
\begin{itemize}
\item combining visualization of and testing for independence
\item better color schemes
\end{itemize}
are addressed. Secondly, the implementation enhancements
\begin{itemize}
\item more modular implementation using \texttt{grid}
\item more flexible plotting parameter specification
\end{itemize}
are described. None of these should be thought of as finished work
but rather as a collection of ideas which still need thorough
refinement and proper implementation.

\subsection{Visualization enhancements}

The extensions of \cite{vcd:Friendly:1994} to mosaic plots provide
substantial improvement of the original mosaic plots and lifted them
from a plot for contingency tables to a visualization technique for
independence problems. However, it has two major drawbacks: Firstly,
the significance level for the hard-coded critical values 2 and 4 is
usually unknown. Secondly, the HSV colors for (light) blue and red
were chosen due to the restrictions of the SAS software (in which
the plots were first implemented), they are not device-independent and
do not provide homogeneous saturations over different colors
and copier proofness. We suggest some ways to overcome these drawbacks.

As pointed out in Section \ref{sec:tests} the critical values for
the maximum statistic $M$ from (\ref{eq:MaxAbs}) can be derived
from the conditional permutation distribution. Instead of the hard-coded
values 2 and 4 the particular critical values, e.g., at the level
$\alpha = 0.9$ and $0.95$, for the table to be visualized could be
used, thus highlighting exactly those residuals who cause a (potential)
dependency within the table. At the moment these critical values
are derived in \texttt{vcd} by simulation of the underlying distribution.

As an alternative to the HSV colors the device-independent
HCL (Hue-Chroma-Luminance) color space \citep{vcd:Ihaka:2003} could be
used. The resulting association plot (using a critical value for
$\alpha = 0.95$) for the Bundesliga data can be seen in
Figure \ref{fig:goals-assoc-signif}. It indicates very clearly that
there is no evidence for rejecting the hypothesis of independence
in this table. Thus, there seems to be neither an advantage for
the home team nor for the away team.

\begin{figure}[htbp]
\begin{center}
<<goals-assoc-signif,fig=TRUE,echo=FALSE,height=5,width=5>>=
grid.assocplot(goals, color = "signif", rot = 0)
@
\caption{Extended association plots for the Bundesliga data.}
\label{fig:goals-assoc-signif}
\end{center}
\end{figure}

To illustrate what a `postive' example looks like we give the association plot
for the famous Hair-Eye-Color data set cross-tabulating the hair and eye
color of 592 statistics students. It is easy to see the hypothesis
of independence is rejected at 95\% level by the maximum test (\ref{eq:MaxAbs})
due to higher frequencies of student with hair and eye color
black/brown, blond/blue and red/green and lower frequencies of
blond/brown and black/blue than would be plausible under independence.

\begin{figure}[htbp]
\begin{center}
<<haireye-assoc-signif,fig=TRUE,echo=FALSE,height=5,width=5>>=
grid.assocplot(haireye, color = "signif")
@
\caption{Extended association plots for the Hair-Eye-Color data.}
\label{fig:haireye-assoc-signif}
\end{center}
\end{figure}

We have seen that this approach performs very well when the test statistic
which should be visualized is the maximum statistic (\ref{eq:MaxAbs}) but
it is difficult to extend this approach to general tests for independence,
in particular in the HCL space, where the range of chroma and luminance are
not independent. HSV space on the other hand has the advantage, that all
three dimensions hue, saturation and value have the range $[0,1]$ and can
be varied independently. Therefore, a different approach than the one
described above would be to use the hue for the sign of the residuals as
before, the saturation for the absolute size of the residuals and the
value as an indicator for the significance of some test statistic, only
using the full value when the overall test for independence rejects
the null hypothesis. The resulting mosaic plots under this paradigm
are depicted in Figure \ref{fig:both-mosaic-signif}. Again, it can
be clearly seen that despite some large residuals there is no
evidence against independence for the Bundesliga, but that there the
null hypothesis has to be rejected for the Hair-Eye-Color data.

\begin{figure}[htbp]
\begin{center}
<<goals-mosaic-signif,fig=TRUE,echo=FALSE,height=5,width=5>>=
grid.mosaicplot(goals, main = NULL)
@
<<haireye-mosaic-signif,fig=TRUE,echo=FALSE,height=5,width=5>>=
grid.mosaicplot(haireye, main = NULL)
@
\caption{Extended mosaic plots for the Bundesliga and Hair-Eye-Color data.}
\label{fig:both-mosaic-signif}
\end{center}
\end{figure}

\subsection{Implementation enhancements}

The current implementation of association and mosaic plots in \textsf{R} suffers
from two main disadvantages: Firstly, it is not easy to recycle the
plots in conditioning plots or pairs plots (like mosaic matrices) as
they have been implemented using \textsf{R}'s base graphics engine.
The new implementation was written from scratch in \texttt{grid} offering
much more versatility amongst some other minor advantages and convenient
improvements. Secondly, the graphics parameters of the rectangles in
association and mosaic plots (like in almost all standard \textsf{R}
plots), like color and line type, cannot be specified for each cell
by the user. To overcome this, the user can specify in the current
implementation in \texttt{vcd} either arrays of graphics parameters
of the same dimensionality as the object that is plotted or a function
which computes these graphics parameters based on the original table
and its Pearson residuals. Functions are provided for the shading schemes
described in the previous section.

\section{Multi-way tables}

Independence problems do not only occur in 2-way tables, although that
is an important special case, but they are also important in tables of
higher dimensionality and can follow much more complex patterns.
These are again defined based on the underlying table of theoretical cell probabilities
$(\pi_{ij\dots})$ with more than two dimensions. Models of interest include the
null hypotheses of:
\begin{itemize}
 \item total independence:
  $\pi_{ijk\dots} = \pi_{i++\dots}\pi_{+j+\dots}\pi_{++k\dots}\dots$
 \item conditional independence:
  $\pi_{ijk\dots} = \pi_{i|k\dots}\pi_{j|k\dots}$
 \item joint independence:
  $\pi_{ijk\dots} = \pi_{ij+\dots}\pi_{++k\dots}$
\end{itemize}
Classical non-graphical methods for these problems include the Chi-square
test, Fisher's exact test, the Cochran-Mantel-Haenzel test (for $2 \times 2 \times
K$-tables), and the analysis of log-linear models for more
complex settings.

Two natural ways to use the visualization techniques described in the previous
sections would be to use (Trellis-like) conditioned plots or pairs plots
(like mosaic matrices) to visualize these more complex patterns of
independence.

Two ideas for the problem of conditional independence are briefly outlined
here and illustrated using the famous admissions data of the University of
California at Berkley (UCB). In this data the question whether there is sex
discrimination at the UCB leads to the result that although women seem to
be disadvantaged at the aggregated level there is no sex discrimination
conditioned on the department---with the very exception of one department
in which women are {\em more} likely to be admitted than would be plausible
under independence. Exactly this is illustrated in the conditioned
assocation plot in Figure \ref{fig:UCB-assoc-signif}.

\begin{figure}[htbp]
\begin{center}
<<UCB-assoc-signif,fig=TRUE,echo=FALSE,height=5,width=5>>=
tabplot(UCBAdmissions, color = "signif")
@
\caption{Extended conditioned association plots for the UCB Admissions data.}
\label{fig:UCB-assoc-signif}
\end{center}
\end{figure}

Similarly, the same data can be visualized using a mosaic matrix where
a conditional independence model is fitted in each plot (see Figure
\ref{fig:UCB-mosaic-signif}).

\begin{figure}[htbp]
\begin{center}
<<UCB-mosaic-signif,fig=TRUE,echo=FALSE,height=5,width=5>>=
grid.mosaicpairs(UCBAdmissions, type = "conditional")
@
\caption{Extended conditioned association plots for the UCB Admissions data.}
\label{fig:UCB-mosaic-signif}
\end{center}
\end{figure}

\section{Conclusion}
We suggest a set of enhancements for visualizing the independence
problem in 2-way tables with some outlook to multi-way tables. The
extensions aim at improving the visualization by displaying both
the size of the residuals and the significance of a test for
independence and by using better color schemes. Furthermore, a new
implementation is outlined based on the graphics package \texttt{grid}
which provides more modular design and more flexible specification
of graphical parameters.

\bibliographystyle{asa}
\bibliography{vcd}

\end{document}
% LocalWords:  proofness