### 2005-07-29 trunk/vcd/inst/doc/strucplot.Rnw
Vignette added.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@416 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[a4paper]{article}
\usepackage{hyperref, graphicx, color, alltt, a4wide}
\usepackage{Sweave}
\usepackage[round]{natbib}
\definecolor{Red}{rgb}{0.7,0,0}
\definecolor{Blue}{rgb}{0,0,0.8}
\definecolor{hellgrau}{rgb}{0.55,0.55,0.55}
\newcommand{\pkg}[1]{{\normalfont\fontseries{b}\selectfont #1}}
\newcommand{\code}[1]{\texttt{#1}}
\newenvironment{smallexample}{\begin{alltt}}{\end{alltt}}

\begin{document}

%\VignetteIndexEntry{strucplots---Visualizing higher-dimensional contigency tables in vcd}
%\VignetteDepends{vcd}
%\VignetteKeywords{graphics, categorical data}
%\VignettePackage{vcd}

\SweaveOpts{engine=R,eps=TRUE}
\setkeys{Gin}{width=0.8\textwidth}

<<preliminaries,echo=FALSE,results=hide>>=
set.seed(1071)
library(vcd)
data(Titanic)
@

\title{strucplots---Visualizing higher-dimensional Contingency Tables in \pkg{vcd}}
\author{by David Meyer\\
  Wirtschaftsuniversit\"at Wien, Austria\\
\url{David.Meyer@R-project.org}
}
\maketitle
\sloppy

The \code{strucplot} framework in \pkg{vcd} can be used to visualize
higher-dimensional contingency tables and integrates techniques such as
mosaic displays and association plots. The main idea is to visualize
the tables' cells arranged in rectangular form. To allow for more than
two dimensions, the plot area is constructed using recursive
conditional splits, given the tiles representing the levels of the
previous dimension.
This principle defines a graph `family' with still many parameters
left such as:

\begin{itemize}
\item the split direction for each dimension
\item the content of the tiles
\item the graphical parameters of the tiles' content
\item the spacing between the tiles
\item the labeling of the tiles
\end{itemize}

\noindent The document at hand gives an introduction to the framework,
whereas labeling and shading issues are described in separate vignettes.

\section{Framework Overview}

The strucplot framework is highly modularized; Figure \ref{fig:struc}
shows the hierarchical relationship between the various components. 
The core function is \code{strucplot()} which processes the parameters,
sets up the graphical layout using grid viewports (see Figure \ref{fig:layout}),
and coordinates panel, labeling, shading, and spacing functions to produce 
the plot. The actual workhorse is the chosen panel function: currently, 
low-level panel functions for association plots (\code{struc\_assoc}) and 
mosaic plots (\code{struc\_mosaic}) are available. In addition, there are
high-level wrapper functions such as \code{mosaic()} and \code{assoc()} 
to \code{strucplot()} that conveniently set the 
parameters---it is these high-level functions that the user normally would like to use, 
although \code{strucplot()} is fully functional on its own. \code{mosaic()} and 
\code{assoc} are generic functions: for both, there exist a formula method and a default 
method. Other functions, such as \code{doubledecker()}, \code{pairs()}, and 
\code{cotabplot}, are on even higher level and make use of 
code{mosaic()} and \code{assoc()} to produce plots.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{struc}
  \caption{Strucplot components.}
  \label{fig:struc}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics{vcdViewport}
  \caption{Viewport layout for strucplots with their names. [A] =
    ``cornerTL'', [B] = ``cornerTR'', [C] = ``cornerBL'', [D] = ``cornerBR''.}
  \label{fig:layout}
\end{figure}

\section{Mosaic and Association Plots}

As an example, consider the famous `Survival on the Titanic' 
data containing two binary variables (survival to the disaster and
gender), as well as two categorical variables (class and age). The
`flattened' contincency table can be obtained using the
\code{structable()} function (quite similar to \code{ftable()}, but
allowing the specification of split directions as in mosaic():

<<structable,fig=FALSE,echo=TRUE,height=6,width=7,results=verbatim>>=
structable(Titanic)
@

Let us
first visualize the contingency table by the means of a mosaic plot
\citep{vcd:Hartigan+Kleiner:1984} which is basically 
an area-proportional visualization of (typically observed) frequencies, composed
of tiles (corresponding to the cells) created by recursive
vertical and horizontal splits of a square. Thus, the area of each tile
is proportional to the corresponding cell entry \emph{given} the
dimensions of previous splits (see Figure \ref{fig:observed}):
\begin{Sinput}
> mosaic(Titanic, type = "expected", main = TRUE, sub = "Expected values")
\end{Sinput}

\noindent The small bullets indicate zero entries in the corresponding cell.

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<Observed,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(Titanic, main = TRUE)
@
\caption{Mosaic plot for the Titanic data.}
\label{fig:observed}
\end{center}
\end{figure}

\noindent  It is also possible to visualize the expected values instead of the
observed values (see Figure \ref{fig:expected}):
\begin{Sinput}
> mosaic(Titanic, type = "expected", main = TRUE, sub = "Expected values")
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]

\begin{center}
<<Expected,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(Titanic, type = "expected", main = TRUE, sub = "Expected values")
@
\caption{Mosaic plot for the Titanic data (expected values).}
\label{fig:expected}
\end{center}
\end{figure}

\noindent The Pearson residuals (standardized deviations of observed from expected values) 
are preferably visualized using association
plots \citep{vcd:Cohen:1980}. In contrast to \code{assocplot()}, \pkg{vcd}'s \code{assoc()}
function scales to more than two variables (see Figure \ref{fig:residuals}):
\begin{Sinput}
> assoc(Titanic, compress = FALSE)
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<Residuals,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
assoc(Titanic, compress = FALSE)
@
\caption{Association plot for the Titanic data.}
\label{fig:residuals}
\end{center}
\end{figure}

\noindent The \code{compress} argument keeps distances between tiles
equal for better comparison.

So far, we have visualized the full table. Parts of the data can be
selected either using \code{margin.table()} on the input data, or the
formula interfaces. The following example compares the age
distribution among classes, given gender, and also shows how to 
use grid's viewport framework to put two plots besides 
each other:
\begin{Sinput}
> pushViewport(viewport(layout = grid.layout(ncol = 2)))

> pushViewport(viewport(layout.pos.col = 1))
> mosaic(~ Class + Age, data = Titanic[,"Male",,], 
+        margin = c(bottom = 0), sub = "Male", newpage = FALSE)
> popViewport()

> pushViewport(viewport(layout.pos.col = 2))
> mosaic(~ Class + Age, data = Titanic[,"Female",,], 
+        margin = c(bottom = 0), sub = "Female", newpage = FALSE)
> popViewport(2)
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<Variables,fig=TRUE,echo=FALSE,height=6,width=16,results=hide>>=
pushViewport(viewport(layout = grid.layout(ncol = 2)))

pushViewport(viewport(layout.pos.col = 1))
mosaic(~ Class + Age, data = Titanic[,"Male",,], 
       margin = c(bottom = 0), sub = "Male", newpage = FALSE,
       labeling_args = list(gp_labels = gpar(fontsize = 24), 
                            gp_varnames = gpar(fontsize = 24)),
       title_gp = gpar(fontsize = 35))
popViewport()

pushViewport(viewport(layout.pos.col = 2))
mosaic(~ Class + Age, data = Titanic[,"Female",,], 
       margin = c(bottom = 0), sub = "Female", newpage = FALSE,
       labeling_args = list(gp_labels = gpar(fontsize = 24), 
                            gp_varnames = gpar(fontsize = 24)),
       title_gp = gpar(fontsize = 35))
popViewport(2)
@
\caption{Age distribution among classes, given gender.}
\label{fig:parttable}
\end{center}
\end{figure}

\noindent note the use of the \code{margin} argument: it takes a
vector with up to four values whose unnamed components are recycled,
but ``overruled'' by the named arguments. Thus, in the example, only
the bottom margin is set to 0, leaving the default size of 2 lines
unchanged for the other margins. This idea applies to almost all
vectorized arguments in the strucplot framework (with
\code{split\_vertical} as a prominent exception).

Since mosaic displays are ``conditional plots'' by definition, we can
also use one single mosaic for stratified plots.
The formula interface of \code{mosaic()} allows the
specification of conditioning variables (see Figure \ref{fig:conditioning}):

\begin{Sinput}
> mosaic(~ Class + Age | Sex, data = Titanic, split = TRUE)
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]

\begin{center}
<<Conditioning,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(~ Class + Age | Sex, data = Titanic, split = TRUE)
@
\caption{Mosaic plot for the Titanic data (expected values).}
\label{fig:conditioning}
\end{center}
\end{figure}

\noindent Note, however, that the plots in the ``pseudo-strata'' are
distorted since they are not corrected for the marginal
distribution(s) of the conditioning variables. Indeed,
the \code{cotabplot()} function would do a much better job here (see
next section).

For both mosaic plots and association plots, 
the splitting of the tiles can be controlled using the
\code{split\_vertical} argument (default: alternating splits starting
with a vertical one). Sensible recycling rules apply. 
For compatibility with \code{mosaicplot()}, the \code{mosaic()} function
also allows the use of a \code{"direction"} argument taking a vector of \code{"h"} and
\code{"v"} characters (see Figures \ref{fig:split1} and \ref{fig:split2}):
\begin{Sinput}
> mosaic(Titanic, direction = "h")
> mosaic(Titanic, split_vertical = c(TRUE, TRUE, TRUE, FALSE))
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<split1,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(Titanic, direction = "h")
@
\caption{Mosaic plot for the Titanic data---alternative splitting (1).}
\label{fig:split1}
\end{center}
\end{figure}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<split2,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(Titanic, split_vertical = c(TRUE, TRUE, TRUE, FALSE),
                labeling_args = list(abbreviate = c(Sex = TRUE, Age = TRUE)))
@
\caption{Mosaic plot for the Titanic data---alternative splitting (2).}
\label{fig:split2}
\end{center}
\end{figure}

\section{Higher-level functions}

The last plot pretty much resembles a doubledecker plot, except for
the spacing. Since spacing is also modular in the strucplot framework,
the \code{doubledecker()} function was straightforward to implement:
it really is just a wrapper for \code{mosaic()}, setting the right
defaults. Figure \ref{fig:doubledecker} shows a doubledecker plot of
the Titanic data, explaining the probability of Survival by Age, given
Sex, given Class:
\begin{Sinput}
> doubledecker(Titanic)
> ## equivalent to:
> doubledecker(Survived ~ Class + Sex + Age, data = Titanic)
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<doubledecker,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
doubledecker(Titanic)
@
\caption{Doubledecker plot for the Titanic data.}
\label{fig:doubledecker}
\end{center}
\end{figure}

Another high-level function is \code{pairs.table()}, that is, the S3-method for objects
of class "table" of the \code{pairs()} generic. This function produces a matrix 
of strucplots (currently, association or mosaic plots) in the off-diagonal cells, and
the variable names (with, optionally, univariate statistics) in the diagonal cells. Figure
\ref{fig:pairs} shows mosaic plots visualizing mutual independence in the lower triangle,
association plots for the same in the upper triangle, and bar charts in the diagonal.

\begin{Sinput}
> pairs(Titanic, upper_panel = pairs_assoc)
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<pairs,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
pairs(Titanic, upper_panel = pairs_assoc)
@
\caption{Pairs plot for the Titanic data.}
\label{fig:pairs}
\end{center}
\end{figure}

\noindent Each cell's row and column define two variables $X$ and $Y$ used for the
specification of four different types of independence: `\code{pairwise}',
     `\code{total}', `\code{conditional}' and `\code{joint}'. The pairwise mosaic
     matrix shows bivariate marginal relations between $X$ and $Y$, collapsed over all
     other variables. The total independence mosaic matrix shows mosaic
     plots for mutual independence, i.e., for marginal and conditional
     independence among all pairs of variables. The conditional
     independence mosaic matrix shows mosaic plots for marginal
     independence of $X$ and $Y$ given all other variables. The joint independence
     mosaic matrix shows mosaic plots for joint independence of all
     pairs ($X$, $Y$) of variables from the others.

Since the matrix is symmetric, the upper and lower part can be used to display different 
types of independence models, or different strucplots (currently, only association and 
mosaic plots are supported). The available panel functions (\code{pairs\_assoc()} and 
\code{pairs\_mosaic()}) are just simple wrappers to \code{assoc()} and 
\code{mosaic()}. Obviously, seeing patterns in strucplot matrices 
becomes increasingly difficult with
higher dimensionality. Typically, therefore, this plot is used with a suitable 
residual-based shading (described in a separate vignette).

Another useful high-level plot is produced by \code{cotabplot()}: this function can be
used for plotting stratified strucplots in a lattice-like display. Whereas \code{pairs()} 
is conditioning on variables, \code{cotabplot()} is conditioning on variable \emph{levels}.
The plot in Figure \ref{fig:cotabplot} shows the Survival rate among the classes, given
all four level combinations of Sex and Age.

\begin{Sinput}
> cotabplot(~ Class + Survived | Sex + Age, data = Titanic)
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<cotabplot,fig=TRUE,echo=FALSE,height=6,width=12,results=hide>>=
cotabplot(~ Class + Survived | Sex + Age, data = Titanic)
@
\caption{cotabplot for the Titanic data.}
\label{fig:cotabplot}
\end{center}
\end{figure}

\section{Technical Notes on Panel Functions}

All strucplot components (including the panel functions) 
are parameterized through the use of generating functions, that is,
functions with all parameters in their signature returning a function
to be called by \code{strucplot()}. This returned function has access
to all parameters passed to the generating function thanks to 
lexical scope. The idea is to use module-specific `code
factories' that generate parameterized workhorses rather than
struggling with monolithic functions with endless parameter lists. As
a simple example, consider \code{struc\_mosaic()} and \code{struc\_assoc},
the generating functions for the mosaic and the association plots:

\begin{smallexample}
struc_mosaic <- function(zero_size = 0.5)
  function(residuals, observed, expected = NULL, spacing, gp, split_vertical) {
  \dots
  }

struc_assoc <- function(compress = TRUE, xlim = NULL, ylim = NULL,
                        yspace = unit(0.5, "lines"), xscale = 0.9)
  function(residuals, observed = NULL, expected, spacing, gp, split_vertical) {
  \dots
  }
\end{smallexample}

\noindent Both functions have specific parameters in their signature,
but both return a function with a standardized interface that
subsequently is called by \code{strucplot} to produce the actual
plot. As can be seen, the only information \code{strucplot()} passes
to the panel functions are residuals, observed and expected
values, a structure containing the between-tile spacing information,
an object of class \code{"gpar"} for the tiles' graphical parameters,
and a logical vector with the splitting information for each
dimension. The workhorse function, of course, is free to use only
part or all of this information. High-level interface functions such
as \code{mosaic()} and \code{assoc{}} call \code{strucplot()} by
specifying the corresponding function along with the specific
parameters. Now, there a several ways of doing the same thing, as
illustrated by the following equivalent calls:

\begin{Sinput}
> strucplot(Titanic, panel = struc_mosaic(zero_size = 1))
> strucplot(Titanic, panel = struc_mosaic, panel_args = list(zero_size = 1))
> strucplot(Titanic, panel = myworkhorse)
\end{Sinput}

\noindent where \code{myworkhorse} could be, e.g., a function with
hard-wired parameters, but with an interface complying with the
workhorse signature. To enable \code{strucplot()} to distinguish
between generating functions and workhorse functions, the generating
functions must be of class \code{"panel\_generator"}.

All panel functions are supposed to produce conditional hierarchical
plots by the means of nested viewports, corresponding to the provided splitting
information. Thus, at the end of the plotting, each tile is associated with
a particular viewport. Each of those viewports has to be
conventionally named, enabling other strucplot modules, in particular the
labeling functions, to access specific tiles after they have been
plotted. Note that the viewport tree is popped by default. The
following example shows how to access parts of the plot after it has
been drawn (see Figure \ref{fig:afterplot}):

\begin{Sinput}
> mosaic(~ Sex + Class, data = Titanic, pop = FALSE)
> current.vpTree() ## output not shown

> seekViewport("cell..Sex.Male")
> grid.rect(gp = gpar(col = "red", lwd = 4))

> seekViewport("cell..Sex.Male..Class.Crew")
> grid.circle(r = 0.2, gp = gpar(fill = "cyan"))
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<afterplot,fig=TRUE,echo=FALSE,height=6,width=12,results=hide>>=
mosaic(~ Sex + Class, data = Titanic, pop = FALSE)

seekViewport("cell..Sex.Male")
grid.rect(gp = gpar(col = "red", lwd = 4))

seekViewport("cell..Sex.Male..Class.Crew")
grid.circle(r = 0.2, gp = gpar(fill = "cyan"))
@
\caption{Adding elements to a mosaic plot after drawing.}
\label{fig:afterplot}
\end{center}
\end{figure}

\noindent The naming convention for the viewports is:
\code{cell..\emph{[Variable 1]}.\emph{[Level 1]}..\emph{[Variable 2]}.\emph{[Level 2]}} \dots. Clearly, these names depend on the splitting.

In addition to the viewports, the main graphical elements get names
following a similar construction method. This allows to change
graphical parameters of plot elements after the plotting (see Figure \ref{fig:changeplot}):

\begin{Sinput}
> assoc(Titanic)
> getNames() ## output not shown
> grid.edit("rect..Class.1st..Age.Adult..Sex.Female..Survived.Yes", 
            gp = gpar(fill = "red"))
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<changeplot,fig=TRUE,echo=FALSE,height=6,width=12,results=hide>>=
assoc(Titanic)
getNames() ## output not shown
grid.edit("rect..Class.1st..Age.Adult..Sex.Female..Survived.Yes", 
          gp = gpar(fill = "red"))
@
\caption{Changing graphical parameters of elements after drawing.}
\label{fig:changeplot}
\end{center}
\end{figure}

\begin{thebibliography}{}

\bibitem[Cohen, 1980]{vcd:Cohen:1980}
Cohen, A. (1980).
\newblock On the graphical display of the significant components in a two-way
  contingency table.
\newblock {\em Communications in Statistics---Theory and Methods},
  A9:1025--1041.

\bibitem[Hartigan and Kleiner, 1984]{vcd:Hartigan+Kleiner:1984}
Hartigan, J. and Kleiner, B. (1984).
\newblock A mosaic of television ratings.
\newblock {\em The American Statistician}, 38:32--35.

\end{thebibliography}

\end{document}