### 2005-08-03 trunk/vcd/paper/JCGS.Rnw
finished Sections 3.1 and 3.2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@427 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[10pt,a4paper]{article}

%% Packages:
\usepackage{graphicx,color,a4wide,thumbpdf,hyperref}

%% \usepackage{Sweave} is essentially
\RequirePackage[T1]{fontenc}
\RequirePackage{ae,fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}

%% BibTeX
\usepackage[authoryear,round,longnamesfirst]{natbib}
\bibliographystyle{asa}
\bibpunct{(}{)}{;}{a}{}{,}

%% new commands
\let\code=\texttt
\let\proglang=\textsf
\newcommand{\pkg}[1]{{\normalfont\fontseries{b}\selectfont #1}}
\newcommand{\given}{\, | \,}

%% Colors
\definecolor{Red}{rgb}{0.7,0,0}
\definecolor{Blue}{rgb}{0,0,0.8}
\definecolor{hellgrau}{rgb}{0.55,0.55,0.55}

%% mimic JCGS style
\renewcommand{\section}{\secdef \mysec \mysecnn}
\newcommand{\mysec}[2][default]{\vspace{1.3\baselineskip}%
  \pdfbookmark[1]{#1}{Section.\thesection.#1}%
  \refstepcounter{section}%
  {\begin{center} \large \bf \thesection. \uppercase{#1} \end{center}}%
  \vspace{0.7\baselineskip}}
\newcommand{\mysecnn}[1]{\vspace{1.3\baselineskip}%
  {\begin{center}\large \bf \uppercase{#1} \end{center}}%
  \vspace{0.7\baselineskip}}

\renewcommand{\subsection}{\secdef \mysubsec \mysubsecnn}
\newcommand{\mysubsec}[2][default]{\vspace{1.2\baselineskip}%
  \pdfbookmark[2]{#1}{Subsection.\thesubsection.#1}%
  \refstepcounter{subsection}%
  \noindent%
  {\large \textbf{\thesubsection{} {\sc#1}}} \vspace{.7\baselineskip}}
\newcommand{\mysubsecnn}[1]{\vspace{1.2\baselineskip}%
  \noindent%
  {\large \textbf{{\sc #1}}} \vspace{.7\baselineskip}}
\renewcommand{\abstractname}{}
\renewcommand{\refname}{REFERENCES}

%% Hyperref
\hypersetup{%
  pdftitle = {Residual-based Shadings for Visualizing (Conditional) Independence},
  pdfsubject = {Residual-based Shadings},
  pdfkeywords = {contingency tables, association plots, mosaic plots, HCL colors, 
    HSV colors, conditional inference, permutation tests},
  pdfauthor = {Achim Zeileis, David Meyer, Kurt Hornik},
  %% change colorlinks to false for pretty printing
  colorlinks = {true},
  linkcolor = {Blue},
  citecolor = {Blue},
  urlcolor = {Red},
  hyperindex = {true},
  linktocpage = {true},
}

\begin{document}

\begin{center}
\textbf{\LARGE Residual-based Shadings for Visualizing\\[.15cm] (Conditional) Independence}

\vspace{0.5cm}
{\large Achim \textsc{Zeileis}, David \textsc{Meyer}, and Kurt \textsc{Hornik}}

%% \vspace{0.3cm}
%% \textit{\large Wirtschaftsuniversit\"at Wien, Augasse 2--6, 1090 Wien, Austria}\\[0.14cm]
%% {\large \textit{E-mail:} \texttt{Firstname.Lastname@wu-wien.ac.at}}
\end{center}

\SweaveOpts{engine=R,eps=FALSE}
\setkeys{Gin}{width=0.8\textwidth}

<<load_data,echo=FALSE,results=hide>>=
library(vcd)

data(Arthritis)
art <- xtabs(~ Treatment + Improved, data = Arthritis, subset = Sex == "Female")

data(UCBAdmissions)
ucb <- aperm(UCBAdmissions)

myseed <- 1071
@

<<inference,echo=FALSE,results=hide>>=
if(FALSE) {
  set.seed(myseed)
  art_max <- coindep_test(art, n = 5000)
  set.seed(myseed)
  art_chisq <- coindep_test(art, n = 5000, indepfun = function(x) sum(x^2))

  set.seed(myseed)
  art_gp <- shading_max(art, n = 5000)

  set.seed(myseed)
  ucb_doublemax <- coindep_test(ucb, "Dept", n = 5000)
  set.seed(myseed)
  ucb_maxChisq <- coindep_test(ucb, "Dept", n = 5000, indepfun = function(x) sum(x^2))
  set.seed(myseed)
  ucb_sumChisq <- coindep_test(ucb, "Dept", n = 5000, indepfun = function(x) sum(x^2), aggfun = sum)

  set.seed(myseed)
  ucb_cotab <- cotab_coindep(ucb, condvars = "Dept", type = "assoc", n = 5000)

  save(art_max, art_chisq, art_gp,
       ucb_doublemax, ucb_maxChisq, ucb_sumChisq, ucb_cotab,
       file = "JCGS-inference.rda")
} else {
  load("JCGS-inference.rda")
}
@

<<auxiliary_functions,echo=FALSE,results=hide>>=
stat <- function(x, digits = 3) as.vector(round(x$statistic, digits = digits))
pval <- function(x, digits = 6) {
  oop <- options()
  on.exit(oop)
  options(scipen = 1000)
  as.vector(round(x$p.value, digits = digits))
}

hue.slice <- function(hue, grid.n = 100, type = c("HCL", "HSV"), plot = TRUE, fixup = FALSE)
{
  type <- match.arg(type)
  if(type == "HCL") {
    chroma = seq(0, 100, length = grid.n)
    luminance = seq(0, 100, length = grid.n)
    nc <- length(chroma)
    nl <- length(luminance)
    color.slice <- outer(chroma, luminance, function(y, x) hcl(hue, x, y, fixup = fixup))
    xlab <- "chroma"
    ylab <- "luminance"
    main <- paste("hue =", round(hue, digits = 0))
  } else {
    chroma = seq(0, 1, length = grid.n)
    luminance = seq(0, 1, length = grid.n)
    nc <- length(chroma)
    nl <- length(luminance)
    color.slice <- outer(chroma, luminance, function(y, x) hsv(hue, x, y))
    xlab <- "saturation"
    ylab <- "value"
    main <- paste("hue =", round(hue, digits = 3))
  }
  if(plot) {
    plot(0.5, 0.5, xlim = range(chroma), ylim = range(luminance), type = "n", axes = FALSE,
         xlab = xlab, ylab = ylab, yaxs = "i", xaxs = "i", main = main)
    for(i in 1:(nc-1)) {
      rect(chroma[i], luminance[-nl], chroma[i] + 100/(nc-1), luminance[-1], border = color.slice[,i+1], col = color.slice[,i+1])
    }
    axis(1)
    axis(2)
    box()
  }
  colnames(color.slice) <- chroma
  rownames(color.slice) <- luminance
  attr(color.slice, "type") <- type
  class(color.slice) <- "slice"
  invisible(color.slice)
}
@

\begin{abstract}
Visualizing independence by association and mosaic plots.

Enhancements: diverging color palette based on HCL color space
combined with visualization of the result of a significance test.

Extensions to visualizing conditional independence.\\[0.3cm]

\textbf{Key Words:} Association plots; Conditional inference; 
  Contingency tables; HCL colors; HSV colors; Mosaic plots; Permutation tests.
\end{abstract}

\section{Introduction} \label{sec:introduction}

independence in 2-way tables

association and mosaic displays

colors and color spaces
    because the three perceptual dimensions (hue, lightness, saturation)
    are not properly mapped to the three dimensions of the color space
    and hence are confounded \citep{vcd:Brewer:1999}

somewhere:
mention Augsburg stuff, visualization of log-linear models via mosaics
\citep{vcd:Theus+Lauer:1999,vcd:Hofmann:2003,vcd:Hofmann:2001}.

Association plots as residual plots for log-linear models
\citep{vcd:Meyer+Zeileis+Hornik:2003}, especially
in coplot or trellis layout.

\section{Independence in 2-way Tables}
\label{sec:2way}

In this section, the basic tools for testing and visualizing 
independence in 2-way tables are briefly reviewed. 
For illustration, a data set about treatment and improvement of
patients with rheumatoid arthritis from \cite{vcd:Koch+Edwards:1988}
is used. The data set is also discussed in \cite{vcd:Friendly:2000}
and the subset of the $59$ female patients from the study is given in
Table~\ref{tab:art}.

\subsection{Tests}
\label{sec:tests}

To fix notations, we consider a 2-way contingency table with cell
frequencies $\{n_{ij}\}$ for $i = 1, \dots, I$ and $j = 1, \dots, J$
and row and column sums $n_{i+} = \sum_i n_{ij}$ and $n_{+j} = \sum_j n_{ij}$
respectively.
For convenience the number of observations is denoted $n = n_{++}$.
Given an underlying distribution with theoretical cell probabilities
$\pi_{ij}$, the null hypothesis of independence of the two categorical
variables can be formulated as
\begin{equation} \label{eq:H0}
H_0: \; \pi_{ij} \quad = \quad \pi_{i+} \pi_{+j}.
\end{equation}

The expected cell frequencies in this model are
$\hat n_{ij} = n_{i+}n_{+j}/n$. The probably best known and most used
measure of discrepancy between observed and expected values are the Pearson residuals
\begin{equation} \label{eq:Pearson}
r_{ij} \quad = \quad \frac{n_{ij} - \hat n_{ij}}{\sqrt{\hat n_{ij}}}.
\end{equation}
The most convenient way to aggregate the $I \times J$ residuals
to one test statistic is their squared sum
\begin{equation} \label{eq:ChiSquare}
X^2 \quad = \quad \sum_{i, j}r_{ij}^2,
\end{equation}
because this is known to have an unconditional limiting $\chi^2$ distribution with $(I-1)(J-1)$
degrees of freedom under the null hypothesis.
This is the well-known $\chi^2$ test which is typically introduced at the very beginning
of the chapter about independence in 2-way tables in statistics textbooks
\citep[see e.g.,][]{vcd:Agresti:2002}.

\begin{table}[t!]
\begin{center}
\caption{\label{tab:art} Treatment and improvement among $59$ patients
with rheumatoid arthritis.}
\begin{tabular}{rr|rrr}
                   &         & \multicolumn{3}{|l}{\textbf{Improved}} \\
                   &         & None & Some & Marked \\ \hline
\textbf{Treatment} & Placebo &   19 &    7 &      6 \\
                   & Treated &    6 &    5 &     16
\end{tabular}
\end{center}
\end{table}

But the sum of squares is not the only plausible way of capturing deviations
from $0$ in the Pearson residuals.
There are many other conceivable functionals $\lambda(\cdot)$ which
lead to reasonable test statistics $\lambda(\{r_{ij}\})$: one which is particularly
suitable for identifying the cells which cause the dependence (if any) is
the maximum of the absolute values
\begin{equation} \label{eq:MaxAbs}
M \quad = \quad \max_{i, j} |r_{ij}|.
\end{equation}
Given a critical value $c_\alpha$ for this test statistic, all residuals whose
absolute value exceeds $c_\alpha$ violate the hypothesis of indendence at significance level
$\alpha$ \cite[][Ch.~7]{vcd:Mazanec+Strasser:2000}. Thus, the interesting cells causing
the dependence can easily be identified.

Furthermore, an important reason for using the unconditional limiting distribution
for the $X^2$ statistic (\ref{eq:ChiSquare}) was the closed form result
for the distribution. Recently, with the improving perfomance of computers,
conditional inference (or permutation tests)---carried out either by simulation
or by computation of the (asymptotic) permutation distribution---have been receiving
increasing attention \citep{vcd:Pesarin:2001,vcd:Strasser+Weber:1999}. 
For testing the independence hypothesis (\ref{eq:H0}), using a permutation test is particularly
intuitive due to the permutation invariance (given row and column sums) of this problem.
Consequently, all results in this paper are based on conditional inference performed
by simulating from the permutation distribution of test statistics of type
$\lambda(\{r_{ij}\})$.

Other measures of discrepancy (e.g., deviance residuals) could, of course, also be used
instead of the the Pearson residuals $\{r_{ij}\}$. But as the ideas discussed here
extend straightforwardly to that situation, we do not discuss it in detail.

For the arthritis data from Table~\ref{tab:art}, both tests indicate a clearly significant
dependence of improvement on treatment: the sum-of-squares statistic from Equation~(\ref{eq:ChiSquare})
is $X^2 = \Sexpr{stat(art_chisq)}$ with a $p$ value of $p = \Sexpr{pval(art_chisq)}$,
and the maximum statistion from Equation~(\ref{eq:MaxAbs}) is $M = \Sexpr{stat(art_max)}$
with a $p$ value of $p = \Sexpr{pval(art_max)}$. Both $p$ values have been computed
by drawing a sample of size $5000$ from the permutation distribution under independence.

\subsection{Visualizations} \label{sec:visualization}

Two well-established visualization techniques for independence in 2-way tables are
mosaic plots and association plots. Both are suitable to bring out
departures of an observed table $\{n_{ij}\}$ from the expected table $\{\hat n_{ij}\}$
in a graphical way. The latter focuses on the visualization of the Pearson residuals
$r_{ij}$ (under independence) while the latter primarily displays the observed frequencies
$n_{ij}$.

\emph{Mosaic plots}
\citep{vcd:Hartigan+Kleiner:1981} can be seen as an extension of grouped
bar charts where width and height of the bars show the relative
frequencies of the two variables: a mosaic plot simply
consists of a collection of tiles whose
sizes are proportional to the observed cell frequencies as shown
in the left panel of Figure~\ref{fig:classic}. A rectangle corresponding
to $100$ percent of the observations is first split horizontally with respect to 
the treatment frequencies and then vertically with respect to the conditional
improvement frequencies. This shows that there have been more placebo than treated 
patients with no improvement and vice versa for marked improvement.
This strategy of splitting with respect to conditional frequencies given all
previous variables can also directly be used for visualizing multi-way tables
\citep[see][for an overview of how to construct and read mosaic displays]{vcd:Hofmann:2003}.

\emph{Association plots}
\citep{vcd:Cohen:1980} visualize the table of Pearson residuals:
each cell is represented by a rectangle that has (signed) height
proportional to the corresponding Pearson residual $r_{ij}$ and width proportional to the
square root of the expected counts $\sqrt{\hat n_{ij}}$. Thus, the area is proportional to
the raw residuals $n_{ij} - \hat n_{ij}$. The association plot for the arthritis
data is shown in the right panel of Figure~\ref{fig:classic} which leads to the
same interpretation as the mosaic plot: there are more placebo patients with no 
improvement and fewer with marked improvement than expected under independence---vice
versa for the treated patiens.

\setkeys{Gin}{width=\textwidth}
\begin{figure}[htbp]
\begin{center}
<<classic_plots,fig=TRUE,echo=FALSE,height=6,width=11,results=hide>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
pushViewport(viewport(layout.pos.col=1, layout.pos.row=1))
mosaic(art, newpage = FALSE, margins = c(2.5, 4, 2.5, 3))
popViewport()
pushViewport(viewport(layout.pos.col=2, layout.pos.row=1))
assoc(art, newpage = FALSE, margins = c(5, 2, 5, 4))
popViewport(2)
@
\caption{Classic mosaic and association plot for the Arthritis data.}
\label{fig:classic}
\end{center}
\end{figure}

\section{Residual-based Shadings} \label{sec:shadings}

Colors are commonly used to enhance mosaic and association plots. To integrate
a visualization of the Pearson residuals $\{r_{ij}\}$ into the mosaic display---which
in its `raw' version only visualizes the observed frequencies $\{n_{ij}\}$---\cite{vcd:Friendly:1994}
suggested a residual-based shading for the mosaic tiles that can also be applied
to the rectangles in association plots \citep{vcd:Meyer+Zeileis+Hornik:2003}. 
In this section, we first briefly review the \cite{vcd:Friendly:1994} shading, 
before we suggest different colors and a combination of visualization and significance
testing to extend these residual-based shadings.

\subsection{Friendly Shading}

The extensions of \cite{vcd:Friendly:1994} to mosaic plots provide a
substantial improvement of the original mosaic plots and enhanced them
from a plot for contingency tables to a visualization technique for
log-linear models---and thus also for independence problems including
2-way tables as the simplest special case.

The idea is to use a color coding for the mosaic tiles that visualizes the
sign and absolute size of each residual $r_{ij}$: Cells corresponding to small
residuals ($|r_{ij}| < 2$) are shaded white. Cells with medium sized residuals
($2 \le |r_{ij}| < 4$) are shaded in light blue and light red for positive negative
residuals respectively. Cells with large residuals ($|r_{ij}| \ge 4$) are shaded
with a fully saturated blue and red respectively. Mosaic plots enhanced by this 
shading can thus also bring out departures from independence (or other log-liner 
models in multi-way tables) graphically and visualize patterns of dependence.
The heuristic for choosing the cut offs $2$ and $4$ is that the Pearson residuals
are approximately standard normal which implies that the highlighted cells are those with
residuals \emph{individually} significant at approximately the $\alpha = 0.05$ and $\alpha = 0.0001$
level. But the main purpose of the shading is not to visualize significance but the \emph{pattern} of
deviation from independence \citep[p.~109]{vcd:Friendly:2000}.

In addition to the shading of the rectangles themselves, the Friendly shading also
encompasses a choice of line type and line color of the borders of the rectangles
with similar ideas as described above. As both mosaic and association plots are
area-proportional visualization techniques, we focus on area shadings and always 
use solid black borders throughout this paper, but the extensions suggested in the
following could also be applied to control line type and color.

\subsection{Colors}

The way the (light) blue and red colors are chosen differs somewhat between
various implementations of Friendly mosaic plots: In his original 
\proglang{SAS} implementation \citep[see][]{vcd:Friendly:2000}, Michael Friendly
uses colors from a palette based on HLS (Hue--Luminance--Saturation) color space.
The implementation in the \proglang{R} system for statistical computing
\citep[\url{http://www.R-project.org/}]{vcd:R:2005} employs colors from
HSV (Hue--Saturation--Value) space. The latter is a very common implementation of
colors in many computer packages and makes the generation of the Friendly shading
very simple.

The HSV space originally looks like a regular cone with black at its
peak (zero value) and full color wheels for different saturations at the other end,
centered around a white (full value). Type and amount of color are controlled by hue
and saturation respectively. Typically, polar coordinates that all range in the 
unit interval are used in this space, such that it looks like a regular cylinder.
For generating colors in the Friendly shading, the following strategy is used:
The hue $h$ codes the sign of the residuals---$h = 0$ (red hue) is used for negative residuals,
$h = 2/3$ (blue hue) for positive residuals. The absolute size of the residuals
is then coded by the saturation $s$ which is set to $0$, $0.5$ and $1$ respectively
for small/medium/large residuals respectively. The value is always fixed at $v = 1$.
This is also depicted in the upper panel of Figure~\ref{fig:shadingHSVHCL} which shows
to slides from the HSV cylinder side to side. The plot shows the saturation/value plane
for the given hues $h = 0$ and $h = 2/3$. The significant color palette shows the
colors used for the Friendly shading when the residuals are increasing from left to right.
The set of non-significant colors will be explained in the next section.

Although this HSV-based shading is already very useful for enhancing mosaic and
association plots and although HSV is very commonly available implementation of
color spaces, HSV colors in general and the Friendly shading in particular have
a number of disadvantages. Most importantly, HSV colors are not perceptually
uniform because the three HSV dimensions map only poorly to the three
perceptual dimensions of the human visual system  \citep{vcd:Brewer:1999,vcd:Ihaka:2003}.
Consequently, the the HSV dimensions are confounded, e.g., saturation
is not uniform across different hues. A fully saturated blue $(2/3, 1, 1)$ is
much darker as a fully saturated red $(0, 1, 1)$ or green $(1/3, 1, 1)$.
This makes it more difficult for the human eye to judge the size of shaded areas
and can therefore lead to color-caused optical illusions when used in statistical graphs
\citep{vcd:Cleveland+McGill:1983}. Furthermore, flashy fully saturated HSV colors are
good for drawing attention to a plot, but hard to look at for a longer time \citep{vcd:Ihaka:2003}
which makes HSV-shaded graphics harder to interpret. Finally, white is employed as
the neutral color for small residuals in the Friendly shading, but typically
grey is found to convey neutrality or uninterestingness much better than white does
\citep{vcd:Brewer:1999}.

Alternative ways to choose colors have been available for a long time, but have
been only slowly adopted for implementations of colors in computer packages in 
general and for shading in statistical graphs in particular. The idea of using
perceptually based colors that are `in harmony' go back to \cite{vcd:Munsell:1905} 
who introduced a color notation for balanced colors. Based on similar principles, Cynthia
Brewer and co-workers suggested different types of palettes (qualitative/sequential/diverging)
and provided the online tool \pkg{ColorBrewer.org} \citep{vcd:Harrower+Brewer:2003} for
selecting an appropriate palette for a specific problem. Furthermore, the
Commission Internationale de l'\'Eclairage (CIE) introduced the two perceptually
based color spaces CIELAB and CIELUV where the latter is typically preferred for 
emissive color technologies such as computer displays. \cite{vcd:Ihaka:2003} discusses
how CIELUV colors can be used for choosing colors for statistical graphics such as barplots.
By taking polar coordinates in CIELUV space, it is called HCL (Hue--Chroma--Luminance) space
and qualitative palettes can easily be chosen by using a range of hues for fixed values
of chroma and luminance. Such colors are always balanced towards the same grey 
and thus do not have the problem of varying saturations as for the HSV colors.

Here, we discuss how similar ideas can be used for deriving a diverging HCL palettes that 
provide a suitable translation of the ideas from the Friendly shading to perceptually
uniform HCL colors. 
The HCL space looks like a double cone with black (zero luminance) at one end and white
at the other (full luminance). In its middle, there is a full color wheel for different
values of chroma (that controls the colorfulness). Unfortunately, the HCL space is not completely
regular as the HSV space is and consequently its dimensions cannot be standardized to 
unit intervals: the hue ranges in $[0, 360]$ degrees and chroma and luminance both in $[0, 100]$
percent. But not all combinations $(h, c, l)$ yield valid HCL colors and the admissable combinations
of $c$ and $l$ vary across different hues $h$. But for the problem of constructing
a diverging palette, this problem can easily be overcome as we just need two different
hues (a `negative' and a `positive' hue) and hence choose two hues that correspond to 
similar shapes in the chroma/luminance plane. The lower panel of Figure~\ref{fig:shadingHSVHCL}
shows two such planes side to side for the hues $h = 0$ and $h = 260$. To obtain a 
sequence of colors with the same properties as the Friendly shading, the palette starts
at a fully saturated red $(0, 100, 50)$, goes via a neutral color, ends at a fully saturated
blue $(260, 100, 50)$, and uses linear interpolation in between. Instead of using white
$(0, 0, 100)$ as the neutral color, a light grey $(0, 0, 90)$ is employed as motivated above.

This diverging palette (see the `significant' colors in Figure~\ref{fig:shadingHSVHCL},
the `non-significant' colors are again discussed in the following section)
uses both chroma, i.e., the colorfulness, and luminance, i.e., 
the amount of grey, to code the absolute size of the quantity visualized, i.e., the
Pearson residuals when applied to the independence problem. By changing the neutral
color or by changing the maximum chroma respectively, this can be changed to using only
chroma or luminance for this purpose, but using both seems to be a very effective way
of visualization.

The advantages of using this HCL shading scheme instead of the HSV scheme is that
it is based on a perceptually based color space that provides uniform saturations
for different hues and is device inependent.

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[p]
\begin{center}
<<shading_HSV,fig=TRUE,echo=FALSE,height=7,width=8>>=
## generate colors
hue23 <- hue.slice(2/3, grid.n = 101, plot = FALSE, type = "HSV")
hue0 <- hue.slice(0, grid.n = 101, plot = FALSE, type = "HSV")
saturation <- as.numeric(colnames(hue23))
value <- as.numeric(rownames(hue23))

## select those with value >= 0.5
hue23 <- hue23[value >= .5, ]
hue0 <- hue0[value >= .5, ]
value <- value[value >= .5]
nl <- nrow(hue23)
nc <- ncol(hue23)

## plot 2 slides from HSV space
plot(0.5, 0.5, xlim = c(-1, 1), ylim = c(0, 1), type = "n", axes = FALSE,
       xlab = "", ylab = "", yaxs = "i", xaxs = "i", main = "")
for(i in 1:(nc-1)) {
  rect(saturation[i], value[-nl], saturation[i] + 1/(nc-1), value[-1], border = hue23[,i+1], col = hue23[,i+1])
}
for(i in 1:(nc-1)) {
  rect(-saturation[i], value[-nl], -(saturation[i] + 1/(nc-1)), value[-1], border = hue0[,i+1], col = hue0[,i+1])
}
axis(2, at = c(50, 75, 100)/100, labels = c(0.5, 0.75, 1))
axis(4, at = c(50, 75, 100)/100, labels = c(0.5, 0.75, 1))
axis(3, at = -4:4*.25, labels=c(4:0*.25, 1:4*.25))
mtext(c("hue = 0", "hue = 2/3"), side = 3, at = c(-.5, .5), line = 3, cex = 1.2)
mtext("saturation", side = 3, at = 0, line = 2)
mtext("value", side = 2, at = .75, line = 2)
mtext("value", side = 4, at = .75, line = 2)
lines(c(-1, 1), c(.5, .5))

## significant colors
rect(-1, 0.95, -.90, 1, col = hsv(0, 1, 1))
rect(-0.45, 0.95, -.55, 1, col = hsv(0, 0.5, 1))
rect(-.05, .95, .05, 1, col = hsv(2/3, 0, 1))
rect(0.45, 0.95, .55, 1, col = hsv(2/3, 0.5, 1))
rect(.90, .95, 1, 1, col = hsv(2/3, 1, 1))

text(-1, .33, "significant", pos = 4, cex = 1.2)
rect(-1, .20, -.80, .30, col = hsv(0, 1, 1))
rect(-.40, .20, -0.6, .30, col = hsv(0, 0.5, 1))
rect(-.20, .20, 0, .30, col = hsv(0, 0, 1))
rect(0, .20, .20, .30, col = hsv(2/3, 0, 1))
rect(0.4, .20, .60, .30, col = hsv(2/3, .5, 1))
rect(.80, .20, 1, .30, col = hsv(2/3, 1, 1))

lines(c(-.9, -.55), c(0.975, .975), lty = 2)
lines(c(-.45, -.05), c(0.975, .975), lty = 2)
lines(c(.45, .05), c(0.975, .975), lty = 2)
lines(c(.9, .55), c(0.975, .975), lty = 2)

## non-significant colors
rect(-1, 0.5, -.90, 0.55, col = hsv(0, 1, 0.5))
rect(-0.45, 0.5, -.55, 0.55, col = hsv(0, 0.5, 0.5))
rect(-.05, .5, .05, 0.55, col = hsv(2/3, 0, 0.5))
rect(0.45, 0.5, .55, 0.55, col = hsv(2/3, 0.5, 0.5))
rect(.90, .5, 1, 0.55, col = hsv(2/3, 1, 0.5))

text(-1, .13, "non-significant", pos = 4, cex = 1.2)
rect(-1, 0, -.80, .10, col = hsv(0, 1, 0.5))
rect(-.60, 0, -.4, .10, col = hsv(0, 0.5, 0.5))
rect(-.20, 0, 0, .10, col = hsv(0, 0, 0.5))
rect(0, 0, .20, .10, col = hsv(2/3, 0, 0.5))
rect(0.4, 0, .60, .1, col = hsv(2/3, .5, 0.5))
rect(.80, 0, 1, .10, col = hsv(2/3, 1, 0.5))

lines(c(-.9, -.55), c(0.525, .525), lty = 2)
lines(c(-.45, -.05), c(0.525, .525), lty = 2)
lines(c(.45, .05), c(0.525, .525), lty = 2)
lines(c(.9, .55), c(0.525, .525), lty = 2)
@

<<shading_HCL,fig=TRUE,echo=FALSE,height=7,width=8>>=
## generate colors
hue260 <- hue.slice(260, grid.n = 101, plot = FALSE)
hue360 <- hue.slice(360, grid.n = 101, plot = FALSE)
mychroma <- as.numeric(colnames(hue260))
luminance <- as.numeric(rownames(hue260))

## select those with lumincance >= 50
hue260 <- hue260[luminance >= 50, ]
hue360 <- hue360[luminance >= 50, ]
luminance <- luminance[luminance >= 50]
nc <- ncol(hue260)
nl <- nrow(hue260)

## plot 2 slides from HCL space
plot(0.5, 0.5, xlim = c(-100, 100), ylim = c(0, 100), type = "n", axes = FALSE,
       xlab = "", ylab = "", yaxs = "i", xaxs = "i", main = "")
for(i in 1:(nc-1)) {
  rect(mychroma[i], luminance[-nl], mychroma[i] + 100/(nc-1), luminance[-1], border = hue260[,i+1], col = hue260[,i+1])
}
for(i in 1:(nc-1)) {
  rect(-mychroma[i], luminance[-nl], -(mychroma[i] + 100/(nc-1)), luminance[-1], border = hue360[,i+1], col = hue360[,i+1])
}
axis(2, at = c(50, 70, 90, 100), labels = c(50, 70, 90, 100))
axis(4, at = c(50, 70, 90, 100), labels = c(50, 70, 90, 100))
axis(3, at = -4:4*25, labels=c(4:0*25, 1:4*25))
mtext(c("hue = 0", "hue = 260"), side = 3, at = c(-50, 50), line = 3, cex = 1.2)
mtext("chroma", side = 3, at = 0, line = 2)
mtext("luminance", side = 2, at = 75, line = 2)
mtext("luminance", side = 4, at = 75, line = 2)
lines(c(-100, 100), c(50, 50))

## significant colors
rect(-100, 47.5, -90, 52.5, col = hcl(0, 100, 50))
rect(-55, 67.5, -45, 72.5, col = hcl(0, 50, 70))
rect(-5, 95, 5, 100, col = hcl(260, 0, 100))       ## grey vs. white
rect(-5, 87.5, 5, 92.5, col = hcl(260, 0, 90))     ## grey vs. white
rect(45, 67.5, 55, 72.5, col = hcl(260, 50, 70))
rect(90, 47.5, 100, 52.5, col = hcl(260, 100, 50))

text(-100, 33, "significant", pos = 4, cex = 1.2)
rect(-100, 20, -80, 30, col = hcl(0, 100, 50))
rect(-60, 20, -40, 30, col = hcl(0, 50, 70))
rect(-20, 20, 0, 30, col = hcl(0, 0, 90))       
rect(0, 20, 20, 30, col = hcl(260, 0, 90))
#white# rect(-20, 20, 0, 30, col = hcl(0, 0, 100))
#white# rect(0, 20, 20, 30, col = hcl(260, 0, 100))
rect(40, 20, 60, 30, col = hcl(260, 50, 70))
rect(80, 20, 100, 30, col = hcl(260, 100, 50))

lines(c(-45, -5), c(72.5, 87.5), lty = 2)
lines(c(45, 5), c(72.5, 87.5), lty = 2)
lines(c(-95, -55), c(52.5, 67.5), lty = 2)
lines(c(95, 55), c(52.5, 67.5), lty = 2)

## non-significant colors
rect(-25, 47.5, -15, 52.5, col = hcl(0, 20, 50))
rect(-15, 67.5, -5, 72.5, col = hcl(0, 10, 70))
rect(5, 67.5, 15, 72.5, col = hcl(260, 10, 70))
rect(25, 47.5, 15, 52.5, col = hcl(260, 20, 50))

text(-100, 13, "non-significant", pos = 4, cex = 1.2)
rect(-60, 0, -40, 10, col = hcl(0, 20, 50))
rect(-40, 0, -20, 10, col = hcl(0, 10, 70))
rect(-20, 0, 0, 10, col = hcl(0, 0, 90))
rect(0, 0, 20, 10, col = hcl(260, 0, 90))
rect(20, 0, 40, 10, col = hcl(260, 10, 70))
rect(40, 0, 60, 10, col = hcl(260, 20, 50))

lines(c(-18.75, -11.25), c(52.5, 67.5), lty = 2)
lines(c(-8.75, -1.25), c(72.5, 87.5), lty = 2)
lines(c(18.75, 11.75), c(52.5, 67.5), lty = 2)
lines(c(8.75, 1.25), c(72.5, 87.5), lty = 2)
@
\caption{Extended shading in HSV (upper) and HCL space (lower).}
\label{fig:shadingHSVHCL}
\end{center}
\end{figure}

\subsection{Significance}

However, it has two major drawbacks: First,
the significance level for the hard-coded critical values 2 and 4 is
usually unknown.

  - for non-significant results, clearly reduce the amount of color, i.e. the maximum
    chroma used.

In particular,
the Friendly shading does not provide a visualization of the maximum test (\ref{eq:MaxAbs})
as in general it is unknown to which significance level $\alpha$ the
values $c_\alpha = 2$ and $4$ correspond for the table being visualized.

As pointed out in Section \ref{sec:tests}, the critical values for
the maximum statistic $M$ from (\ref{eq:MaxAbs}) can be derived
from the permutation distribution. Instead of the hard-coded
values 2 and 4, the particular critical values, e.g., at the levels
$\alpha = 0.1$ and $0.01$, for the table to be visualized could be
used. Hence, exactly those residuals causing the (potential)
dependency within the table are highlighted. At the moment, these critical values
are derived in \texttt{vcd} by simulation of the underlying distribution.

To illustrate what a `postive' example looks like we give the association plot
for the well-known Hair-Eye-Color data set (avilable in base \textsf{R}) which cross-tabulates the hair and eye
color and gender of 592 statistics students: Figure \ref{fig:haireye-assoc-signif}
shows the association plot for the 328 female students. It is easy to see that
the hypothesis of independence is rejected at 1\% level by the maximum test (\ref{eq:MaxAbs})
due to higher frequencies of students with hair and eye color
black/brown and blond/blue and lower frequencies of
blond/brown and brown/blue than would be plausible under independence. Furthermore,
it can also be seen that the residual corresponding to the black/blue
cell is significant at the 10\% level.

We have seen that this approach performs very well when the test statistic
which should be visualized is the maximum statistic (\ref{eq:MaxAbs}). However,
it is difficult to extend this approach to general tests for independence,
in particular in the HCL space, where the ranges of chroma and luminance are
not independent. In contrast the HSV space has the advantage that all
three dimensions hue, saturation and value have the range $[0,1]$ and can
be varied independently. Therefore, a different approach than the one
described above could be to use the hue for the sign of the residuals as
before, the saturation for the absolute size of the residuals, and the
value as an indicator for the significance of some test statistic (only
using the full value when the overall test for independence rejects
the null hypothesis). The resulting mosaic plots under this paradigm
are depicted in Figure \ref{fig:both-mosaic-signif}. Again, it can
clearly be seen that despite some large residuals there is no
evidence against independence for the Bundesliga data, but that the
null hypothesis has to be rejected for the Hair-Eye-Color data.

\setkeys{Gin}{width=\textwidth}
\begin{figure}[htbp]
\begin{center}
<<shadings,fig=TRUE,echo=FALSE,height=6,width=11,results=hide>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
pushViewport(viewport(layout.pos.col=1, layout.pos.row=1))
mosaic(art, newpage = FALSE, gp = shading_Friendly, margins = c(2.5, 4, 2.5, 3), gp_args = list(eps = NULL, lty = 1))
popViewport()
pushViewport(viewport(layout.pos.col=2, layout.pos.row=1))
mosaic(art, newpage = FALSE, gp = art_gp, margins = c(2.5, 3, 2.5, 4))
popViewport(2)
@
\caption{Mosaic plot for the Arthritis data with Friendly shading (left) and extended shading (right).}
\label{fig:shadings}
\end{center}
\end{figure}

\section{Extensions to Conditional Independence}

UCB admissions data \citep{vcd:Bickel+Hammel+O'Connell:1975}
double max: \Sexpr{stat(ucb_doublemax)}, p-value = \Sexpr{pval(ucb_doublemax)}
max Chisq: \Sexpr{stat(ucb_maxChisq)}, p-value = \Sexpr{pval(ucb_maxChisq)}
sum Chisq: \Sexpr{stat(ucb_sumChisq)}, p-value = \Sexpr{pval(ucb_sumChisq)}
simulated from $5000$ repetitions

\cite{vcd:Friendly:1999} discusses grouping similar to coplots (conditioning
plots) \citep{vcd:Cleveland:1993} that lead to trellis graphics
\citep{vcd:Becker+Cleveland+Shyu:1996}.

Independence problems do not only occur in 2-way tables, although that
is an important special case, but they are also important in tables of
higher dimensionality and can follow much more complex patterns.
These are again defined based on the underlying table of theoretical cell probabilities
$(\pi_{ij\dots})$ with more than two dimensions. Models of interest include the
null hypotheses of conditional independence:
  \[ \pi_{ijk\dots} = \pi_{i|k\dots}\pi_{j|k\dots} \]

Classical non-graphical methods for these problems include the $\chi^2$
test, Fisher's exact test, the Cochran-Mantel-Haenzel test (for $2 \times 2 \times
K$-tables), and the analysis of log-linear models for more
complex settings.

As an example, two natural ways to use the visualization techniques described in the previous
sections would be to use (Trellis-like) conditioning plots or pairs plots
(like mosaic matrices) to visualize these more complex patterns of
independence.

Two ideas for the problem of conditional independence are briefly outlined
here and illustrated using the famous admissions data of the University of
California at Berkley (UCB) which is available in base \textsf{R}.
In this data, the question whether there is gender
discrimination at the UCB leads to the result that although women seem to
be disadvantaged at the aggregated level there is no gender discrimination
conditioned on the department---with the very exception of one department
in which women are {\em more} likely to be admitted than would be plausible
under independence. Exactly this is illustrated in the conditioning
assocation plot in Figure \ref{fig:UCB-assoc-signif}.

Similarly, the same data can be visualized using a mosaic matrix where
a conditional independence model is fitted in each plot (see Figure
\ref{fig:UCB-mosaic-signif}).

\setkeys{Gin}{width=\textwidth}
\begin{figure}[htbp]
\begin{center}
<<conditional_assoc,fig=TRUE,echo=FALSE,height=6,width=9>>=
cotabplot(~ Gender + Admit | Dept, data = ucb, panel = ucb_cotab)
@
\caption{Conditional association plot for UCB admissions.}
\label{fig:emp-mosaic1}
\end{center}
\end{figure}

\section{Conclusions}

\bibliography{vcd}

\end{document}

%% for reproducibility, print also

<<graphics>>
## load packages and data
library(vcd)
data(Arthritis)
art <- xtabs(~ Treatment + Improved, data = Arthritis, subset = Sex == "Female")
data(UCBAdmissions)
ucb <- aperm(UCBAdmissions)

## plain classic plots
mosaic(art)
assoc(art)

## new max shading with HCL colors
mosaic(art, gp = shading_Friendly)
set.seed(1234)
mosaic(art, gp = shading_max)

## conditional independence
set.seed(1234)
cotabplot(~ Admit + Gender | Dept, data = ucb, panel = cotab_coindep, type = "assoc")
## which uses the double max functional
set.seed(1234)
coindep_test(ucb, "Dept")

## alternatively use Peason functional
## via simulation of conditional distribution:
set.seed(1234)
coindep_test(ucb, "Dept", indepfun = function(x) sum(x^2), aggfun = sum)
## via asymptotic distribution:
fm <- loglm(~ Dept * (Gender + Admit), data = ucb)
fm
plot(fm, labeling_args = list(abbreviate = c(Admit = 2)))
@

%% deprecated stuff

%% employment data
<<employment>>
## data
data(Employment)
emp <- aperm(Employment)
emp1 <- co_table(Employment, "LayoffCause")[["Replaced"]]

## plain
assoc(emp1)
mosaic(emp1)

## new shadings
mosaic(emp1, gp = shading_Friendly)
mosaic(emp1, gp = shading_max)

## conditional
mosaic(emp, expected = ~(EmploymentStatus+EmploymentLength)*LayoffCause)
cotabplot(emp, "LayoffCause", split_vertical = TRUE)
cotabplot(emp, panel = cotab_coindep, type = "assoc")
cotabplot(emp, "LayoffCause", panel = cotab_coindep, type = "assoc", split_vertical = TRUE)
@

%% Friendly shading in detail
\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[htbp]
\begin{center}
<<Friendly-hue,fig=TRUE,echo=FALSE,height=4.8,width=8>>
plot(0.5, 0.5, xlim = c(0, 1), ylim = c(0, 1), type = "n", axes = FALSE, xlab = "", ylab = "", yaxs = "i")
dummy <- 0:100/100
dummy.col <- hsv(dummy[-101] + 1/200, 1, 1)
rect(dummy[-101], 0.6, dummy[-1], 1, border = dummy.col, col = dummy.col)
rect(0, 0.6, 1, 1)
axis(3)
mtext("hue", line = 2, cex = 1.5)
text(1, 0.45, "saturation = 1\nvalue = 1", pos = 2, cex = 1.2)
lines(c(0.1, 0), c(0.3, 0.6), lty = 2)
lines(c(2/3, 2/3), c(0.3, 0.6), lty = 2)
rect(0, 0.1, 0.2, 0.3, col = hsv(0, 1, 1))
rect(2/3 - 0.1, 0.1, 2/3 + 0.1, 0.3, col = hsv(2/3, 1, 1))
text(0.1, 0.1, expression(r[ij] < 0), pos = 1, cex = 1.5)
text(2/3, 0.1, expression(r[ij] > 0), pos = 1, cex = 1.5)
@

<<Friendly-sat,fig=TRUE,echo=FALSE,height=4.8,width=8>>
plot(0.5, 0.5, xlim = c(0, 1), ylim = c(0, 1), type = "n", axes = FALSE, xlab = "", ylab = "", yaxs = "i")
dummy <- 0:100/100
dummy.col <- hsv(2/3, dummy[-101] + 1/200, 1)
rect(dummy[-101], 0.8, dummy[-1], 1, border = dummy.col, col = dummy.col)
dummy.col <- hsv(0, dummy[-101] + 1/200, 1)
rect(dummy[-101], 0.6, dummy[-1], 0.8, border = dummy.col, col = dummy.col)
rect(0, 0.8, 1, 1)
rect(0, 0.6, 1, 0.8)
axis(3)
mtext("saturation", line = 2, cex = 1.5)
mtext(c("h = 0", "h = 2/3"), side = 2, at = c(0.7, 0.9), las = 1, cex = 1.2)
text(0.8, 0.5, "value = 1", pos = 2, cex = 1.2)
lines(c(0.1, 0), c(0.3, 0.6), lty = 2)
lines(c(0.5, 0.5), c(0.3, 0.6), lty = 2)
lines(c(0.9, 1), c(0.3, 0.6), lty = 2)
rect(0, 0.2, 0.2, 0.3, col = hsv(2/3, 0, 1))
rect(0, 0.1, 0.2, 0.2, col = hsv(0, 0, 1))
rect(0.4, 0.2, 0.6, 0.3, col = hsv(2/3, 0.5, 1))
rect(0.4, 0.1, 0.6, 0.2, col = hsv(0, 0.5, 1))
rect(0.8, 0.2, 1, 0.3, col = hsv(2/3, 1, 1))
rect(0.8, 0.1, 1, 0.2, col = hsv(0, 1, 1))
text(0.1, 0.1, expression(abs(r[ij]) < 2), pos = 1, cex = 1.5)
text(0.5, 0.1, expression((2 < abs(r[ij])) < 4), pos = 1, cex = 1.5)
text(0.9, 0.1, expression(abs(r[ij]) > 4), pos = 1, cex = 1.5)
@

<<Friendly-val,fig=TRUE,echo=FALSE,height=4.8,width=8>>
plot(0.5, 0.5, xlim = c(0, 1), ylim = c(0, 1), type = "n", axes = FALSE, xlab = "", ylab = "", yaxs = "i")
dummy <- 0:100/100
dummy.col <- hsv(2/3, 1, dummy[-101] + 1/200)
rect(dummy[-101], 0.8, dummy[-1], 1, border = dummy.col, col = dummy.col)
dummy.col <- hsv(0, 1, dummy[-101] + 1/200)
rect(dummy[-101], 0.6, dummy[-1], 0.8, border = dummy.col, col = dummy.col)
rect(0, 0.8, 1, 1)
rect(0, 0.6, 1, 0.8)
axis(3)
mtext("value", line = 2, cex = 1.5)
mtext(c("h = 0", "h = 2/3"), side = 2, at = c(0.7, 0.9), las = 1, cex = 1.2)
text(0, 0.5, paste("saturation =", 1), pos = 4, cex = 1.2)
lines(c(0.85, 1), c(0.3, 0.6), lty = 2)
rect(0.7, 0.2, 0.8, 0.3, col = hsv(2/3, 0, 1))
rect(0.8, 0.2, 0.9, 0.3, col = hsv(2/3, 0.5, 1))
rect(0.9, 0.2, 1, 0.3, col = hsv(2/3, 1, 1))
rect(0.7, 0.1, 0.8, 0.2, col = hsv(0, 0, 1))
rect(0.8, 0.1, 0.9, 0.2, col = hsv(0, 0.5, 1))
rect(0.9, 0.1, 1, 0.2, col = hsv(0, 1, 1))
lines(c(0.5, 0.5), c(0.3, 0.6), lty = 2)
lines(c(0.85, 1), c(0.3, 0.6), lty = 2)
rect(0.35, 0.2, 0.45, 0.3, col = hsv(2/3, 0, 0.5))
rect(0.45, 0.2, 0.55, 0.3, col = hsv(2/3, 0.5, 0.5))
rect(0.55, 0.2, 0.65, 0.3, col = hsv(2/3, 1, 0.5))
rect(0.35, 0.1, 0.45, 0.2, col = hsv(0, 0, 0.5))
rect(0.45, 0.1, 0.55, 0.2, col = hsv(0, 0.5, 0.5))
rect(0.55, 0.1, 0.65, 0.2, col = hsv(0, 1, 0.5))
rect(0.7, 0.2, 0.8, 0.3, col = hsv(2/3, 0, 1))
rect(0.8, 0.2, 0.9, 0.3, col = hsv(2/3, 0.5, 1))
rect(0.9, 0.2, 1, 0.3, col = hsv(2/3, 1, 1))
rect(0.7, 0.1, 0.8, 0.2, col = hsv(0, 0, 1))
rect(0.8, 0.1, 0.9, 0.2, col = hsv(0, 0.5, 1))
rect(0.9, 0.1, 1, 0.2, col = hsv(0, 1, 1))
text(0.5, 0.1, "non-significant", pos = 1, cex = 1.5)
text(0.85, 0.1, "significant", pos = 1, cex = 1.5)
@

\caption{Extended Friendly shading in HSV space.}
\label{fig:Friendly-HSV}
\end{center}
\end{figure}