### 2005-08-07 trunk/vcd/inst/doc/shading.Rnw
New shading vignette added.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@437 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[a4paper]{article}
\usepackage{hyperref, graphicx, color, alltt, a4wide}
\usepackage{Sweave}
\usepackage[round]{natbib}
\definecolor{Red}{rgb}{0.7,0,0}
\definecolor{Blue}{rgb}{0,0,0.8}
\definecolor{hellgrau}{rgb}{0.55,0.55,0.55}
\newcommand{\pkg}[1]{{\normalfont\fontseries{b}\selectfont #1}}
\newcommand{\code}[1]{\texttt{#1}}
\newenvironment{smallexample}{\begin{alltt}}{\end{alltt}}

\begin{document}

%\VignetteIndexEntry{Colors and residual-based Shadings in Strucplots}
%\VignetteDepends{vcd}
%\VignetteKeywords{graphics, categorical data}
%\VignettePackage{vcd}

\SweaveOpts{engine=R,eps=TRUE}
\setkeys{Gin}{width=0.8\textwidth}

<<preliminaries,echo=FALSE,results=hide>>=
set.seed(1071)
library(vcd)
data(Titanic)
@

\title{Colors and residual-based Shadings in Strucplots}
\author{by David Meyer\\
  Wirtschaftsuniversit\"at Wien, Austria\\
\url{David.Meyer@R-project.org}
}
\maketitle
\sloppy

Unlike other R graphics functions, strucplots allow almost full
control over the graphical parameters of all plot elements. In
particular, the user can \emph{independently} modify the appearance of the 
tiles of association plots and mosaic plots. Built on top of this functionality, the
framework supplies a set of shading functions that choose colors
appropriate to the visualization of independence models fit to the
visualized data.
The tiles' graphical parameters are set using the \code{gp} argument
of the strucplot functions. This argument basically expects an object
of class \code{gpar} whose components have the same shape (length and dimensionality) 
than the data table (see Section \ref{sec:gp}). For convenience, however, the user can also
supply a function that computes such an object given a vector of
residuals, or, alternatively, a generating function that takes
parameters and returns such a function (see Section \ref{sec:shading}).

\section{Specifying graphical parameters in strucplots}
\label{sec:gp}

As an example, consider the Admissions to the University of California
in Berkely (UCB) data. In the table collapsed over departments, we
would like to highlight the (incidentally wrong) impression that there
were to much male students accepted compared to the presumably
discriminated female students (see Figure \ref{fig:ucb}):

<<ucb,fig=FALSE,echo=TRUE,height=6,width=7,results=verbatim>>=
ucb <- margin.table(UCBAdmissions, 1:2)
ucb

fill_colors <- matrix(c("dark cyan","gray","gray","dark magenta"), ncol = 2)
fill_colors

mosaic(ucb, gp = gpar(fill = fill_colors, col = 0))
@

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<ucbfig,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(ucb, gp = gpar(fill = fill_colors, col = 0))
@
\caption{Mosaic plot for the UCBAdmissions data with highlighted cells.}
\label{fig:ucb}
\end{center}
\end{figure}

\noindent As the example shows, we therefore create a fourfold table with appropriate colors and supply them to the \code{fill} component of the \code{gpar} object passed to the \code{gp} argument of \code{mosaic()}. In addition, we use the \code{col} component to turn off the borders for cosmetic reasons.

If the parameters specified in the \code{gpar} object are incomplete,
strucplot functions will recycle them along the last splitting
dimension. In the following example based on the Survival on the
Titanic data, we will highlight all cells corresponding to survived
passengers (see Figure \ref{fig:recycling}):

\begin{Sinput}
> mosaic(Titanic, gp = gpar(fill = c("gray","dark magenta")))
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<recycling,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(Titanic, gp = gpar(fill = c("gray","dark magenta")))
@
\caption{Recycling of parameters.}
\label{fig:recycling}
\end{center}
\end{figure}

\section{Using residual-based shadings}
\label{sec:shading}

This new flexible way of specifying graphical parameters is the basis for a suite of
shading functions that modify the tiles' appearance with respect to a vector of 
residuals, resulting from deviations of the observed from the expected
values under a given independence model. 
The idea is to visualize at least the size of the residuals, but some
shadings, additionally, indicate overall significance. One particular
shading, the maximum shading, even allows to identify those cells that
cause the rejection of the independence hypothesis.

All shading functions are parameterized be the means of generating
functions that return the actual shading functions. As an example,
consider the \code{shading\_binary()} function that only visualizes the
sign of the residuals:

\begin{Sinput}
shading_binary <- function(observed = NULL, residuals = NULL, expected = NULL, 
                           df = NULL, col = hcl(c(260, 0), 50, 70))
{
\dots
}
\end{Sinput}

\noindent Each generating function, passed to a strucplot functions' \code{gp}
argument, takes a few arguments that describe the model
(\code{observed}, \code{expected}, \code{residuals}, and \code{df}), and a
number of parameters that specify the appearance (in the case of
\code{shading\_binary()} only \code{col}, a vector of length two, 
that specifies the two colors corresponding to negative and 
positive residuals, respectively). The
generating function then return the actual shading function taking 
only one argument: a vector of residuals, and returning an object
of class \code{gpar} that contains the graphical parameters for all
tiles corresponding to the residuals. In the following example, positive 
and negative residuals of an hypothesized independence model shall be 
visualized by different colors (see Figure \ref{fig:binary}):

\begin{Sinput}
> obs <- margin.table(UCBAdmissions, 1:2)
> exp <- independence_table(obs)
> res <- (obs - exp) / sqrt(exp)
> res
          Gender
Admit           Male    Female
  Admitted  4.784093 -5.793466
  Rejected -3.807325  4.610614

> gpobj <- shading_binary()(res)
> gpobj
$fill
     [,1]      [,2]     
[1,] "#96A2DF" "#E18E9E"
[2,] "#E18E9E" "#96A2DF"

> mosaic(obs, gp = gpobj)
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<binary,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
obs <- margin.table(UCBAdmissions, 1:2)
exp <- independence_table(obs)
res <- (obs - exp) / sqrt(exp)
gpobj <- shading_binary()(res)
mosaic(obs, gp = gpobj)
@
\caption{Binary shading.}
\label{fig:binary}
\end{center}
\end{figure}

\noindent The shading-generating functions available
in \pkg{vcd} include model-fitting, and hence the {\sf R} code
corresponding to the example above in fact reduces to:

\begin{Sinput}
> mosaic(obs, gp = shading_binary)
\end{Sinput}

\section{An overview of the shading functions in \pkg{vcd}}

Currently, there are three basic shadings available
(\code{shading\_hcl(), shading\_hsv(), and shading\_binary()}, as well
as two derived functions (\code{shading\_Friendly()} building upon
\code{shading\_hsv()} and \code{shading\_max()} building upon \code{shading\_hcl}).
\code{shading\_binary()} has been explained in the previous
paragraph. \code{shading\_hsv()} and \code{shading\_hcl()} are quite
similar in their functionality, but use different color spaces: the
Hue-Saturation-Value (HSV) and the Hue-Chroma-Luminance (HCL) scheme,
respectively. We will focus on the latter, since the HCL space is
preferable to the HSV space in many respects (e.g., uniform saturations over
different hues, device independentness, \dots).

In the HCL space, colors are specified in three dimensions: hue, chroma (`colorfulness', saturation), and luminance (`brightness', amount of gray). These three dimensions are used by \code{shading\_hcl()} to visualize information about the residuals and the underlying independence model. A residuals' hue indicates its sign (by default blue and red for positive and negative ones, respectively). An overall full or poor chroma indicates significant or non-significant results, respectively. Finally, the luminance of a residual is set according to its size.

As an example, we will visualize the (in)dependence of hair and eye
color in the HairEyeColor data set (see Figure \ref{fig:haireye})

\begin{Sinput}
> haireye <- margin.table(HairEyeColor, 1:2)
> mosaic(haireye, shade = TRUE)
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<haireye,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
haireye <- margin.table(HairEyeColor, 1:2)
mosaic(haireye, shade = TRUE)
@
\caption{Shaded residuals in the HairEyeColor data set---two cut-off points}
\label{fig:haireye}
\end{center}
\end{figure}

\noindent The default shading scheme is \code{shading\_hcl()}, so it
suffices to turn shading on to bring colors to the mosaic plot. Large
positive residuals (greater than $4$) can be found for brown eyes/black hair and blue
eyes/blond hair, they are colored in saturated red. 
On the other side, there is a large negative residual
(less than $-4$) for brown eyes/blond hair, colored deep blue. There are also three
medium-sized positive (negative) residuals between 2 and 4 ($-2$ and
$-4$): the colors for them are less saturated. 
The heuristic for choosing the cut-off points $2$ and $4$ is that the Pearson residuals
are approximately standard normal which implies that the highlighted cells are those with
residuals \emph{individually} significant at approximately the $\alpha
= 0.05$ and $\alpha = 0.0001$ level. These default cut-off points can
be changed to alternative values using the \code{interpolate}
argument (see Figure \ref{fig:interpolate}):

\begin{Sinput}
> mosaic(haireye, shade = TRUE, 
         gp_args = list(interpolate = 1:4, h = c(100, 200)))
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<interpolate,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(haireye, shade = TRUE, 
       gp_args = list(interpolate = 1:4, h = c(100, 200)))
@
\caption{Shaded residuals in the HairEyeColor data set---four cut-off points}
\label{fig:interpolate}
\end{center}
\end{figure}

\noindent Note that we also changed the hue values for purely cosmetic
reasons. The third remaining dimension, the luminance, is used for
visualizing the significance of a test statistic. The $p$ value can
be supplied to the \code{p.value} argument. If none is specified, it
is computed from a Chi-squared distribution with \code{df} degrees of
freedom. The \code{level} argument is used to specify the confidence
level:  if \code{p.value} is smaller than \code{1 - level}, bright colors are used, 
otherwise dark colors are employed. These colors are computed using
the two luminance values supplied to the \code{l} argument. The
following example shows the relationship of home goals and away
goals of the German Bundesliga in 1995: although there are two
``larger'' residuals (one greater than 2, one less then $-2$), the
Chi-squared test does not reject the null hypothesis of
independence. Consequently, the colors appear dark (see Figure \ref{fig:bundesliga}):

\begin{Sinput}
> bl <- xtabs(~ HomeGoals + AwayGoals, data = Bundesliga, subset = Year == 1995)
> mosaic(bl, shade = TRUE)
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<bundesliga,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
bl <- xtabs(~ HomeGoals + AwayGoals, data = Bundesliga, subset = Year == 1995)
mosaic(bl, shade = TRUE)
@
\caption{Non-significant Chi-squared test using part of the Bundesliga data.}
\label{fig:bundesliga}
\end{center}
\end{figure}

\noindent The \code{interpolate} argument also accepts a function that
automatically chooses the cut-off points in a data-driven way. 
This functionality is used by \code{shading\_max()} that uses the
maximum statistic to visualize significant \emph{cells} causing the
rejection of the independence hypothesis. The
\code{level} argument of \code{shading\_max()} 
then can be used to specify several confidence
levels from which the corresponding cut-off points are computed. 
By default, two cut-off points are computed corresponding 
to confidence levels of $90\%$ and $99\%$, respectively.
In the following example, we investigate the effect of a new treatment
for rheumatoid arthritis on a group of female patients using the
maximum shading (see Figure \ref{fig:maximum}):

\begin{Sinput}
> art <- xtabs(~ Improved + Treatment, data = Arthritis, subset = Sex == "Female")
> mosaic(art, shade = TRUE, gp = shading_max)
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<arthritis,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
art <- xtabs(~ Improved + Treatment, data = Arthritis, subset = Sex == "Female")
mosaic(art, gp = shading_max)
@
\caption{Significant maximum test on female patients of the Arthritis
  data.}
\label{fig:maximum}
\end{center}
\end{figure}

\noindent The maximum test is significant although the residuals are
all in the $[-2,2]$ interval, as the cut-off points are even smaller.
The \code{shading\_hcl} function with default cut-off points would
not have shown any color. In addition, since the test statistic is the
maximum of the absolute pearson residuals, \emph{every} colored
residual violates the hypotheses of independence, and thus, the
`culprits' can immediately be identified.

The last shading function is \code{shading\_Friendly()}. This shading
mimics the shading introduced by \cite{vcd:Friendly:2000}. This
shading corresponds to the HSV shading with default cut-off points (2 and 4), and in
addition uses the border color and line type to visualize the
residuals' sign. The following example again visualizes the Bundesliga
data from above, this time using the Friendly scheme and, in addition, an
alternative legend:

\begin{Sinput}
> mosaic(bl, gp = shading_Friendly, legend = legend_fixed, zero_size = 0)
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<friendly,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(bl, gp = shading_Friendly, legend = legend_fixed, zero_size = 0)
@
\caption{The Bundesliga data for 1995 using the Friendly shading.}
\label{fig:friendly}
\end{center}
\end{figure}

\noindent (The bullets indicating zero observed values are removed here
since this feature does not exist in the original SAS implementation
of the Friendly mosaic plots).

\section{Creating new shading functions}

The example with four cut-off points (Figure \ref{fig:interpolate})
might suggest that a continuous shading simply visualizing the complete
range of confidence levels might be useful. The following code shows
how such a shading function might be constructed:

\begin{Sinput}
shading_continuous <- function(observed = NULL, residuals = NULL,
                               expected = NULL, df = NULL, n = 5000, ...)
{
  ## function works only for two-way tables
  stopifnot(length(dim(observed)) == 2)

  ## sample permutation distribution of the maximum statistic
  obs.test <- coindep_test(observed, n = n)

  ## Use distribution to map residuals to [0,1] range
  ipol <- function(x) abs(obs.test$pdist(x))

  ## call HCL shading 
  shading_hcl(observed = NULL, residuals = NULL, expected = NULL,
              df = NULL, interpolate = ipol, p.value = obs.test$p.value, ...)
}
class(shading_continuous) <- "panel_generator"

mosaic(haireye, gp = shading_continuous)
\end{Sinput}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<continuous,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
shading_continuous <- function(observed = NULL, residuals = NULL,
                               expected = NULL, df = NULL, n = 5000, ...)
{
  ## function works only for two-way tables
  stopifnot(length(dim(observed)) == 2)

  ## sample permutation distribution of the maximum statistic
  obs.test <- coindep_test(observed, n = n)

  ## Use distribution to map residuals to [0,1] range
  ipol <- function(x) abs(obs.test$pdist(x))

  ## call HCL shading 
  shading_hcl(observed = NULL, residuals = NULL, expected = NULL,
              df = NULL, interpolate = ipol, p.value = obs.test$p.value, ...)
}
class(shading_continuous) <- "panel_generator"

mosaic(haireye, gp = shading_continuous)
@
\caption{The HairEyeColor data with continuous shading.}
\label{fig:continuous}
\end{center}
\end{figure}

\noindent We first compute the permutation distribution for the
maximum statistic using \code{coindep\_test}, and then simply use it
as the interpolation function passed to \code{shading\_hcl}. The result
using the HairEyeColor data, however, is disappointing: too much color
makes it difficult to interpret the image, and the subtle color
differences are hard to catch. Therefore, we only included shadings
with discrete cut-off points.

\begin{thebibliography}{}

\bibitem[Friendly, 2000]{vcd:Friendly:2000}
Friendly, M. (2000).
\newblock {\em Visualizing Categorical Data}.
\newblock SAS Insitute, Carey, NC.

\end{thebibliography}

\end{document}