### 2005-10-02 trunk/vcd/inst/doc/strucplot.Rnw

\documentclass[a4paper]{article}
\usepackage{hyperref, graphicx, color, alltt, a4wide}
\usepackage{Sweave}
\usepackage[round]{natbib}
\definecolor{Red}{rgb}{0.7,0,0}
\definecolor{Blue}{rgb}{0,0,0.8}
\definecolor{hellgrau}{rgb}{0.55,0.55,0.55}
\newcommand{\pkg}[1]{{\normalfont\fontseries{b}\selectfont #1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\class}[1]{\textsf{#1}}
\newenvironment{smallexample}{\begin{alltt}}{\end{alltt}}
\newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
\newcommand{\data}[1]{`\texttt{#1}'}

%% \code without `-' ligatures
\def\nohyphenation{\hyphenchar\font=-1 \aftergroup\restorehyphenation}
\def\restorehyphenation{\hyphenchar\font=`-}
{\catcode`\-=\active%
  \global\def\code{\bgroup%
    \catcode`\-=\active \let-\codedash%
    \Rd@code}}
\def\codedash{-\discretionary{}{}{}}
\def\Rd@code#1{\texttt{\nohyphenation#1}\egroup}

\newcommand{\codefun}[1]{\code{#1()}}

\begin{document}

%\VignetteIndexEntry{The Strucplot Framework---Visualizing Multi-way Contingency Tables}
%\VignetteDepends{vcd}
%\VignetteKeywords{graphics, categorical data}
%\VignettePackage{vcd}

\SweaveOpts{engine=R,eps=TRUE,height=6,width=7,results=hide,fig=FALSE,echo=TRUE}
\setkeys{Gin}{width=0.7\textwidth}

<<preliminaries,echo=FALSE,results=hide>>=
set.seed(1071)
library(vcd)
data(Titanic)
data(HairEyeColor)
@

\title{The Strucplot Framework---Visualizing Multi-way Contingency Tables}
\author{by David Meyer, Achim Zeileis, and Kurt Hornik}
\date{}

\maketitle
\sloppy

\section{Framework Overview}

The strucplot framework in the {\sf R} package \pkg{vcd} visualizes
multi-way contingency tables and integrates techniques such as
mosaic displays, association plots, and sieve plots. The main idea is to visualize
the tables' cells arranged in rectangular form. For multi-way tables,
the variables are nested into rows and columns using recursive
conditional splits, given the margins. The result is a 
`flat' representation that can be visualized in
ways similar to a two-dimensional table.
This principle defines a class of conditional displays which allows
for granular control of graphical appearance aspects, including:

\begin{itemize}
\item the content of the tiles
\item the split direction for each dimension
\item the graphical parameters of the tiles' content
\item the spacing between the tiles
\item the labeling of the tiles
\end{itemize}

\noindent This document gives an introduction to the framework,
whereas labeling and shading issues are described in separate vignettes.

The strucplot framework is highly modularized: Figure \ref{fig:struc}
shows the hierarchical relationship between the various components. 
On the lowest level, there are several groups of workhorse and 
parameter functions that directly or indirectly influence the final
appearance of the plot. These `graphical appearance control'
(`grapcon') functions are created by generating functions, allowing
flexible parameterization and extensibility. 
The first part of their names (\code{\emph{group\_foo()}}) 
reflects the group they belong to (strucplot core, labeling,
legend, shading, or spacing). The workhorse functions (created by 
\code{\emph{struc\_foo()}}, 
\code{\emph{labeling\_foo()}}, and \code{\emph{legend\_foo()}}) 
directly produce graphical output (``add ink to the canvas''), whereas
the parameter functions
(created by \code{\emph{spacing\_foo()}} and \code{\emph{shading\_foo()}}) compute
graphical parameters used by the others. The grapcon functions returned by 
\code{\emph{struc\_foo()}} implement the core functionality, creating the tiles and their
content.  On the second level of the framework, a suitable combination
of grapcon functions (or, alternatively, corresponding generating functions)
is passed as ``hyperparameters'' to \codefun{strucplot}.
This central function 
sets up the graphical layout using grid viewports (see Figure \ref{fig:layout}),
and coordinates the specified core, labeling, shading, and spacing functions to produce 
the plot. On the third level, we provide
several convenience functions such as \codefun{mosaic},
\codefun{sieve}, \codefun{assoc}, and \codefun{doubledecker} which 
interface \codefun{strucplot} through sensible parameter defaults
and support for model formulas. Finally, on the fourth
level, there are `related' \pkg{vcd} functions (such as \codefun{cotabplot}
and the \codefun{pairs} methods for table objects) 
arranging collections of plots of the strucplot 
framework into more complex displays (e.g., by means of panel functions).

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.8\textwidth]{struc}
    \caption{Components of the strucplot framework.}
    \label{fig:struc}
  \end{center}
\end{figure}

\setkeys{Gin}{width=0.6\textwidth}
\begin{figure}[h]
\begin{center}
<<vcdlayout,fig=TRUE,echo=FALSE,width=6,height=6>>=
    pushViewport(vcd:::vcdViewport(legend = T, mar =4))
    seekViewport("main")
    grid.rect(gp = gpar(lwd = 3))
    grid.text("main", gp = gpar(fontsize = 20))
    seekViewport("sub")
    grid.rect(gp = gpar(lwd = 3))
    grid.text("sub", gp = gpar(fontsize = 20))
    seekViewport("plot")
    grid.rect(gp = gpar(lwd = 3))
    grid.text("plot", gp = gpar(fontsize = 20))
    seekViewport("legend")
    grid.text("legend", rot = 90, gp = gpar(fontsize = 20))
    grid.rect(gp = gpar(lwd = 3))
    seekViewport("legend_sub")
    grid.rect(gp = gpar(lwd = 3))
    grid.text("[F]", gp = gpar(fontsize = 20))
    seekViewport("legend_top")
    grid.rect(gp = gpar(lwd = 3))
    grid.text("[E]", gp = gpar(fontsize = 20))
    seekViewport("margin_top")
    grid.rect(gp = gpar(lwd = 3))
    grid.text("margin_top", gp = gpar(fontsize = 20))
    seekViewport("margin_bottom")
    grid.rect(gp = gpar(lwd = 3))
    grid.text("margin_bottom", gp = gpar(fontsize = 20))
    seekViewport("margin_right")
    grid.rect(gp = gpar(lwd = 3))
    grid.text("margin_right", rot = 90, gp = gpar(fontsize = 20))
    seekViewport("margin_left")
    grid.rect(gp = gpar(lwd = 3))
    grid.text("margin_left", rot = 90, gp = gpar(fontsize = 20))
    seekViewport("corner_top_left")
    grid.rect(gp = gpar(lwd = 3))
    grid.text("[A]", gp = gpar(fontsize = 20))
    seekViewport("corner_top_right")
    grid.rect(gp = gpar(lwd = 3))
    grid.text("[B]", gp = gpar(fontsize = 20))
    seekViewport("corner_bottom_left")
    grid.rect(gp = gpar(lwd = 3))
    grid.text("[C]", gp = gpar(fontsize = 20))
    seekViewport("corner_bottom_right")
    grid.rect(gp = gpar(lwd = 3))
    grid.text("[D]", gp = gpar(fontsize = 20))
@ 
\caption{Viewport layout for strucplot displays with their names. [A] =
  ``corner\_top\_left'', [B] = ``corner\_top\_right'', 
  [C] = ``corner\_bottom\_left'', [D] = ``corner\_bottom\_right'', [E]
  = ``legend\_top'', [F] = ``legend\_sub''.}
\label{fig:layout}
\end{center}
\end{figure}
\setkeys{Gin}{width=0.7\textwidth}

\section{Mosaic, Association, and Sieve Plots}

As an example, consider the \data{HairEyeColor} 
data containing two polytomous variables (hair and eye color), 
as well as one (artificial) dichotomous variable (sex, i.e., gender). The
`flattened' contingency table can be obtained using the
\codefun{structable} function (quite similar to \codefun{ftable} in
base {\sf R}, but allowing the specification of split directions):

<<structable,results=verbatim>>=
(hec <- structable(Eye ~ Sex + Hair, data = HairEyeColor))
@

Let us first visualize the contingency table by means of a mosaic plot
\citep{vcd:Hartigan+Kleiner:1984} which is basically 
an area-proportional visualization of (typically, observed) frequencies, composed
of tiles (corresponding to the cells) created by recursive
vertical and horizontal splits of a square. Thus, the area of each tile
is proportional to the corresponding cell entry \emph{given} the
dimensions of previous splits. Figure \ref{fig:observed} depicts 
the effect of

<<Observed>>=
mosaic(hec)
@

\noindent equivalent to

<<Observed2>>=
mosaic(~ Sex + Eye + Hair, data = HairEyeColor)
@

\begin{figure}[p]
\begin{center}
<<Observedfig,fig=TRUE,echo=FALSE>>=
<<Observed>>
@
\caption{Mosaic plot for the \data{HairEyeColor} data.}
\label{fig:observed}
\end{center}
\end{figure}

\noindent The small bullets indicate zero entries in the corresponding
cell. Note that in contrast to, e.g.,  \codefun{mosaicplot} in base
{\sf R}, the default split direction and level ordering in all strucplot
displays correspond to the textual representation.
It is also possible to visualize the expected values instead of the
observed values (see Figure \ref{fig:expected}):

<<Expected>>=
mosaic(hec, type = "expected")
@

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[p]
\begin{center}
<<Expectedfig,fig=TRUE,echo=FALSE>>=
<<Expected>>
@
\caption{Mosaic plot for the \data{HairEyeColor} data (expected values).}
\label{fig:expected}
\end{center}
\end{figure}

\noindent In order to compare observed and expected values,
a sieve plot \citep{vcd:riedwyl+schuepbach:1994} 
could be used (see Figure \ref{fig:sieve}):

<<sieve>>=
sieve(hec)
@

\begin{figure}[h]
\begin{center}
<<sievefig,fig=TRUE,echo=FALSE>>=
<<sieve>>
@
\caption{Sieve plot for the \data{HairEyeColor} data visualizing simultaneously
  observed and expected values.}
\label{fig:sieve}
\end{center}
\end{figure}

\noindent Alternatively, we can directly inspect the residuals. 
The Pearson residuals (standardized deviations of observed from expected values) 
are preferably visualized using association
plots \citep{vcd:Cohen:1980}. In contrast to \codefun{assocplot} in
base {\sf R}, \pkg{vcd}'s \codefun{assoc}
function scales to more than two variables (see Figure \ref{fig:residuals}):

<<Residuals>>=
assoc(hec, compress = FALSE)
@

\begin{figure}[p]
\begin{center}
<<Residualsfig,fig=TRUE,echo=FALSE,width=6, height=10>>=
<<Residuals>>
@
\caption{Association plot for the \data{HairEyeColor} data.}
\label{fig:residuals}
\end{center}
\end{figure}

\noindent The \code{compress} argument keeps distances between tiles
equal for better comparison.

For both mosaic plots and association plots, 
the splitting of the tiles can be controlled using the
\code{split\_vertical} argument (default: alternating splits starting
with a vertical one).

<<split1>>=
mosaic(hec, split_vertical = c(TRUE, FALSE, TRUE))
@

\noindent For compatibility with \codefun{mosaicplot} in base {\sf R},
the \codefun{mosaic} function also allows the use of a \code{"direction"} 
argument taking a vector of \code{"h"} and
\code{"v"} characters (see Figure \ref{fig:split}):

<<split2>>=
mosaic(hec, direction = c("v","h","v"))
@

\begin{figure}[p]
\begin{center}
<<splitfig,fig=TRUE,echo=FALSE,width=7,height=7>>=
<<split1>>
@
\caption{Mosaic plot for the \data{HairEyeColor} data---alternative splitting.}
\label{fig:split}
\end{center}
\end{figure}

By a suitable combination of splitting, spacing, and
labeling settings, the functions provided by the strucplot framework
can be customized in a quite flexible way. For example, 
\codefun{doubledecker} is simply a wrapper for 
\codefun{mosaic}, setting the right defaults. 
Figure \ref{fig:doubledecker} shows a doubledecker plot of
the \data{Titanic} data, explaining the probability of survival (`survived') by age, given
sex, given class. It is created by:

<<doubledecker1>>=
doubledecker(Titanic)
@

\noindent equivalent to:

<<doubledecker2>>=
doubledecker(Survived ~ Class + Sex + Age, data = Titanic)
@

\begin{figure}[h]
\begin{center}
<<doubledeckerfig,fig=TRUE,echo=FALSE>>=
<<doubledecker1>>
@
\caption{Doubledecker plot for the \data{Titanic} data.}
\label{fig:doubledecker}
\end{center}
\end{figure}

\section{Conditional and partial views}

So far, we have visualized full tables. For objects of class \class{table},
conditioning on levels (i.e., choosing a table subset for fixed levels
of the conditioning variable(s)) is simply done by indexing.
However, subsetting \code{"structable"} objects is more restrictive 
because of their inherent conditional structure. Since the variables
on both the row and the columns side are nested, 
conditioning is only possible ``outside-in'':

<<subsetting,results=verbatim>>=
hec
hec["Male",]
hec[c("Male", "Brown"),]
hec["Male","Green"]
@

\noindent Now, there are several ways for visualizing 
conditional independence structures. The ``brute force'' method is to
draw separate plots for the strata. The following example compares 
the association between hair and eye color, given gender, by using
subsetting on the flat table and \pkg{grid}'s viewport 
framework to visualize the two groups besides each other:

<<Variables>>=
pushViewport(viewport(layout = grid.layout(ncol = 2)))

pushViewport(viewport(layout.pos.col = 1))
mosaic(hec["Male"], margins = c(top = 2, 0), sub = "Male", newpage = FALSE)
popViewport()

pushViewport(viewport(layout.pos.col = 2))
mosaic(hec["Female"], margins = c(top = 2, 0), sub = "Female", newpage = FALSE)
popViewport(2)
@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[p]
\begin{center}
<<Variablesfig,fig=TRUE,echo=FALSE,height=5,width=12>>=
<<Variables>>
@
\caption{Distribution of hair and eye color, given gender.}
\label{fig:parttable}
\end{center}
\end{figure}
\setkeys{Gin}{width=0.7\textwidth}

\noindent Note the use of the \code{margins} argument: it takes a
vector with up to four values whose unnamed components are recycled,
but ``overruled'' by the named arguments. Thus, in the example, only
the top margin is set to 2 lines, and all other to 0.
This idea applies to almost all
vectorized arguments in the strucplot framework (with
\code{split\_vertical} as a prominent exception).

Since mosaic displays are ``conditional plots'' by definition, we can
also use one single mosaic for stratified plots.
The formula interface of \codefun{mosaic} allows the
specification of conditioning variables (see Figure \ref{fig:conditioning}):

<<Conditioning>>=
mosaic(~ Hair + Eye | Sex, data = hec, split_vertical = TRUE, keep_aspect_ratio = FALSE)
@

\begin{figure}[p]
\begin{center}
<<Conditioningfig,fig=TRUE,echo=FALSE,width=12>>=
<<Conditioning>>
@
\caption{Mosaic plot for conditional independence structures.}
\label{fig:conditioning}
\end{center}
\end{figure}

\noindent The effect of using this kind of formula is that
conditioning variables are permuted ahead of the the conditioned variables
in the table, and that \codefun{spacing\_conditional} is used as
default to better
distinguish conditioning from conditioned dimensions.
This spacing uses equal space between tiles of conditioned variables,
and increasing space between tiles of conditioning variables. In
addition, we release the fixed aspect ratio to get less distorted margins.

The \codefun{cotabplot} function does a much better job on this task:
it arranges stratified strucplot displays 
in a lattice-like layout, conditioning on variable \emph{levels}.
The plot in Figure \ref{fig:cotabplot} shows hair and eye color, given sex:

<<cotabplot>>=
cotabplot(~ Hair + Eye | Sex, data = hec, panel_args = list(margins = 3), labeling = labeling_left(clip = FALSE))
#cotabplot(~ Hair + Eye | Sex, data = hec, labeling_args = list(abbreviate = c(Eye = 2)))
@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[h]
\begin{center}
<<cotabplotfig,fig=TRUE,echo=FALSE,height=6,width=11>>=
<<cotabplot>>
@
\caption{Conditional table plot for the \data{HairEyeColor} data.}
\label{fig:cotabplot}
\end{center}
\end{figure}
\setkeys{Gin}{width=0.7\textwidth}

\noindent The \code{labeling\_args} argument ensures that `eye' levels
are abbreviated (see the vignette on labeling for detailed information).

Another high-level function for visualizing conditional independence
models are the \codefun{pairs} methods for table and structable objects.
In contrast to \codefun{cotabplot} which conditions on variables,
the \codefun{pairs} methods create pairwise views of the table. 
The function produces, by default, a plot matrix 
having strucplot displays in the off-diagonal panels, and
the variable names (optionally, with univariate statistics) in the diagonal cells. 
Figure \ref{fig:pairs} shows a pairs display with mosaic plots visualizing mutual 
independence in the lower triangle,
association plots for the same in the upper triangle, and bar charts in the diagonal.

<<pairs>>=
pairs(hec, lower_panel = pairs_assoc, space = 0.3, 
      diag_panel_args = list(rot = -45, just_leveltext = c("left","bottom")))
@

%\setkeys{Gin}{width=\textwidth}
\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[h]
\begin{center}
<<pairsfig,fig=TRUE,echo=FALSE,width=8,height=8>>=
<<pairs>>
@
\caption{Pairs plot for the \data{HairEyeColor} data.}
\label{fig:pairs}
\end{center}
\end{figure}
\setkeys{Gin}{width=0.7\textwidth}

\noindent (The labels of the variables are to be read from left to
right and from top to bottom.) In plots produced by \codefun{pairs}, 
each panel's row and column define two variables $X$ and $Y$ used for the
specification of four different types of independence: pairwise,
total, conditional, and joint. The pairwise mosaic
matrix shows bivariate marginal relations between $X$ and $Y$, collapsed over all
other variables. The total independence mosaic matrix shows mosaic
plots for mutual independence, i.e., for marginal and conditional independence
among all pairs of variables. The conditional
independence mosaic matrix shows mosaic plots for marginal
independence of $X$ and $Y$, given all other variables. The joint independence
mosaic matrix shows mosaic plots for joint independence of all
pairs ($X$, $Y$) of variables from the others.

Since the matrix is symmetric, the upper and lower parts can
independently be used to display different 
types of independence models, or different strucplots displays
(mosaic, association, or sieve plots).
The available panel functions (\codefun{pairs\_assoc}, 
\codefun{pairs\_mosaic}, and \codefun{pairs\_sieve}) 
are simple wrappers to \codefun{assoc}, 
\codefun{mosaic}, and \codefun{sieve}, respectively. 
Obviously, seeing patterns in strucplot matrices 
becomes increasingly difficult with
higher dimensionality. Therefore, this plot is typically used with a suitable 
residual-based shading (described in the vignette on ``Colors and
Residual-based Shadings in the Strucplot Framework'').

\section{Interactive plot modifications}

All strucplot core functions are supposed to produce conditional hierarchical
plots by the means of nested viewports, corresponding to the provided splitting
information. Thus, at the end of the plotting, each tile is associated with
a particular viewport. Each of those viewports has to be
conventionally named, enabling other strucplot modules, in particular the
labeling functions, to access specific tiles after they have been
plotted. The naming convention for the viewports is:

\begin{center}
\code{cell:\emph{Variable1}=\emph{Level1},\emph{Variable2}=\emph{Level2}} \dots
\end{center}

\noindent Clearly, these names depend on the splitting.
The following example shows how to access parts of the plot after it has
been drawn (see Figure \ref{fig:afterplot}):

<<viewportnames>>=
mosaic(~ Hair + Eye, data = hec, pop = FALSE)

seekViewport("cell:Hair=Blond")
grid.rect(gp = gpar(col = "red", lwd = 4))

seekViewport("cell:Hair=Blond,Eye=Blue")
grid.circle(r = 0.2, gp = gpar(fill = "cyan"))
@

\noindent Note that the viewport tree is removed by
default. Therefore, the \texttt{pop} argument has to be set to
\texttt{FALSE} when viewports shall be accessed.

\setkeys{Gin}{width=0.6\textwidth}
\begin{figure}[h]
\begin{center}
<<viewportnamesfig,fig=TRUE,echo=FALSE>>=
<<viewportnames>>
@
\caption{Adding elements to a mosaic plot after drawing.}
\label{fig:afterplot}
\end{center}
\end{figure}

In addition to the viewports, the main graphical elements get names
following a similar construction method. This allows to change
graphical parameters of plot elements \emph{after} 
the plotting (see Figure \ref{fig:changeplot}):

<<changeplot,results=verbatim>>=
assoc(Eye ~ Hair, data = hec, pop = FALSE)
getNames()[1:6]
grid.edit("rect:Hair=Blond,Eye=Blue", gp = gpar(fill = "red"))
@

%% code-chunk reuse does not work with parameter changing
\begin{figure}[h]
\begin{center}
<<changeplotfig,fig=TRUE,echo=FALSE>>=
x <- tab <- margin.table(HairEyeColor, 1:2)
x[] <- "light gray"
x["Blond","Blue"] <- "Red"
assoc(tab, gp = gpar(fill = x))
@
\caption{Changing graphical parameters of elements after drawing.}
\label{fig:changeplot}
\end{center}
\end{figure}

\bibliographystyle{jss}
\bibliography{vcd}

\end{document}