### 2006-04-10 trunk/vcd/paper/hsc.Rnw
went through HSC revision
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@518 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass{svmult}

% choose options for [] as required from the list
% in the Reference Guide, Sect. 2.2

\usepackage{Sweave,hyperref,color,graphicx} 
\usepackage[bottom]{footmisc}% places footnotes at page bottom

\definecolor{Red}{rgb}{0.7,0,0}
\definecolor{Blue}{rgb}{0,0,0.8}
\definecolor{hellgrau}{rgb}{0.55,0.55,0.55}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\SweaveOpts{engine=R,eps=TRUE,height=6,width=7,results=hide,fig=FALSE,echo=TRUE}
\setkeys{Gin}{width=0.7\textwidth}

%% Hyperref
\hypersetup{%
  pdftitle = {Visualizing Contingency Tables},
  pdfsubject = {Handbook of Statistical Computing},
  pdfkeywords = {association plots, contingency tables, HCL colors, HSV colors, mosaic plots, sieve plots},
  pdfauthor = {David Meyer, Achim Zeileis, Kurt Hornik},
  %% change colorlinks to false for pretty printing
  colorlinks = {false},
  linkcolor = {Blue},
  citecolor = {Blue},
  urlcolor = {Red},
  hyperindex = {true},
  linktocpage = {true},
}

<<preliminaries,echo=FALSE,results=hide>>=
myseed <- 1071
set.seed(myseed)
library(vcd)
library(scatterplot3d)
library(xtable)
data(Arthritis)
data(UCBAdmissions)
hospital <- as.table(matrix(c(43,16,3,6,11,10,9,18,16), c(3,3), byrow = TRUE, 
dimnames = list("Visit frequency" = c("Regular","Less than monthly","Never"), 
"Length of stay" = c("2-9","10-19","20+"))))
art <- xtabs(~Treatment + Improved, data = Arthritis, subset = Sex == "Female")

set.seed(myseed)
art_max <- coindep_test(art, n = 5000)

set.seed(myseed)
ucb_max <- coindep_test(aperm(UCBAdmissions, c(3,2,1)), margin = "Dept", n = 5000)

set.seed(myseed)
art_chisq <- coindep_test(art, n = 5000, indepfun = function(x) sum(x^2))

set.seed(myseed)
hos_chisq <- coindep_test(hospital, n = 5000, indepfun = function(x) sum(x^2))

@

%\newcommand{\data}[1]{`\texttt{#1}'}
\newcommand{\data}[1]{#1}
\newcommand{\proglang}[1]{\textsf{#1}}
\newcommand{\pkg}[1]{\texttt{#1}}

\begin{document}

\title*{Visualizing Contingency Tables}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{David Meyer\inst{1}\and
Achim Zeileis\inst{2}\and
Kurt Hornik\inst{2}}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Department of Information Systems and Operations
\and Department of Statistics and Mathematics\\Wirtschaftsuniversität
Wien\\Augasse 2--6, 1090 Vienna, Austria
\texttt{\{David.Meyer,Achim.Zeileis,Kurt.Hornik\}@wu-wien.ac.at}}
\maketitle

\section{Introduction}
\label{sec:intro}

Categorical data analysis is typically based on two- or
higher-dimensional contingency tables, cross-tabulating the
co-occurences of levels of nominal and/or ordinal data. 
In order to explain these, statisticians
typically look for (conditional) independence structures using common
methods such as independence tests and
log-linear models. One idea of visualization techniques is to
use the human visual system to detect structures in the data that
possibly are not obvious from solely numeric output (e.g., test
statistics). Whether the
task is purely exploratory or model based, techniques such as mosaic, sieve, and
association plots offer good support for visualization. They have been
extended over the last two decades, and implementations exist in many
statistical environments.

All three graphical methods visualize aspects of (possibly higher-dimensional)
contingency tables. A \emph{mosaic plot} \cite{vcd:Hartigan+Kleiner:1984} is basically 
an area-proportional visualization of (typically, observed) frequencies, composed
of tiles (corresponding to the cells) created by recursive
vertical and horizontal splits of a rectangle. Thus, the area of each tile
is proportional to the corresponding cell entry \emph{given} the
dimensions of previous splits. 
\emph{Sieve plots} \cite{vcd:riedwyl+schuepbach:1994} 
are similar to mosaic plots, but the area of each
tile is proportional to the \emph{expected} cell entry, and each tile
is filled with a number of rectangles corresponding to the observed value.
An \emph{association plot} \cite{vcd:Cohen:1980}
visualizes the standardized deviations of observed frequencies from
those expected under a certain independence hypothesis.
Each cell is represented by a rectangle that has
(signed) height proportional to the residual and width proportional
to the square root of the expected counts, 
so that the area of the box is proportional
to the difference in observed and expected frequencies.

Over the years, extensions to these techniques have mainly been focused on
the following aspects:

\begin{itemize}
  \item Varying the shape of mosaic plots (as well as bar plots) to yield, e.g.,
  double-decker plots \cite{vcd:Hofmann:2001}, spine 
  plots \cite{vcd:riedwyl+schuepbach:1994}, or spinograms \cite{vcd:hofmann+theus}.
\item Using residual-based shadings to visualize log-linear models
  \cite{vcd:Friendly:1994,vcd:Friendly:2000} and significance of
  statistical tests \cite{vcd:Meyer+Zeileis+Hornik:2003}.
\item Using pairs plots and trellis-like layouts for marginal, conditional and
  partial views \cite{vcd:Friendly:1999}.
\item Adding direct user interaction, allowing quick exploration and
  modification of the visualized models \cite{vcd:Unwin+Hawkins+Hofmann:1996,vcd:Theus:2003}.
\item Providing a modular and flexible implementation to easily allow
  user extensions \cite{vcd:Meyer+Zeileis+Hornik:2005a}.
\end{itemize}

\noindent Current implementations of mosaic displays can be found,
e.g., for \proglang{SAS} \cite{vcd:SAS:2005}, \pkg{ViSta} \cite{vcd:young:1996}, 
\pkg{MANET} \cite{vcd:Unwin+Hawkins+Hofmann:1996},
\pkg{Mondrian} \cite{vcd:Theus:2003},
\proglang{R} \cite{vcd:R:2005}, and \proglang{S-PLUS} \cite{vcd:SPLUS:2005}. 
Implementations of association and sieve plots can only be found in
\proglang{R} and \proglang{SAS} (in the latter, these plots are available
for two-way tables only).
Table~\ref{tab:compare} gives an overview of the available functionality in these systems.
The figures in this chapter have all been produced using the
\proglang{R} system, using the extension packages \pkg{vcd} 
\cite{vcd:Meyer+Zeileis+Hornik:2005} and
\pkg{scatterplot3d} \cite{vcd:ligges+maechler:2003}
(Fig.~\ref{fig:threedee} only), all freely available from the
Comprehensive \proglang{R} Archive Network (\url{http://CRAN.R-project.org/}).

\begin{table}[h]
  \centering
  \begin{tabular}{|l|c|c|c|c|c|}
    \hline
                             & \proglang{SAS} & \proglang{S-PLUS} & \proglang{R} & \pkg{ViSta} & \pkg{MANET}/\pkg{Mondrian}\\\hline
    Basic functionality      &  $\times$  &   $\times$    & $\times$ &   $\times$   & $\times$\\
    Shape                    &     &        & $\times$ &       & $\times$\\
    Residual-based shadings  &  $\times$  &        & $\times$ &   $\times$   & ($\times$)\\
    Conditional Views        &  $\times$  &        & $\times$ &       & $\times$\\
    Interaction              &     &        &   &   $\times$   & $\times$\\
    Extensible Design        &     &        & $\times$ &       &  \\\hline
  \end{tabular}
  \caption{Comparison of current software environments.}
  \label{tab:compare}
\end{table}
%FIXME% For the JSS paper there was criticism about haven MANET and Mondrian in the
%FIXME% same column, how did we decide ro resolve this? Different columns, I suppose...?!

This chapter will give an overview of the state of the art 
of mosaic, sieve, and association plots, both for exploratory visualization and
model-based analysis. Exploratory techniques will include specialized displays for
the bivariate case, as well as pairs plot-like displays for
higher-dimensional tables. As for the model-based tools, particular
emphasis will be given to methods suitable for the visualization of 
conditional independence tests (including permutation tests), as well
as for the visualization of particular GLMs (such as log-linear models). 
In Sect.~\ref{sec:twoway}, we start with the
simple bivariate case.  Section~\ref{sec:extensions} explains how the use of color
in residual-based shadings can support data exploration, and
even promotes the methods to diagnostic and model-based tools
by visualizing test statistics and residuals of independence models.
In Sect.~\ref{sec:multiway}, we show how the
basically bivariate methods straightforwardly extend to the multivariate case by using
`flat' representations of the multi-way tables. In this section, we
also introduce specialized displays for conditional independence
structures. Section~\ref{sec:concl} concludes the chapter.

\section{Two-way Tables}
\label{sec:twoway}

Throughout this section, our examples will be based on
the \data{hospital} data \cite{vcd:wing:1962} given in Tab.~\ref{tab:hospital}.

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|rrr|r|}
\hline
 & \multicolumn{4}{l|}{\textbf{Length of stay (in years)}}\\
\textbf{Visit frequency} &~~2--9 & ~~10--19 & ~~20+ & $\Sigma$ \\
\hline
Regular & 43 & 16 & 3 & 62 \\
Less than monthly & 6 & 11 & 10 & 27 \\
Never & 9 & 18 & 16 & 43 \\\hline
$\Sigma$ & 58 & 45 & 29 & 132 \\
\hline
\end{tabular}
\caption{The \data{hospital} data.}
\label{tab:hospital}
\end{center}
\end{table}

\noindent The table relates the length of stay (in years) of 132
long-term schizophrenic patients in two London mental hospitals with
the frequency of visits (from relatives or friends). 
The length of stay (LOS) has been categorized into
2--9 years, 10--19 years, and more than 19 years. There are also three
categories for the visit frequency: regular (including
%FIXME: lookup regular in Wing-paper.
patients that were allowed to go home), less than monthly, and never.
Wing \cite{vcd:wing:1962} concludes from this data 
that the longer the length of stay in hospital, the less frequent the
visits, which can be seen from the column-standardized table 
(see Tab.~\ref{tab:colhospital}).

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|rrr|}
\hline
 & \multicolumn{3}{l|}{\textbf{Length of stay (in years)}}\\
\textbf{Visit frequency} &~~2--9 & ~~~~~~10--19 & ~~20+ \\
\hline
Regular & 0.74 & 0.36 & 0.10  \\
Less than monthly & 0.10 & 0.24 & 0.35  \\
Never & 0.16 & 0.40 & 0.55  \\\hline
$\Sigma$ & 1.00 & 1.00 & 1.00 \\
\hline
\end{tabular}
\caption{The \data{hospital} data, corrected for the column margin.}
\label{tab:colhospital}
\end{center}
\end{table}

\noindent In addition, \cite{vcd:haberman:1974} notes that this pattern is not
significantly different in the ``less than monthly'' and ``never''
strata. From the row-standardized table (see Tab.~\ref{tab:rowhospital}), 
it seems indeed that LOS is homogeneous with respect to these two visit frequency strata.

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|rrr|r|}
\hline
 & \multicolumn{4}{l|}{\textbf{Length of stay (in years)}}\\
\textbf{Visit frequency} &~~2--9 & ~~10--19 & ~~20+ & $\Sigma$ \\
\hline
Regular & 0.69 & 0.26 & 0.05 & 1.00 \\
Less than monthly & 0.22 & 0.41 & 0.37 & 1.00 \\
Never & 0.21 & 0.42 & 0.37 & 1.00 \\
\hline
\end{tabular}
\caption{The \data{hospital} data, corrected for the row margin.}
\label{tab:rowhospital}
\end{center}
\end{table}

Although far from optimal, contingency tables are frequently visualized using
grouped bar plots (see Fig.~\ref{fig:grouped}) 
or even by means of 3D-bar charts (see Fig.~\ref{fig:threedee}). It seems hard
to detect the aforementioned pattern in these, especially in the 3D
plot where the perspective view tends to distort the true proportions
of the bars. In the following, we
will introduce three graphical methods that are better suited for
contingency tables.

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[p]
\begin{center}
<<grouped,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
barplot(hospital, legend = rownames(hospital), beside = TRUE, 
        xlab = "Length of stay (years)", ylab = "Number of patients")
@
\caption{Bar plot for the \data{hospital} data.}
\label{fig:grouped}
\end{center}
\end{figure}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[p]
\begin{center}
<<threedeebars,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
myhospital = t(hospital)[,3:1]
mydat = data.frame("Length of stay" = as.vector(row(myhospital)), 
                   "Visit frequency" = as.vector(col(myhospital)), 
                   "Number of patients" = as.vector(myhospital))
scatterplot3d(mydat, type = "h", pch = " ", lwd = 10, 
x.ticklabs = c("2-9","","10-19","","20+"), 
              y.ticklabs = c("Never","","Less than monthly","","Regular"), 
              xlab = "Length of stay (years)", ylab = "Visit frequency", zlab = "Number of patients",
              y.margin.add = 0.2,
              color = "black", box = F)
@
\caption{3D-bar chart for the \data{hospital} data.}
\label{fig:threedee}
\end{center}
\end{figure}

\subsection{Mosaic Displays}
\label{sec:mosaic}

Mosaic displays have been introduced by 
\cite{vcd:Hartigan+Kleiner:1981,vcd:Hartigan+Kleiner:1984} and
extended, e.g., by \cite{vcd:Friendly:1994,vcd:Friendly:1999,vcd:Friendly:2000}. 
They visualize the observed values of a contingency table by
area-proportional tiles, arranged in a rectangular mosaic. The tiles are
obtained by recursive partitioning splits of a rectangle.
In the following, we describe the main idea of mosaic plots, 
chapter XXX in this book \emph{\bf (Reference to other contributions)} 
provides more detailed information. Consider our
example of the hospital data from above. Step 1 consists of splitting a square
according to the marginals of one of the variables. To be consistent with the textual
representation, we choose LOS with vertical splits (see Fig.~\ref{fig:mosaic1fig}). 
The result is similar to a bar plot where not the height, but the width is
adapted to visualize the counts for each level, which is also 
called a \emph{spine plot} \cite{vcd:hummel:1996}. From this plot, we see that 
the number of patients 
decreases with the length of stay. Step 2 now is to add further splits in the
other direction, i.e., horizontal splits, for the second
variable. This means that each vertical bar is split according to the
marginals of the second variable, \emph{given} the first variable (see
Fig.~\ref{fig:mosaic2fig}). The resulting plot visualizes the 
contingency table where each cell has a size proportional to the
corresponding table entry.
We can still see the marginal distribution of LOS, %'
and additionally, the visit frequency \emph{given} each category of LOS. %'
If the two variables were independent, the grid would be regular. Clearly, compared
to a length of stay of 10--19 years, more patients get regular visits for stays from 2--9 
years, and conversely, less patients get regular visits for stays for more than 19 years.
For patients that get no visits, the pattern is reversed.

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[p]
\begin{center}
<<mosaic1,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(margin.table(hospital, 2), split = TRUE)
@ 
\caption{Construction of a mosaic plot for a two-way table, step 1.}
\label{fig:mosaic1fig}
\end{center}
\end{figure}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[p]
\begin{center}
<<mosaic2,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(t(hospital), split = TRUE, mar = c(left = 3.5),
       labeling_args = list(offset_labels = c(left = 0.5), 
         offset_varnames = c(left = 1, top = 0.5), set_labels = 
         list("Visit frequency" = c("Regular","Less than\nmonthly","Never"))))
@ 
\caption{Construction of a mosaic plot for a two-way table, step 2.}
\label{fig:mosaic2fig}
\end{center}
\end{figure}

Since mosaic plots are asymmetric by construction, the choice of the
variable order matters, as the first splitting variable dominates the 
plot. In our example, if we use `Visit frequency' %'
as the first splitting variable, the impression is very different compared to the previous
mosaic (see Fig.~\ref{fig:mosaic3fig}). 
In this alternative display, we see the marginal distribution of `Visit frequency' %'
in the rows: about half of the patients get visited regularly. This
group is dominated by patients staying between 2 and 9 years.
It seems apparent that the distribution of LOS %'
is similar for monthly and never visited patients, so this two categories actually
represent one homogenous group (patients visited only casually). Since
the first splitting variable dominates the plot, it should be chosen
as to be the explanatory variable.
%FIXME% causality is not needed, hence omitted: in (hypothesized) causal relationships.

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}
\begin{center}
<<mosaic3,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(hospital)
@ 
\caption{Mosaic plot for the hospital data, using `Visit frequency' %'
as first splitting variable.}
\label{fig:mosaic3fig}
\end{center}
\end{figure}

\subsection{Sieve Plots}
\label{sec:sieve}

%FIXME% Maybe formulate the following theory a bit more like:
%FIXME% "To fix notations, ...", "It is well-known that..."

When we try to explain data, we assume the validity of a certain
model for the generating process. In the case of two-way contingency
tables, the two most common hypotheses \cite{vcd:Agresti:2002} are

\begin{enumerate}
\item independence of the two variables
\item homogeneity of one variable among the strata defined by
the second.
\end{enumerate}

\noindent It is easy to compute the \emph{expected} table under either
of these hypotheses. Consider a 
2-way contingency table with $I$ rows and $J$ columns,
cell frequencies $\{n_{ij}\}$ for
$i = 1, \dots, I$ and $j = 1, \dots, J$, and row and column sums
$n_{i+} = \sum_i n_{ij}$ and $n_{+j} = \sum_j n_{ij}$, respectively. For
convenience, the number of observations is denoted $n = n_{++}$.
Given an underlying distribution with theoretical cell probabilities
$\pi_{ij}$, the null hypothesis of independence of the two categorical
variables can be formulated as
\begin{equation} \label{eq:H0}
H_0: \; \pi_{ij} \quad = \quad \pi_{i+} \pi_{+j}.
\end{equation}

\begin{table}[ht]
\begin{center}
\begin{tabular}{|l|rrr|r|}
\hline
 & \multicolumn{4}{l|}{\textbf{Length of stay (in years)}}\\
\textbf{Visit frequency} & ~~2--9 & ~~10--19 & ~~20+ & $\Sigma$ \\
\hline
Regular & 27.24 & 21.14 & 13.62 & 62 \\
Less than monthly & 11.86 & 9.20 & 5.93 & 27 \\
Never & 18.89 & 14.66 & 9.45 & 43 \\\hline
$\Sigma$ & 58.00 & 45.00 & 29.00 & 132 \\
\hline
\end{tabular}
\caption{The \data{hospital} data---expected values.}
\label{tab:hospital2}
\end{center}
\end{table}

Now, the expected cell frequencies in this model are simply
$\hat n_{ij} = n_{i+}n_{+j}/n$.
The expected table for our sample data is given in Tab.~\ref{tab:hospital2}.
It could again be visualized using a mosaic plot, this
time applied to the table of expected frequencies. 
If we cross-tabulate each tile to fill it with a number of squares equal to
the corresponding number of \emph{observed} frequencies, we get a \emph{sieve plot}
(see Fig.~\ref{fig:sievefig}). It implicitly compares
expected and observed values since the density of the grid will
increase with the deviation of the observed from the expected
values. This allows the detection of general association patterns (for
nominal variables) and of linear association (for ordinal variables),
the latter producing tiles of either very high or very low density 
along one of the diagonals.
In the case of our data, the density of the rectangles is marked along the secondary
diagonal, indicating a negative association of 
the two variables. This gives evidence to the fact that for these patients,
visit frequency decreases with the length of stay.

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}
\begin{center}
<<sievefig,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
sieve(t(hospital), split = TRUE, pop = FALSE, gp = gpar(lty =
"dotted", col = "black"))
labeling_cells(text = t(hospital), clip = FALSE, gp = gpar(fontface = 2, fontsize = 15))(t(hospital))
@ 
\caption{Sieve plot for the \data{hospital} data.}
\label{fig:sievefig}
\end{center}
\end{figure}

\subsection{Association Plots}
\label{sec:assoc}

In the last section, we have seen how to compare observed and expected
values of a contingency table using sieve plots. We can do this
more straightforwardly by using a plot that directly visualizes the residuals. The
most widely used residuals are the Pearson residuals

\begin{equation} \label{eq:Pearson}
r_{ij} \quad = \quad \frac{n_{ij} - \hat n_{ij}}{\sqrt{\hat n_{ij}}}.
\end{equation}

\noindent which are standardized raw residuals. In an 
association plot \cite{vcd:Cohen:1980}, each cell is represented 
by a rectangle that has (signed) height
proportional to the corresponding Pearson residual $r_{ij}$ and width proportional to the
square root of the expected counts $\sqrt{\hat n_{ij}}$. Thus, the area is proportional to
the raw residuals $n_{ij} - \hat n_{ij}$. The sign is visualized by
its position relative to the baseline (upward tiles for positive, and downward 
tiles for negative residuals). Figure~\ref{fig:assoc}
shows the association plot for the hospital data. 
Consistent with the corresponding mosaic and sieve plots, 
we clearly see that too many (too few) patients that stay from 
2 to 9 (more than 19) years get visited regularly
than would be expected under the null of independence, and that this pattern is 
reversed for patients visited less than monthly and that get never visited.

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}
\begin{center}
<<assoc,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
assoc(t(hospital), split = TRUE)
@ 
\caption{Association plot for the \data{hospital} data.}
\label{fig:assoc}
\end{center}
\end{figure}

\subsection{Summary}

Mosaic plots are an instrument to visualize the \emph{observed} frequencies
of a contingency table by recursive conditional splits. If the variables can be
separated into explanatory and dependent variable(s), the explanatory variable(s)
should be be used first for splitting to display the conditional distribution of
the dependent variable(s) given the explanatory variable(s). %FIXME% causality removed again
Sieve plots basically visualize the table of \emph{expected} frequencies, and
in addition the deviations from the observed frequencies by the density of a
grid added to each tile. They complement mosaic plots in detecting dependency
patterns for ordinal variables. An alternative way of enhancing mosaic plots
to display deviations from expected frequencies is to use
residual-based shadings (see next section) which are typically more intellegible
than sieve plots, in particular for nominal variables.
Association plots directly
visualize Pearson and raw residuals, i.e., standardized and
non-standardized deviations of observed from expected frequencies,
respectively. These plots should be used if the diagnostics of
independence models is of primary interest.

\section{Using Colors for Residual-based Shadings}
\label{sec:extensions}

As introduced in the previous section for association plots, the
investigation of residuals from a posited independence model is of
major interest in analyzing contingency tables. In the following, we
will demonstrate how the use of colors can greatly facilitate the
detection of interesting patterns. Before, we start with some general 
remarks on colors and color palettes.

\subsection[A Note on Colors and Color Palettes]{A Note on Colors and Color Palettes\footnote{The electronic version of this article uses colors.}}
\label{sec:colors} 
%FIXME% Maybe extend the footnote and say that the printed version is monochrome,
%FIXME% and that the electronic version uses color.
%FIXME% We can already say earlier that the R code for reproducing the figures
%FIXME% is available from a Web page. And then we can say here that the colored versions
%FIXME% can also be obtained from the Web page.
%FIXME% Where do we want to set up the URL? I would suggest
%FIXME%   http://statmath.wu-wien.ac.at/projects/vcd/

The plots introduced in the previous section are basically composed of
tiles whose areas represent characteristics derived from the
contingency tables---observed and expected frequencies in the case of
mosaic and sieve plots, or residuals as visualized by association
plots. When using color for these tiles, it is imperative
to choose the right color palettes, derived from suitable color
spaces. Apart from aesthetic considerations, wrongly chosen colors
might serously affect the analysis. For example,

\begin{itemize}
\item Using high-chroma colors for large
  areas tends to produce after-image effects which can be
  distracting \cite{vcd:Ihaka:2003}.
\item Lighter colors tend to make areas look larger than darker
  colors, therefore using colors with unequal luminance values  makes it
  difficult to compare area sizes \cite{vcd:Cleveland+McGill:1983}.
\item Color palettes derived from
  non-uniform color spaces may contain unbalanced colors with respect to
  their colorfulness or brightness. When tiles are shaded using such a
  palette, some of them might appear more important than others in an
  uncontrolled way.
\end{itemize}

\noindent 
Due to the three-dimensional nature of human color
perception \cite{vcd:Mollon:1995}, 
it has been common to specify colors using the three primaries
red, green, and blue (RGB colors), especially for computer devices. 
The appearance of the on-screen colors is affected by the device
characteristics: For example, the intensity $I$ of a primary 
on a particular device follows the rule $I=L^\gamma$, with $L$
the value for the primary color, and $\gamma$ device-dependent 
(but typically close to 2.2). Therefore, a first caveat is that if colors are to
appear identical on different devices, their gamma characteristics
have to be taken into account.

Choosing colors and color palettes in RGB space can be a bit inconvenient
and hence software implementations are often based on
Hue-Saturation-Value (HSV) colors (or the comparable Hue-Luminance-Satu\-ration 
color scheme). Both spaces are rather similar transformations of the RGB space
\cite{vcd:Brewer:1999,vcd:Poynton:2000} and are very common implementations of
colors in many computer packages \cite{vcd:Moretti+Lyons:2002}. Each color in HSV space is
represented by three dimensions $(H, S, V)$: the hue $H$ (dominant wavelength in the spectrum, in $[0, 360]$), 
the saturation $S$ (`colorfulness' or `pureness', in $[0, 100]$), and 
the value $V$ (`brightness', amount of gray, in $[0, 100]$).\footnote{In many implementations,
all three dimensions are scaled to the unit interval instead of the coordinates used here.}
These intuitive dimensions
make HSV colors easyier to specifiy than, e.g, RGB colors. 
However, HSV colors have several disadvantages.
Most importantly, HSV colors are not perceptually
uniform because the three HSV dimensions map only poorly to the three
perceptual dimensions of the human visual system  \cite{vcd:Brewer:1999,vcd:Ihaka:2003}.
One important issue here is that HSV dimensions are confounded, e.g., saturation
is not uniform across different hues.
As an example, see Fig.~\ref{fig:colors} (left) 
showing a qualitative color palette (colors $(H, 100, 100)$ for varying hues $H$) in the HSV
space: although saturation and value are fixed, the fully saturated blue is
perceived much darker than the fully saturated red or green,
making it difficult to judge the size of shaded areas.
Furthermore, the flashy fully saturated HSV colors are hard 
to look at for a longer time. For similar reasons, it is equally difficult
to derive acceptable diverging palettes from the HSV space, i.e.,
bipolar scales containing colors ranging between two very distinct colors.
The upper part of Fig.~\ref{fig:diverging} shows a diverging
palette in the HSV space with colors ranging from a saturated red $(0, 100, 100)$
over a neutral white $(H, 0, 100)$ to a saturated blue $(240, 100, 100)$.
Although the palette should be balanced with
respect to colorfulness and brightness, the red colors are perceived
to be more more intense and flashy than the corresponding blue
colors.

\setkeys{Gin}{width=\textwidth}
\begin{figure}
\begin{center}
<<colorfig,fig=TRUE,echo=FALSE,height=3,width=7,results=hide>>=
par(mfrow = c(1,2), mar = c(1,1,1,1), oma = c(0,0,0,0))
pie(rep(1,9), radius = 1, col = rainbow(9), labels = 360 * 0:8/9)
pie(rep(1,9), radius = 1, col = rainbow_hcl(9), labels = 360 * 0:8/9)
@ 
\caption{Qualitative color palette for the HSV (left) 
         and HCL (right) spaces. The HSV colors are $(H, 100, 100)$ and
	 the HCL colors $(H, 50, 70)$ for the same hues $H$. Note that
         in a monochrome version of this paper, all pies in the right
         wheel will be shaded with the same gray, i.e., appear identical.}
\label{fig:colors}
\end{center}
\end{figure}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}
\begin{center}
<<diverging,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
par(mfrow = c(1,1), mar = c(1,1,1,1), oma = c(0,0,0,0))
plot.new()
rect(0:4/5, 0.2, 1:5/5, 0.5, border = 0, col = diverge_hcl(5))
rect(0, 0.2, 1, 0.5, border = 1, col = NULL)
rect(0:4/5, 0.55, 1:5/5, 0.85, border = 0, col = diverge_hsv(5))
rect(0, 0.55, 1, 0.85, border = 1, col = NULL)
text(c(1:5/5 - 0.1), 0.11, c("(260, 100, 50)", "(260, 50, 70)", "(H, 0, 90)", "(0, 50, 70)", "(0, 100, 50)"))
text(c(1:5/5 - 0.1), 0.91, c("(240, 100, 100)", "(240, 50, 100)", "(H, 0, 100)", "(0, 50, 100)","(0, 100, 100)"))
@ 
\caption{Diverging color palettes for the HSV space (upper part) and
  the HCL space (lower part), ranging from blue over a neutral color to
  red. The triples indicate the settings for 
  the three dimensions: (Hue, Saturation, Value)
  in the upper part, and (Hue, Chroma, Luminance) in the lower part.
  In a monochrome version of the paper, the right and left hand
  sides of the HCL colors palette appear identical. }
\label{fig:diverging}
\end{center}
\end{figure}

The use of colors that are more `in harmony' goes back to Munsell 
\cite{vcd:Munsell:1905} %'
who introduced a notation for balanced colors. Based on those,
tools producing better palettes for specific tasks have been developed 
\cite{vcd:Harrower+Brewer:2003}. Other perceptually-based color
spaces, especially suited for computer displays, are the CIELAB and CIELUV
spaces \cite{vcd:CIE:2004} from which qualitative palettes for
statistical graphics have been derived \cite{vcd:Ihaka:2003}. A
transformation of the CIELUV space leads to the HCL (Hue-Chroma-Luminance) space.
These colors are again specified by triplets $(H, C, L)$:
chroma $C$ loosely corresponds to colorfulness and luminance $L$ to
brightness, but in contrast to HSV colors, chroma is an
\emph{absolute} measure valid for all hues, and luminance can be
varied independently from the other two dimensions.
Qualitative color palettes can easily be obtained by holding constant
chroma and luminance, and using varying hues. HCL colors with fixed luminance 
are always balanced towards the same grey 
and thus do not have the problem of varying saturations as the HSV
colors (see Fig.~\ref{fig:colors}, right). 
Similarly, diverging HCL color palettes can be derived
\cite{vcd:Zeileis+Meyer+Hornik:2005} by interpolating again between a
neutral color such as $(H, 0, 90)$ and two colors with full chroma
such as blue $(260, 100, 50)$ and red $(0, 100, 50)$. The resulting
palette is shown in the lower part of Fig.~\ref{fig:diverging}. 
In contrast to the HSV counterpart, matching
colors (light red/blue and dark red/blue) 
are balanced to the same gray, and thus receive the same perceptual ``weight''.
Note, that chroma and luminance are varied simultaneously, i.e.,
the full chroma colors are also darker than the neutral color. Of course,
it would also be possible to choose a darker neutral color (or lighter 
full chroma colors), however, by varying both chroma and luminace better 
contrasts can be achieved.

\subsection{Highlighting and Color-based Shadings}
\label{sec:shadings}

All plots introduced in Sect.~\ref{sec:twoway} are composed of empty
tiles. It seems intuitive to use filled tiles to highlight information
of interest. Consider again the hospital example: in Fig.~\ref{fig:highlight},
we mark tiles for patients never or seldom visited to visualize their proportion in
the LOS strata. For optical clarity, we set the
spacing between the visit frequency tiles to 0. Clearly, the
proportion of these patients increases with LOS. In fact, such a
``stacked'' plot can also be interpreted as spine plot with highlighting
\cite{vcd:hummel:1996}, particularly useful in analyzing relationships among
categorical data with a binary dependent variable. More variations on
this theme, such as doubledecker plots, are treated in chapter XXX
(\textbf{REFERENCE TO OTHER CHAPTER(S)}).

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}
\begin{center}
<<highlight,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(t(hospital), 
       mar = c(left = 3.5), 
       labeling_args = list(offset_labels = c(left = 0.5), 
         offset_varnames = c(left = 1, top = 0.5), set_labels = 
         list("Visit frequency" = c("Regular","Less than\nmonthly","Never"))),
       split = TRUE, spacing = spacing_highlighting, 
       gp = gpar(fill = rep(grey.colors(2)[2:1], 1:2)))
@ 
\caption{Spine plot with highlighting for the hospital data.}
\label{fig:highlight}
\end{center}
\end{figure}

Using colors, even more complementary information can be visualized,
either by adding additional information, or by redundantly coding information
already visualized by the `raw' plot to support our perceptual system. %'
First, we consider the sieve plots. The density 
of the grid in the raw version implicitly gives us an 
idea of the residuals' %'
size, but since the plot does not include the density corresponding to zero
residuals (the null model) for comparison, we 
cannot easily assess whether there are more, of less counts in a cell 
than expected under the null hypothesis. 
%It would help if we knew the sign of the residuals. 
Using color, we can add the sign information, for example using blue for
positive, red for negative, and gray for zero residuals.

% Figure~\ref{fig:colorsieve} 
% demonstrates the effect of applying such a color shading to a sieve plot for the 
% \data{hospital} data. Color clearly emphasizes the linear association
% pattern in this data: Patients with a LOS from 2--9 years are
% associated with regular visits, whereas patients with a LOS from
% 10--19 and more than 19 years are associated with a visit frequency
% of less than monthly and no visits.

% \setkeys{Gin}{width=0.7\textwidth}
% \begin{figure}
% \begin{center}
% <<sievecolor,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
% sieve(t(hospital), split = TRUE, shade = TRUE)
% @ 
% \caption{Sieve plot with color coding of the residuals.}
% \label{fig:colorsieve}
% \end{center}
% \end{figure}

Mosaic plots in their initial version are monochrome
displays. Friendly \cite{vcd:Friendly:1994} introduced a
residual-based shading of the tiles to additionally visualize the
residuals from a given independence model fitted to the table. 
The idea is to use a color coding for the mosaic tiles that visualizes the
sign and absolute size of each residual $r_{ij}$: Cells corresponding to small
residuals ($|r_{ij}| < 2$) have no color. Cells with medium sized residuals
($2 \le |r_{ij}| < 4$) are shaded light blue and light red for positive and negative
residuals, respectively. Cells with large residuals ($|r_{ij}| \ge 4$) are shaded
with a fully saturated blue and red, respectively. 
The heuristic for choosing the cut-offs $2$ and $4$ is that the Pearson residuals
are asymptotically standard normal which implies that the highlighted cells are those with
residuals \emph{individually} significant at approximately the $\alpha
= 0.05$ and $\alpha = 0.0001$ levels. However, the main purpose of the
Friendly shading is not to visualize significance but the \emph{pattern} of
deviation from independence \cite{vcd:Friendly:2000}. 
In addition to the shading of the rectangles themselves, the Friendly shading also
encompasses a choice of line type and line color of the rectangle borders
with similar ideas as described above, useful when no color can be used.

In Fig.~\ref{fig:Friendly}, we show again the mosaic for the \data{hospital}
data, this time using a Friendly-like color shading (with HCL
instead of HSV colors, and no line type coding).
\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}
\begin{center}
<<friendly,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(t(hospital), split = TRUE, shade = TRUE, 
       mar = c(left = 3.5), 
       gp_args = list(p.value = hos_chisq$p.value),
       labeling_args = list(offset_labels = c(left = 0.5), 
         offset_varnames = c(left = 1, top = 0.5), set_labels = 
         list("Visit frequency" = c("Regular","Less than\nmonthly","Never"))))
@ 
\caption{Mosaic display with Friendly-like color coding of the residuals.}
\label{fig:Friendly}
\end{center}
\end{figure}
Clearly, the asymmetry for regular and never visited students,
and the pattern inversion for lengths of stay of 2--9 and more than 19
years are emphasized using the color shading.

For association plots, residual-based shadings are redundant since all relevant
information is already contained in the plot by construction. Nevertheless, using one of the
shadings discussed above will support the analysis process and is therefore recommended.
For example, applying the Friendly shading in Fig.~\ref{fig:colorassoc} 
on the one hand facilitates the discrimination between
positive and negative residuals, and on the other, more importantly,
allows an easier comparison of the tiles' sizes. Thus, the shading
supports the detection of the pattern.

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}
\begin{center}
<<assoccolor,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
assoc(t(hospital), split = TRUE, shade = TRUE, keep = TRUE,
      gp_args = list(p.value = hos_chisq$p.value))
@ 
\caption{Association plot with Friendly-like color coding of the residuals.}
\label{fig:colorassoc}
\end{center}
\end{figure}

\subsection{Visualizing Test Statistics}
\label{sec:test}

Figures~\ref{fig:Friendly} and \ref{fig:colorassoc} 
include the (same) $p$ value of a $\chi^2$ test
of independence, frequently used to assess the significance of the
hypothesis of independence (or homogeneity for stratified data) in
two-way tables. The test statistic is just the sum of the squared
Pearson residuals

\begin{equation} \label{eq:ChiSquare}
X^2 \quad = \quad \sum_{i, j}r_{ij}^2~,
\end{equation}

\noindent known to have a limiting $\chi^2$ distribution with $(I-1)(J-1)$
degrees of freedom under the null hypothesis.
An important reason for using the unconditional limiting distribution
for the $X^2$ statistic from (\ref{eq:ChiSquare}) was the closed form result
for the distribution. Recently, with the improving perfomance of computers,
conditional inference (or permutation tests)---carried out either by simulation
or by computation of the (asymptotic) permutation distribution---have been receiving
increasing attention \cite{vcd:Ernst:2004,vcd:Pesarin:2001,vcd:Strasser+Weber:1999}. 
For testing the independence hypothesis from (\ref{eq:H0}),
using a permutation test is particularly intuitive due to the 
permutation invariance (given row and column sums) of this problem.
Consequently, all results in this paper are based on conditional inference performed
by simulating the permutation distribution of test statistics of type
$\lambda([r_{ij}])$.

Since the HCL space is three-dimensional and we only
used two `degrees of freedom' so far for coding information %'
(hue for the sign and a linear combination of chroma and luminance 
for the size of the residuals), we can add a third piece of information to the plot.
For example, we can visualize the significance of some specified test statistic
(e.g., the $\chi^2$ test statistic) 
using less colorful (`uninteresting') colors for non-significant
results. These can again be derived using the same procedure
described in Sect.~\ref{sec:colors} but using a smaller amount of color, i.e.,
a smaller maximal chroma (e.g., 20 instead of 100).

% The following example using the \data{Bundesliga} data \cite{vcd:Knorr-Held:1999}
% shows the relationship of home goals and away
% goals of Germany's premier soccer league in 1995: although there are two
% ``larger'' residuals (one greater than 2, one less then $-2$), the
% $\chi^2$ test does not reject the null hypothesis of
% independence on a significance level of $5\%$ since the $p$ value is $0.1213$. 
% Consequently, the employed shading uses a darker overall color scheme
% (see Fig.~\ref{fig:bundesliga}).

% \begin{figure}[p]
% \begin{center}
% <<bundesligafig,fig=TRUE,echo=FALSE,height=6,width=7>>=
% bl <- xtabs(~ HomeGoals + AwayGoals, data = Bundesliga, subset = Year == 1995)
% mosaic(bl, shade = TRUE)
% @
% \caption{Mosaic plot for part of the \data{Bundesliga}, using a $\chi^2$ test and
% fixed cut-off points for the residuals.}
% \label{fig:bundesliga}
% \end{center}
% \end{figure}

The heuristic for choosing the cut-off points 
in the Friendly shading may lead to wrong conclusions: especially in
large tables, the test of
independence may not be significant even though some of the residuals are
``large''. On the other hand, the test might be significant even
though the residuals are ``small''. In fact, the cut-off points are really
data-dependent. Consider the case of the \data{arthritis} data 
\cite{vcd:Koch+Edwards:1988}, resulting from a double-blind clinical trial
investigating a new treatment for rheumatoid arthritis, stratified by
gender (see Tab.~\ref{tab:arthritis} for the female patients):

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|rrr|r|}
\hline
 & \multicolumn{4}{c|}{\textbf{Improved}}\\
\textbf{Treatment} & None & Some & Marked & $\Sigma$ \\
\hline
Placebo & 19 & 7 & 6 & 32 \\
Treatment & 6 & 5 & 16 & 27 \\\hline
$\Sigma$ & 25 & 12 & 22 & 59 \\
\hline
\end{tabular}
\caption{The \data{arthritis} data (female patients).}
\label{tab:arthritis}
\end{center}
\end{table}

\noindent Figure~\ref{fig:arthritis} visualizes the results for the female
patients, again by means of a mosaic display. Clearly, the hypothesis
of independence is rejected by the $\chi^2$ test even on a $1\%$ level 
($p=\Sexpr{round(art_chisq$p.value,4)}$), but
since all residuals are in the 
$[\Sexpr{round(min(chisq.test(art)$residuals),4)},\Sexpr{round(max(chisq.test(art)$residuals),4)}]$ %$
interval, the tiles remain uncolored. A solution to this issue is to use a different
test statistic, for example the maximum of the absolute values of the
Pearson residuals \cite{vcd:Meyer+Zeileis+Hornik:2003} instead of the sum of squares:

\begin{equation} \label{eq:MaxAbs}
M \quad = \quad \max_{i, j} |r_{ij}|~.
\end{equation}

\noindent Given a critical value $c_\alpha$ for this test statistic, all residuals whose
absolute value exceeds $c_\alpha$ violate the hypothesis of indendence at level
$\alpha$ \cite[ch. 7]{vcd:Mazanec+Strasser:2000}. Thus, the interesting cells giving
evidence for the rejection of the independence hypothesis can easily be identified.
As explained above, 
the conditional distribution of this test statistic under the null can be obtained
by simulation, sampling tables with the same row and column sums $n_{i+}$ and $n_{+j}$
using, e.g., the Patefield algorithm \cite{vcd:Patefield:1981} 
and computing the maximum statistic for each of these tables.
In Fig.~\ref{fig:arthritis2}, we visualize again the \data{arthritis} data, this
time using the maximum test statistic and its 10\% and 1% critical values at cut-off points.
The shading of the
tiles now clearly shows that the treatment is effective: significantly more patients
in the treatment group exhibit marked improvement than would be
expected under independence.

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[p]
\begin{center}
<<Arthritis,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
mosaic(art, shade = TRUE, gp_args = list(p.value = art_chisq$p.value))
@
\caption{Mosaic plot for the \data{arthritis} data, using the $\chi^2$ test 
and fixed cut-off points for the shading.}
\label{fig:arthritis}
\end{center}
\end{figure}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[p]
\begin{center}
<<Arthritis2fig,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
set.seed(myseed)
mosaic(art, gp = shading_max, gp_args = list(n = 5000))
@
\caption{Mosaic plot for the \data{arthritis} data, using the maximum test and 
data-driven cut-off points for the residuals.}
\label{fig:arthritis2}
\end{center}
\end{figure}

\subsection{Summary}

For the visualization of areas perceptually uniform colors and color palettes
should always be used and HCL color space provides a convenient way to do so.
Shadings can be used to add information to the basic plots and to support the
analysis. Highlighting of tiles can support the analysis relationships with a 
single dependent variable. Residual-based shadings can be used for model diagnostics, e.g.,
by using diverging color palettes to code the sign and size of a residual.
In addition, the significance of test statistics can be visualized by using 
overall less colorful palettes. Using the maximum instead of the $\chi^2$ (or other) test
statistic(s) allows for data-driven cut-offs in the diverging palettes and
precise diagnostic identification of the cells in conflict with the
null hypothesis of independence.

\section{Selected Methods for Multi-way Tables}
\label{sec:multiway}

In Sect.~\ref{sec:twoway}, we have presented basic displays
for two-way tables, based on the visualization of information on 
the tables' cells, arranged in rectangular form. For multi-way tables,
mosaic plots can directly be used by simply adding further splits for each 
additional variable. For sieve and association plots,
we apply the basic idea of mosaic plots to the table itself,
i.e., simply nest the variables into rows and columns using recursive
conditional splits, given the margins. The result is a 
`flat' representation of the multi-way table that can be visualized in %'
ways similar to a two-dimensional table.

In the following, we will first treat specialized displays for
explorative purposes, followed by model-based methods for
conditional independence models.

\subsection{Explorative Visualization Techniques}
\label{sec:flat}

As an example, consider the well-known UCB admissions data 
\cite{vcd:Bickel+Hammel+O'Connell:1975}
on applicants to graduate school at the University of California in Berkeley for
the six largest departments in 1973 classified by admission and
gender. The `flattened' contingency table, putting department into the
columns and admit---nested in gender---into the rows, is given by
Tab.~\ref{tab:ucb}. The table aggregated over all
departments gives wrong evidence of gender bias, 
i.e., higher admission rates for male students.

\begin{table}[ht]
\begin{center}
\begin{tabular}{|ll|rrrrrr|}
\hline
 & & \multicolumn{6}{c|}{\textbf{Department}} \\
\textbf{Gender} & \textbf{Admit} & A & B & C & D & E & F\\\hline
Male   & Admitted & 512 & 353 & 120 & 138 & 53 & 22\\
       & Rejected & 313 & 207 & 205 & 279 & 138 & 351\\
Female & Admitted & 89 & 17 & 202 & 131 & 94 & 24 \\
       & Rejected & 19 & 8 & 391 & 244 & 299 & 317 \\
\hline
\end{tabular}
\caption{The \data{UCB admissions} data, in flat representation.}
\label{tab:ucb}
\end{center}
\end{table}

A first step in the explorative analysis of more complex tables is to
get a quick overview on the data. For that, 
all basic plots can be combined in pairwise displays,
arranged in in a matrix similar to scatterplots in a pairs plot.
The diagonal cells contain the variable names, optionally with
univariate statistics, whereas the off-diagonal cells contain plots whose
variables are implicitly specified by the cells' position in the matrix. %'
In Fig.~\ref{fig:pairs}, the diagonal cells show bar plots for the
distribution of the variables, and the off-diagonal cells mosaic plots
for the corresponding pair of variables. The plots suggest that
admission differs between male and female students, and between the departments,
and that the proportion of male and female students varies across the departments.

\setkeys{Gin}{width=\textwidth}
\begin{figure}
\begin{center}
<<pairs,fig=TRUE,echo=FALSE,height=8,width=8,results=hide>>=
pairs(UCBAdmissions)
@
\caption{Pairs plot for the \data{UCB admissions} data.}
\label{fig:pairs}
\end{center}
\end{figure}

%FIXME% Didn't we decide to use fewer plots in the paper and simply
%FIXME% point out that other combinations could also be used?
%FIXME% I would use for UCB:
%FIXME%   o explorative: highlighted doubledecker (currently Fig. 17)
%FIXME%   o model-based: conidtional shaded assoc (currently Fig. 20)
%FIXME% For Punishment:
%FIXME%   o explorative: highlighted mosaic( currently Fig. 21)
%FIXME%   o model-based: conditional shaded mosaic (currently Fig. 23)
%FIXME% The others should be omitted (I think) and the space gained should
%FIXME% be used for making the individual graphs larger. Preferably, we
%FIXME% should have the explorative and model-based plots for the same data
%FIXME% set on the same page, emphasizing that different aspects of the data
%FIXME% are brought out by the graphs.

The next step is to investigate three-way interactions among the variables.
In this example, we have a binary variable of interest (admission)
that shall be ``explained'' by the others. 
A possible way of representing such a three-way table is 
to draw a mosaic plot of gender and
department and to highlight admission in the tiles (see Fig.~\ref{fig:ucbhighlight}).
From the widths of the tiles it can be seen that an
unequal number of students apply for the six
departments, and from the highlighting we see that the admission rates differ as
well among the departments. The rates 
are equal for male and female students, except for department A where
\textit{more} female than male students are admitted. This result
is even better depicted by a doubledecker plot of admission, explained
by gender given department (see Fig.~\ref{fig:doubleucb}): here, the
difference of admission rates in department A is more marked.

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}
\begin{center}
<<ucbhighlight,fig=TRUE,echo=FALSE,height=6,width=9,results=hide>>=
mosaic(aperm(UCBAdmissions), gp = gpar(fill = grey.colors(2)),
   spacing = spacing_highlighting, split = TRUE, keep = FALSE, rep = FALSE)
@
\caption{Mosaic plot with highlighting for the \data{UCB admissions} data.}
\label{fig:ucbhighlight}
\end{center}
\end{figure}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}
\begin{center}
%FIXME% right margin labels get clipped
<<doubleucb,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
doubledecker(aperm(UCBAdmissions), col = rev(grey.colors(2)))
@
\caption{Dobledecker plot for the \data{UCB admissions} data.}
\label{fig:doubleucb}
\end{center}
\end{figure}

\subsection{Model-based Displays for Conditional Independence Models}
\label{sec:condind}

In addition to the explorative approaches 
to the visualization of multi-way tables outlined in 
the previous section, there are specialized displays 
designed for the visualization of conditional independence structures.

A first approach is to use pairs plots for model search:
we can use the position of the cells in the pairs matrix to select an
independence model and add a corresponding shading to the tiles to
visualize the residuals.
More precisely, each cell $a_{ij}$ in such a matrix defines two
variables $i$ and $j$ that can be used to specify the model visualized
in that cell. Typical hypotheses are: Variables $i$ and $j$ are
marginally independent; variables $i$ and $j$ are conditionally
independent, given all others; variables $i$ and $j$ are jointly
independent from all others, etc.
% Figure~\ref{fig:pairs} shows a pairs display with mosaic plots visualizing mutual 
% independence in the upper triangle, 
% mosaics for \emph{joint} independence 
% in the lower triangle, and bar charts on the diagonal.
% In the upper part of the matrix, we can immediately 
% detect the independence of hair and gender, and eye and gender from the
% non-colored cells in the corresponding mosaic plots. On the other hand, hair and eye color 
% clearly are associated. In the lower part of the matrix, the mosaic
% at the intersection of hair and eye is the less colored one. Its shading
% represents the model of hair and eye color being jointly independent from
% gender, i.e., their association being homogeneous over the gender
% strata. It is the model that fits best, although still rejected at the
% 5\% level 
% ($p=\Sexpr{round(summary(loglm(~ Hair * Eye + Sex, data = HairEyeColor))$tests[2,3],4)}$): %$
% possibly, because too many brown-haired male
% students have blue eyes, and too few of them have brown eyes, than expected under
% the null model?

In our example, from the explorative analysis, we already suspect that
admission and gender are independent, given department, except for
department A. 
Therefore, a sensible way of representing the table is to 
nest admission into department instead of gender to get a stratified view.
The corresponding mosaic and association plots are shown
in Figs.~\ref{fig:ucbmosaic}--\ref{fig:ucbassoc}. The shading visualizes the 
model of admission and gender being conditionally independent, given department. Clearly,
there is no gender bias in the departments except for department A: here, 
significantly \textit{more} female students are admitted then expected under independence.
Mosaic displays are recognized as an excellent
visualization tool for log-linear models 
\cite{vcd:Friendly:1999,vcd:Theus+Lauer:1999,vcd:Hofmann:2001}.
Using the flat table representation, the association plot similarly extends from
the simple two-way case to a structured residual plot usable for 
the diagnostics of log-linear models. Especially if a
mosaic display contains very small or empty cells, the corresponding association plot
might provide an easier way to detect deviation patterns than the
residual-based shading added to the mosaic plot.

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}
\begin{center}
<<ucbmosaic,fig=TRUE,echo=FALSE,height=6,width=12,results=hide>>=
mosaic(aperm(UCBAdmissions, c(3,2,1)), data = UCBAdmissions, split = TRUE,
shade = TRUE, keep = FALSE, residuals = ucb_max$residuals,
gp_args = list(p.value = ucb_max$p.value)) 
@
\caption{Mosaic plot for the \data{UCB admissions} data.}
\label{fig:ucbmosaic}
\end{center}
\end{figure}

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}
\begin{center}
<<ucbassoc,fig=TRUE,echo=FALSE,height=6,width=12,results=hide>>=
assoc(aperm(UCBAdmissions, c(3,2,1)), split = c(TRUE,FALSE,TRUE), shade = TRUE, 
residuals_type = "Pearson", residuals = ucb_max$residuals,
gp_args = list(p.value = ucb_max$p.value))
@
\caption{Association plot for the \data{UCB admissions} data.}
\label{fig:ucbassoc}
\end{center}
\end{figure}

% As an example, consider the \data{PreSex} data
% \cite{vcd:thornes+collard:1979,vcd:gilbert:1981} on
% pre- and extra-marital sex and divorce: 494 divorced and 542 married
% people (willing to divorce) were asked two questions about their pre- and
% extramarital sexual experiences, resulting in Tab.~\ref{tab:presex}.

% \begin{table}[ht]
% \begin{center}
% \begin{tabular}{|ll|rrrrr|}
% \hline
%  & & \textbf{Marital Status} & \multicolumn{2}{l}{Divorced} & \multicolumn{2}{l|}{Married}  \\
% \textbf{Gender} & \textbf{Extramarital Sex} & \textbf{Premarital Sex} & Yes & No & Yes & No\\\hline
% Woman & Yes && 17 & 36 &  4 & 4 \\
%       & No  && 54 &214 & 25 &322 \\
% Men   & Yes && 28 & 17 & 11 & 4 \\
%       & No  && 60 & 68 & 42 &130 \\
% \hline
% \end{tabular}
% \caption{The \data{PreSex} data, in flat representation.}
% \label{tab:presex}
% \end{center}
% \end{table}

% \noindent Figure~\ref{fig:presex}
% illustrates the conditional independence of premarital and
% extramarital sex, given gender and marital status. The $\chi^2$
% test of independence rejects the null hypothesis: possibly, because
% too many men who had premarital sex also had extramarital sex? For the conditioned
% variables premarital and extramarital sex, we use a fixed spacing, whereas for 
% the conditioning variables gender and marital status, the spacing is increasing.

% \setkeys{Gin}{width=0.7\textwidth}
% \begin{figure}
% \begin{center}
% <<PreSex,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
% mosaic(PreSex, condvars = c(1, 4), shade = TRUE)
% @
% \caption{Mosaic plot for the pre- and extramarital sex data.}
% \label{fig:presex}
% \end{center}
% \end{figure}

% Since mosaic displays are conditional plots by construction, we can use them
% for stratified data by using the conditioning variables as first splitting variables. 
% The mosaic in Fig.~\ref{fig:ucbmosaic} shows the distribution of admission
% and gender for all departments. A larger spacing for
% the department variable has been used to simplify the distinction between the 
% groups. 
As we have seen, we can use all plots for the analysis of stratified
data by using the conditioning variables first for splitting.
However, using a single plot, the strata are not corrected for the
marginal distribution of the stratifying variables. Especially for
very unevenly distributed marginals, the strata can become very
distorted, and the tiles, along with their shading, inscrutable.
An alternative is to use trellis-like layouts for visualizing \emph{partial}
tables for given strata, defined by the conditioning variables. 
All subtables are corrected for the marginals and thus, all
corresponding mosaics in the panels have the same size. 
In addition, trellis layouts help reducing the complexity of bigger tables. 
Figure~\ref{fig:ucbadmissions} 
visualizes the data by means of a conditional 
association plot. Each panel corresponds to a department, and contains
an association plot corresponding to the partial (fourfold) table of admission
and gender. Accordingly, the shading visualizes the
residuals from the corresponding conditional independence model (independence of gender
and admission, given department), stratified by department. Clearly,
the situation in department A (more women/less men admitted than
expected under the null hypothesis) causes the rejection of the
hypothesis of conditional independence. 
% The overall deviation pattern
% thus indicates that in contrast to the impression given by the aggregated table, male
% students were not advantaged, and that in department A,
% \emph{female} students were even more successful than their male colleagues.

\setkeys{Gin}{width=\textwidth}
\begin{figure}
\begin{center}
<<UCBAdmissions,fig=TRUE,echo=FALSE,height=6,width=10,results=hide>>=
cotabplot(aperm(UCBAdmissions, c(2,1,3)), panel = cotab_coindep, shade = TRUE, 
          legend = FALSE, margins = c(3,1,1,3), panel_args = list(type = "assoc"))
@
\caption{Conditional mosaic plot for the \data{UCB admissions} data.}
\label{fig:ucbadmissions}
\end{center}
\end{figure}

\subsection{A four-way example}

As a four-way example, we use the punishment data \cite{vcd:andersen:1991} 
from a study of the Gallup Institute in Denmark in 1979 about
the attitude of a random sample of 783 persons towards corporal
punishment of children (see Tab.~\ref{tab:punish}).      
The data consists of four variables: Attitude is a binary variable
indicating whether a person approves no or moderate punishment of children.
Memory indicates whether the person has got memories of 
corporal punishment as a child. Education indicates the
highest level of education (elementary, secondary, high). Finally, 
age indicates the age group in years (15--24, 25--39, $40+$).

\begin{table}[ht]
\begin{center}
\begin{tabular}{|ll|rrrrrrr|}
\hline
 & & \textbf{Age} & \multicolumn{2}{c}{15--24} & \multicolumn{2}{c}{25--39} & \multicolumn{2}{c|}{40+}  \\
\textbf{Education} & \textbf{Attitude} & \textbf{Memory} & Yes & No & Yes & No & Yes & No\\\hline
 Elementary & No && 1 & 26 &  3 & 46&20&109 \\
     & Moderate  && 21 &93 & 41 &119&143&324 \\
Secondary   & No && 2 & 23 & 8 & 52&4&44 \\
     & Moderate  && 5 & 45 & 20 &84&20&56 \\
     High   & No && 2 & 26 & 6 & 24&1&13 \\
      & Moderate && 1 & 19 & 4 &26&8&17 \\
\hline
\end{tabular}
\caption{The \data{punishment} data.}
\label{tab:punish}
\end{center}
\end{table}

%FIXME% Better structure:
%FIXME%   o dependent variable: attitude
%FIXME%   o dependence on age and education is rather clear
%FIXME%     (approval of moderate punishment increases with age and
%FIXME%     decreases with education)
%FIXME%   o question is conditional independence of attitude and
%FIXME%     memory given the other margins
%FIXME%   o hence for exploratory analysis: split in conditioning
%FIXME%     variables (age, education) first, then explanatory variable (memory),
%FIXME%     then dependent variable (attitude) for which highlighting is used.
%FIXME%   o for model-based view: suppress marginal distribution of
%FIXME%     conditioning variables (and their interaction), split first
%FIXME%     in explanatory variable in each cell, use shading based on
%FIXME%     residuals in conditional independence model.
%FIXME%   o advantage of exploratory view: joint distribution of all variables
%FIXME%     can be seen. disadvantage: visualization strongly influence
%FIXME%     by the marginal distribution of education over age (in particular
%FIXME%     the large proportion of 40+ people with elementary education)
%FIXME%     which is not of interest in the conditional independence problem.

To get an overview, we first plot a mosicplot of memory, age, and
education, and highlight attitude (see Fig.~\ref{fig:punishhigh}).
We can see that half of the people have more than 40 years of
age. Most people in this group have only completed elementay
school. For people over 25, a majority has completed secondary
school. Among the younger people, fewer have memories of having been
punished as a child than among the older ones. But sad enough, 
for people with elementary or secondary school education,
a higher proportion \emph{with} memories tends to
have a moderate attitude to punishment than those without.
However, for people under 40, the proportion with
no attitude to punishment increases with the educational level.

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}
\begin{center}
<<punishhigh,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
tab <- xtabs(Freq ~ memory + attitude + age + education, data = Punishment)
mosaic(~ age + education + memory + attitude, data = tab, keep = FALSE,
       gp = gpar(fill = grey.colors(2)), spacing = spacing_highlighting, 
       rep = c(attitude = F), set_varnames = c(attitude = "attitude to punishment"),
       set_labels = list(age = c("15-24","25-39","40+")))
@
\caption{Mosaic plot with highlighting for the \data{punishment} data.}
\label{fig:punishhigh}
\end{center}
\end{figure}

So both age and education seem to matter here. Is there any influence
of memory on attitude to punishment apart from that? To answer this
question, we test the hypothesis of memory and attitude being
conditionally independent, given age and education, and visualize the
model using an extended mosaic plot (see Fig.~\ref{fig:punishmosaic}). 
We can see that the model fits rather good, except for people over 25
with elementary school education: under the null, more people should
say ``no'' to physical punishment of children. This pattern can again also
be visualized using a trellis layout, this time using partial mosaic
plots (see Fig.~\ref{fig:punishmosaic}).

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}
\begin{center}
<<punishmosaic,fig=TRUE,echo=FALSE,height=6,width=7,results=hide>>=
tab_test <- coindep_test(tab, margin = c(3, 4))
mosaic(~ age + education + memory + attitude, data = tab, keep = FALSE,
       expected = ~ (memory + attitude) * education * age,
       gp_args = list(interpolate = tab_test$qdist(c(0.9, 0.99))), 
       rep = c(attitude = F), set_varnames = c(attitude = "attitude to punishment"),
       set_labels = list(age = c("15-24","25-39","40+")))
@
\caption{Mosaic plot for the \data{punishment} data.}
\label{fig:punishmosaic}
\end{center}
\end{figure}

\setkeys{Gin}{width=\textwidth}
\begin{figure}
\begin{center}
<<punishcotab,fig=TRUE,echo=FALSE,height=6,width=10,results=hide>>=
cotabplot(tab, panel = cotab_coindep, panel_args = list(margin = c(2.5, 0, 0.5, 0)))
@
\caption{Conditional mosaic plot for the \data{punishment} data.}
\label{fig:punishcotab}
\end{center}
\end{figure}

\subsection{Summary}

Mosaic, association, and sieve plots can be used for multi-way tables
by applying them to flat representations. In addition, these basic
plots are leveraged by specialized frameworks 
for exploratory model search (pairs plots, highlighting, doubledecker plots) 
and stratified analysis (conditional plots).

\section{Conclusion}
\label{sec:concl}

This chapter reviews several alternatives for the visualization of
multi-way contingency tables. For two-way tables, mosaic, sieve, and
association plots are suitable for the visualization of observed and
expected values and the Pearson residuals, respectively. This basic
methods can be enhanced by using residual-based shadings, preferably
based on perceptual color palettes such as those derived from the HCL
space. Residual-based shadings can be used to visualize sign and size
of the residuals, as well as the significance of test statistics such
as the $\chi^2$ or the maximum test statistic. The latter has the
advantage of detecting residuals causing the rejection of the hypothesis
of independence. The methods directly extend to the multi-way case by
using `flat' representations of the multi-way tables, and specialized
displays for conditional independence such as trellis layouts of
partial tables and pairs plots.

\bibliographystyle{plain}
\bibliography{vcd}

\end{document}