### 2006-05-26 trunk/vcd/paper/Color/draft.Rnw
re-organized papers, added color draft
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@536 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{Z}
%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, echo=FALSE, results=hide}

\title{Choosing Color Palettes for Statistical Graphics}
\author{Achim Zeileis and Kurt Hornik\\Wirtschaftsuniversit\"at Wien}
\Plainauthor{Achim Zeileis, Kurt Hornik}
\Keywords{qualitative palette, sequential palette, diverging palette, HCL colors, HSV colors, perceptual-based color space}

\Abstract{
  Statistical graphs are often augmented by the use of color
  coding information contained in some variable. When this involves
  the shading of areas (and not only points or lines)---e.g., as in barplots, pie charts,
  mosaic displays or heatmaps---it is important that the colors are perceptual-based
  and do not introduce optical illusions or systematic bias. Here, we discuss
  how the perceptual based HCL color space can be used for deriving suitable
  color palettes for coding categorical data (qualitative palettes) and numerical
  variables (sequential and diverging palettes).
}

\begin{document}

<<preliminaries>>=
library("vcd")
source("functions.R")
generate_png <- TRUE
@

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[p]
\begin{center}
<<hsv-hcl>>=
if(generate_png) {
  files <- c("hcl-dcone", "hsv-cylinder")
  PNG <- function(x) paste(x, ".png", sep = "")

  png(file = PNG(files[1]), height = 480, width = 720)
  hcl.wheel(luminance = seq(5, 95, length = 6))
  dev.off()
  png(file = PNG(files[2]), height = 480, width = 720)
  hsv.wheel(value = seq(0, 1, length = 6), cone = FALSE)
  dev.off()
}
@
\includegraphics{hsv-cylinder}

\vspace*{2.5cm}

\includegraphics{hcl-dcone}
\caption{\label{fig:hcl-hsv} HSV and HCL space}
\end{center}
\end{figure}

\begin{figure}[p]
\begin{center}
<<hsv-hcl-gray>>=
if(generate_png) {
  for(i in files) {
    file.copy(PNG(i), PNG(paste(i, "-gray", sep = "")), overwrite = TRUE)
    system(paste("mogrify -type Grayscale", PNG(paste(i, "-gray", sep = ""))))
  }
}
@
\includegraphics{hsv-cylinder-gray}

\vspace*{2.5cm}

\includegraphics{hcl-dcone-gray}
\caption{\label{fig:hcl-hsv-gray} HSV and HCL space (in gray levels)}
\end{center}
\end{figure}

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[h]
\begin{center}
<<pal-q-construct>>=
if(generate_png) {
  png(file = "pal-q-construct.png", height = 480, width = 480)
  hcl.wheel(luminance = 70, do.label = FALSE, inner.circle = TRUE)
  rect(0.45, -0.05, 0.55, 0.05, col = hcl(0, 50, 70))
  rect(-cos(pi/3)/2-0.05, sin(pi/3)/2-0.05, -cos(pi/3)/2+0.05, sin(pi/3)/2+0.05, col = hcl(120, 50, 70))
  rect(-cos(pi/3)/2-0.05, -sin(pi/3)/2-0.05, -cos(pi/3)/2+0.05, -sin(pi/3)/2+0.05, col = hcl(240, 50, 70))
  dev.off()
}
@
\includegraphics{pal-q-construct}
\caption{\label{fig:pal-q-construct} Constructing qualitative palettes}
\end{center}
\end{figure}

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[h]
\begin{center}
<<pal-s-construct>>=
if(generate_png) {
  png(file = "pal-s-construct.png", height = 480, width = 400)
  hue.slice(260, grid.n = 301)
  lines(c(0, 100), c(90, 50), lty = 2)
  lines(c(0, 80), c(90, 30), lty = 2)
  lines(c(2.5, 2.5), c(90, 30), lty = 2)
  rect(0, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(47.5, 67.5, 52.5, 72.5, col = hcl(260, 50, 70))
  rect(95, 47.5, 100, 52.5, col = hcl(260, 100, 50))
  rect(37.5, 57.5, 42.5, 62.5, col = hcl(260, 40, 60))
  rect(77.5, 27.5, 82.5, 32.5, col = hcl(260, 80, 30))
  rect(0, 57.5, 5, 62.5, col = hcl(260, 0, 60))
  rect(0, 27.5, 5, 32.5, col = hcl(260, 0, 30))
  dev.off()
}
@
\includegraphics{pal-s-construct}
\caption{\label{fig:pal-s-construct} Constructing sequential palettes}
\end{center}
\end{figure}

\setkeys{Gin}{width=\textwidth}
\begin{figure}[h]
\begin{center}
<<pal-d-construct>>=
if(generate_png) {
  png(file = "pal-d-construct.png", height = 480, width = 480)
  hue260 <- hue.slice(260, grid.n = 301, plot = FALSE)
  hue360 <- hue.slice(360, grid.n = 301, plot = FALSE)
  mychroma <- as.numeric(colnames(hue260))
  luminance <- as.numeric(rownames(hue260))
  nc <- ncol(hue260)
  nl <- nrow(hue260)
  plot(0.5, 0.5, xlim = c(-100, 100), ylim = c(0, 100), type = "n", axes = FALSE,
       xlab = "", ylab = "", yaxs = "i", xaxs = "i", main = "")
  for(i in 1:(nc-1)) {
    rect(mychroma[i], luminance[-nl], mychroma[i] + 100/(nc-1), luminance[-1], border = hue260[,i+1], col = hue260[,i+1])
  }
  for(i in 1:(nc-1)) {
    rect(-mychroma[i], luminance[-nl], -(mychroma[i] + 100/(nc-1)), luminance[-1], border = hue360[,i+1], col = hue360[,i+1])
  }
  axis(1, labels = FALSE, at = c(-100, 0, 100))
  axis(2, at = c(0, 50, 70, 90, 100), labels = c(0, 50, 70, 90, 100))
  axis(3, at = -4:4*25, labels=c(4:0*25, 1:4*25))
  axis(4, at = c(0, 50, 70, 90, 100), labels = c(0, 50, 70, 90, 100))
  mtext(c("hue = 0", "hue = 260"), side = 3, at = c(-50, 50), line = 3, cex = 1.2)
  mtext("chroma", side = 3, at = 0, line = 2)
  mtext("luminance", side = 2, at = 50, line = 2)
  mtext("luminance", side = 4, at = 50, line = 2)
  lines(c(-100, 0, 100), c(50, 90, 50), lty = 2)
  rect(-100, 47.5, -90, 52.5, col = hcl(0, 100, 50))
  rect(-55, 67.5, -45, 72.5, col = hcl(0, 50, 70))
  rect(-5, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(45, 67.5, 55, 72.5, col = hcl(260, 50, 70))
  rect(90, 47.5, 100, 52.5, col = hcl(260, 100, 50))
  lines(c(-80, 0, 80), c(30, 90, 30), lty = 2)
  rect(-85, 27.5, -75, 32.5, col = hcl(0, 80, 30))
  rect(-45, 57.5, -35, 62.5, col = hcl(0, 40, 60))
  rect(-5, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(35, 57.5, 45, 62.5, col = hcl(260, 40, 60))
  rect(75, 27.5, 85, 32.5, col = hcl(260, 80, 30))
  dev.off()
}
@
\includegraphics{pal-d-construct}
\caption{\label{fig:pal-d-construct} Constructing diverging palettes}
\end{center}
\end{figure}

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[h]
\begin{center}
<<pal-q-examples, fig=TRUE, height=6, width=6>>=
par(mfrow = c(2, 2), mar = c(0, 0, 3, 0))
pal(rainbow_hcl(4, start = 30, end = 300), main = "dynamic (360 deg)")
pal(rainbow_hcl(4, start = 60, end = 240), main = "harmonic (180 deg)")
pal(rainbow_hcl(4, start = 270, end = 150), main = "cold (120 deg)")
pal(rainbow_hcl(4, start = 90, end = -30), main = "warm (120 deg)")
@
\caption{\label{fig:pal-q-examples} Examples for qualitative palettes}
\end{center}
\end{figure}

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[h]
\begin{center}
<<pal-s-examples, fig=TRUE, height=9, width=6>>=
par(mfrow = c(5, 1), mar = c(0, 0, 0, 0))
pal(sequential_hcl(12, c = 0, increase = 2.2))
pal(sequential_hcl(12, increase = 2.2))
pal(terrain_hcl(12, c = c(65, 0), l = c(45, 95), increase = c(1/3, 1.5)))
pal(heat_hcl(12, c = c(80, 30), l = c(30, 90), increase = c(1/5, 1.5)))
pal(rev(heat_hcl(12, h = c(0, -100), l = c(75, 40), c = c(40, 80), increase = 1)))
@
\caption{\label{fig:pal-s-examples} Examples for sequential palettes}
\end{center}
\end{figure}

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[h]
\begin{center}
<<pal-d-examples, fig=TRUE, height=8, width=6>>=
par(mfrow = c(4, 1), mar = c(0, 0, 0, 0))
pal(diverge_hcl(7))
pal(diverge_hcl(7, h = c(180, 330), c = 59, l = c(75, 95)))
pal(diverge_hcl(7, h = c(130, 43), c = 100, l = c(70, 90)))
pal(diverge_hcl(7, c = 100, l = c(50, 90), increase = 1))
@
\caption{\label{fig:pal-d-examples} Examples for diverging palettes}
\end{center}
\end{figure}

\end{document}