### 2006-05-27 trunk/vcd/paper/Color/draft.Rnw
first version of color palettes draft finished
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@537 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{Z}
%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, echo=FALSE, results=hide}

\title{Choosing Color Palettes for Statistical Graphics}
\author{Achim Zeileis and Kurt Hornik\\Wirtschaftsuniversit\"at Wien}
\Plainauthor{Achim Zeileis, Kurt Hornik}
\Keywords{qualitative palette, sequential palette, diverging palette, HCL colors, HSV colors, perceptually-based color space}

\Abstract{
  Statistical graphs are often augmented by the use of color
  coding information contained in some variable. When this involves
  the shading of areas (and not only points or lines)---e.g., as in barplots, pie charts,
  mosaic displays or heatmaps---it is important that the colors are perceptually-based
  and do not introduce optical illusions or systematic bias. Here, we discuss
  how the perceptually-based HCL color space can be used for deriving suitable
  color palettes for coding categorical data (qualitative palettes) and numerical
  variables (sequential and diverging palettes).
}

\begin{document}

\section{Introduction} \label{sec:intro}

<<preliminaries>>=
library("vcd")
source("functions.R")
generate_png <- TRUE
@

First, we motivate why HCL colors \citep{vcd:Ihaka:2003} are suitable for choosing color palettes
by contrasting them with the more commonly implemented HSV colors.

Then, we describe strategies how color palettes for categorical and numerical
data can be chosen in this space. Following \cite{vcd:Brewer:1999}, we distinguish
three types of palettes: qualitative, sequential and diverging.

All palettes described are 
available in \proglang{R} \citep{vcd:R:2006} in the package \pkg{vcd} \citep{vcd:Meyer+Zeileis+Hornik:2006}
using the HCL color implementation from \pkg{colorspace} \citep{vcd:Ihaka:2004}.
Technical documentation to the \proglang{R} implementations along with a large
collection of examples is available via \verb/help("rainbow_hcl")/ that provides 
more comparisons between existing \proglang{R} palettes (based on HSV colors) and
the HCL color palettes.

\section{Color spaces} \label{sec:spaces}

For choosing color palettes, it is imperative to have an understanding
how colors are perceived. For this it is helpful to have an idea how
human color vision evolved. It has been hypothesized that it developed in three
distinct stages: 1.~perception of \textit{light/dark} contrasts (monochrome only),
2.~\textit{yellow/blue} (usually associated with our notion of warm/cold colors),
3.~\textit{green/red} (helpful for assessing the ripeness of fruit). See 
\cite{vcd:Ihaka:2003} for more details and references.

Due to these three color axes, colors are typically described as
locations in a 3-dimensional spaces. However, human perception of color
does not correspond to the physiological axes above, but rather to polar
coordinates in the color plane (yellow/blue vs.~green/red) plus a third
light/dark axis. Thus, perceptually-based color spaces are defined by three dimensions
that try to capture
\begin{enumerate}
  \item \textbf{hue} (dominant wavelength)
  \item \textbf{chroma} (colorfulness)
  \item \textbf{luminance} (brightness, `amount of gray')
\end{enumerate}

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[p]
\begin{center}
<<hsv-hcl>>=
if(generate_png) {
  files <- c("hcl-dcone", "hsv-cylinder")
  PNG <- function(x) paste(x, ".png", sep = "")

  png(file = PNG(files[1]), height = 480, width = 720)
  hcl.wheel(luminance = seq(5, 95, length = 6))
  dev.off()
  png(file = PNG(files[2]), height = 480, width = 720)
  hsv.wheel(value = seq(0, 1, length = 6), cone = TRUE)
  dev.off()
}
@
\includegraphics{hsv-cylinder}

\vspace*{2.5cm}

\includegraphics{hcl-dcone}
\caption{\label{fig:hsv-hcl} HSV and HCL space}
\end{center}
\end{figure}

\begin{figure}[p]
\begin{center}
<<hsv-hcl-gray>>=
if(generate_png) {
  for(i in files) {
    file.copy(PNG(i), PNG(paste(i, "-gray", sep = "")), overwrite = TRUE)
    system(paste("mogrify -type Grayscale", PNG(paste(i, "-gray", sep = ""))))
  }
}
@
\includegraphics{hsv-cylinder-gray}

\vspace*{2.5cm}

\includegraphics{hcl-dcone-gray}
\caption{\label{fig:hsv-hcl-gray} HSV and HCL space (in gray levels)}
\end{center}
\end{figure}

A popular implementation of such a color space, available in many graphics and
statistics software packages, are \textit{HSV} (hue, saturation, value) colors.
They are a simple transformation of \textit{RGB} (red, green, blue) colors
and are defined by a triplet $(H, S, V)$ with $H \in [0, 360]$, $S, V \in [0, 100]$.
HSV space has the shape of a single regular cone (often inflated to a regular
cylinder). Vertical sections through this space are shown in the upper panel
of Figure~\ref{fig:hsv-hcl}, depicting hue and saturation given value.
Although simple to specify and easily available in many computing environments,
HSV colors have a fundamental drawback: its three dimensions map to the three
dimensions of human color perception very poorly. The three dimensions are
confounded which is most easily seen when converting the vertical sections to 
gray scale images in Figure~\ref{fig:hsv-hcl-gray}. Clearly, the brightness of
colors is not uniform of hues and saturations given the value---therefore,
HSV colors are often not considered to be perceptually based.

To overcome these drawbacks, various color spaces have been suggested
that properly map to the perception dimensions, the most prominent of which are
the CIELUV and CIELAB spaces developed by the \cite{vcd:CIE:2004}. \cite{vcd:Ihaka:2003}
argues that CIELUV colors are typically preferred for use with emissive technologies
such as computer screens which makes them an obvious candidate for implementation
in statistical software packages. By taking polar coordinates in the UV plane of
CIELUV, \textit{HCL} (hue, chroma, luminance) colors are obtained, defined by
a triplet $(H, C, L)$ with $H \in [0, 360]$ and $C, L \in [0, 100]$. HCL space
has the shape of a distorted double cone: the admissable chroma and luminance values
depend on the hue chosen. The lower panel of Figure~\ref{fig:hsv-hcl} shows
vertical sections through this space: each of the resulting hue/chroma planes (given
luminance) is now properly balanced towards the same gray (going from black to white
with increasing luminance) which becomes obvious when converting the colors to
a gray scale is in Figure~\ref{fig:hsv-hcl-gray}. This balancing of HCL colors
gives us the opportunity to conveniently choose color palettes which code categorical
and/or numerical information by translating it to the three perceptual dimensions.

\section{Qualitative palettes} \label{sec:qualitative}

Qualitative palettes are a set of colors for depicting different categories,
i.e., for coding a categorical variable. Usually, these should give the same
perceptual weight to each category so that no group is perceived to be larger/more
important than any other one. Typical applications of qualitative palettes in
statistics would be bar plots, pie charts or highlighted spine plots.

\cite{vcd:Ihaka:2003} describes a simple strategy for choosing such palettes: chroma
and luminance are kept fixed and only the hue is varied for obtaining different colors
which are consequently all balanced towards the same gray. If colors from the full
color wheel (i.e., $H \in [0, 360]$) should be used, not all combinations of chroma
and luminance are feasible. Figure~\ref{fig:pal-q-construct} depicts how three colors
are chosen, given $C = 50$ and $L = 70$.

Various strategies for choosing the hues in a certain palette are conceivable.
A simple and intuitive one is often to use colors as metaphors for categories
(e.g., for political parties). Figure~\ref{fig:pal-q-examples} shows a few
further examples for generating qualitative sets of colors $(H, 50, 70)$.
In the upper left panel colors from the full 360 degrees are used ($H = 30, 120, 210, 300$)
creating a `dynamic' set of colors. The upper right panel shows a `harmonic'
set with $H = 60, 120, 180, 240$. Warm ($H = 270, 230, 190, 150$) and cold 
($H = 90, 50, 10, 330$) are shown in the lower left and right panel, respectively.

In \pkg{vcd}, these palettes are available in the function
\begin{Soutput}
rainbow_hcl(n, c = 50, l = 70, start = 0, end = 360*(n-1)/n, ...)
\end{Soutput}

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-q-construct>>=
if(generate_png) {
  png(file = "pal-q-construct.png", height = 480, width = 480)
  hcl.wheel(luminance = 70, do.label = FALSE, inner.circle = TRUE)
  rect(0.45, -0.05, 0.55, 0.05, col = hcl(0, 50, 70))
  rect(-cos(pi/3)/2-0.05, sin(pi/3)/2-0.05, -cos(pi/3)/2+0.05, sin(pi/3)/2+0.05, col = hcl(120, 50, 70))
  rect(-cos(pi/3)/2-0.05, -sin(pi/3)/2-0.05, -cos(pi/3)/2+0.05, -sin(pi/3)/2+0.05, col = hcl(240, 50, 70))
  dev.off()
}
@
\includegraphics{pal-q-construct}
\vspace*{-2cm}
\caption{\label{fig:pal-q-construct} Constructing qualitative palettes}
\end{center}
\end{figure}

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-q-examples, fig=TRUE, height=4, width=6>>=
par(mfrow = c(2, 2), mar = c(0, 0, 3, 0))
pal(rainbow_hcl(4, start = 30, end = 300), main = "dynamic (30 - 300)")
pal(rainbow_hcl(4, start = 60, end = 240), main = "harmonic (60 - 240)")
pal(rainbow_hcl(4, start = 270, end = 150), main = "cold (270 - 150)")
pal(rainbow_hcl(4, start = 90, end = -30), main = "warm (90 - -30)")
@
\caption{\label{fig:pal-q-examples} Examples for qualitative palettes}
\end{center}
\end{figure}

\section{Sequential palettes} \label{sec:sequential}

Sequential palettes are used for coding numerical information that simply ranges
in a certain interval where low values are considered to be uninteresting and
high values are interesting. Without loss of generality, we assume that we want
to visualize an intensity or interestingness $i \in [0, 1]$. A typical application
in statistics are heatmaps.

The simplest solution to this task is to employ light/dark contrasts, i.e., employ the
oldest and easiest axis of the perceptual system. The interestingness is thus coded
by an increasing amount of gray (i.e., decreasing luminance) 
  \[ (H, 0, 90 - i \cdot 60), \]
where the hue $H$ used does not matter, chroma is set to 0 (i.e., no color), and 
luminance ranges in $[30, 90]$ avoiding the extreme colors white ($L = 100$) and 
black ($L = 0$). Instead of going linearly from the light to the dark gray, luminance
could also be increased nonlinearly, e.g., by some function $i^p$ where the power $p$
controls wether intensity/luminance is increase quickly or not.

Alternatively, the intensity $i$ could additionally be code by colorfulness (chroma), e.g.,
  \[ (H, 0 + i^p \cdot C_{\max}, L_{\max} - i^p \cdot (L_{\max} - L_{\min})). \]
This strategy is depicted in Figure~\ref{fig:pal-s-construct} for a blue hue $H = 260$
and different combinations of maximal chroma (0, 80 and 100, respecitvely) and
minimal luminance (30, 30 and 50, respectively). The first two combinations are also
shown in the first two rows of Figure~\ref{fig:pal-s-examples}.

\setkeys{Gin}{width=.7\textwidth}
\begin{figure}[b!]
\begin{center}
<<pal-s-construct>>=
if(generate_png) {
  png(file = "pal-s-construct.png", height = 420, width = 420)
  hue.slice(260, grid.n = 301)
  lines(c(0, 100), c(90, 50), lty = 2)
  lines(c(0, 80), c(90, 30), lty = 2)
  lines(c(2.5, 2.5), c(90, 30), lty = 2)
  rect(0, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(47.5, 67.5, 52.5, 72.5, col = hcl(260, 50, 70))
  rect(95, 47.5, 100, 52.5, col = hcl(260, 100, 50))
  rect(37.5, 57.5, 42.5, 62.5, col = hcl(260, 40, 60))
  rect(77.5, 27.5, 82.5, 32.5, col = hcl(260, 80, 30))
  rect(0, 57.5, 5, 62.5, col = hcl(260, 0, 60))
  rect(0, 27.5, 5, 32.5, col = hcl(260, 0, 30))
  dev.off()
}
@
\includegraphics{pal-s-construct}
\caption{\label{fig:pal-s-construct} Constructing sequential palettes}
\end{center}
\end{figure}

In \pkg{vcd}, this strategy is implemented in the function
\begin{Soutput}
sequential_hcl(n, h = 260, c = c(80, 0), l = c(30, 90), power = 1.5, ...) 
\end{Soutput}

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-s-examples, fig=TRUE, height=9, width=6>>=
par(mfrow = c(5, 1), mar = c(0, 0, 0, 0))
pal(sequential_hcl(12, c = 0, power = 2.2))
pal(sequential_hcl(12, power = 2.2))
pal(terrain_hcl(12, c = c(65, 0), l = c(45, 95), power = c(1/3, 1.5)))
pal(heat_hcl(12, c = c(80, 30), l = c(30, 90), power = c(1/5, 1.5)))
pal(rev(heat_hcl(12, h = c(0, -100), l = c(75, 40), c = c(40, 80), power = 1)))
@
\caption{\label{fig:pal-s-examples} Examples for sequential palettes}
\end{center}
\end{figure}

To increase the contrast between the colors in the palette even further, the ideas from
the previous sequential palettes can also be combined with qualitative palettes by
simultaneously varying the hue as well:
  \[ (H_2 - i \cdot (H_1 - H_2), 0 + i^{p_1} \cdot C_{\max}, L_{\max} - i^{p_2} \cdot (L_{\max} - L_{\min})). \]
A typical application would be heat colors that increase from a light yellow to a full
red. To make the change in hue visible, the chroma needs to increase rather quickly for low
values of $i$ and then only slowly for higher values of $i$. This can be achieved by
choosing a power $p_1 < 1$.

In \proglang{R}, these are available in the function
\begin{Soutput}
heat_hcl(n, h = c(0, 90), c = c(100, 30), l = c(50, 90), power = c(1/5, 1), ...)
\end{Soutput}
with which the lower three rows in Figure~\ref{fig:pal-s-examples} are produced.

\section{Diverging palettes} \label{sec:diverging}

Diverging palettes are also used for coding numerical information ranging in a certain
interval---however, this interval includes a neutral value. Examples for this include
residuals (with the neutral value 0) or binary classification probabilities (with neutral
value 0.5) that could be visualized in mosaic plots \citep{vcd:Zeileis+Meyer+Hornik:2005}
or classification maps. Without loss of generality, we assume that we want
to visualize an intensity or interestingness $i \in [-1, 1]$.

\setkeys{Gin}{width=\textwidth}
\begin{figure}[b!]
\begin{center}
<<pal-d-construct>>=
if(generate_png) {
  png(file = "pal-d-construct.png", height = 480, width = 540)
  hue260 <- hue.slice(260, grid.n = 301, plot = FALSE)
  hue360 <- hue.slice(360, grid.n = 301, plot = FALSE)
  mychroma <- as.numeric(colnames(hue260))
  luminance <- as.numeric(rownames(hue260))
  nc <- ncol(hue260)
  nl <- nrow(hue260)
  plot(0.5, 0.5, xlim = c(-100, 100), ylim = c(0, 100), type = "n", axes = FALSE,
       xlab = "", ylab = "", yaxs = "i", xaxs = "i", main = "")
  for(i in 1:(nc-1)) {
    rect(mychroma[i], luminance[-nl], mychroma[i] + 100/(nc-1), luminance[-1], border = hue260[,i+1], col = hue260[,i+1])
  }
  for(i in 1:(nc-1)) {
    rect(-mychroma[i], luminance[-nl], -(mychroma[i] + 100/(nc-1)), luminance[-1], border = hue360[,i+1], col = hue360[,i+1])
  }
  axis(1, labels = FALSE, at = c(-100, 0, 100))
  axis(2, at = c(0, 30, 50, 60, 70, 90, 100), labels = c(0, 30, 50, 60, 70, 90, 100))
  axis(3, at = -4:4*25, labels=c(4:0*25, 1:4*25))
  axis(4, at = c(0, 30, 50, 60, 70, 90, 100), labels = c(0, 30, 50, 60, 70, 90, 100))
  mtext(c("hue = 0", "hue = 260"), side = 3, at = c(-50, 50), line = 3, cex = 1.2)
  mtext("chroma", side = 3, at = 0, line = 2)
  mtext("luminance", side = 2, at = 50, line = 2)
  mtext("luminance", side = 4, at = 50, line = 2)
  lines(c(-100, 0, 100), c(50, 90, 50), lty = 2)
  rect(-100, 47.5, -90, 52.5, col = hcl(0, 100, 50))
  rect(-55, 67.5, -45, 72.5, col = hcl(0, 50, 70))
  rect(-5, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(45, 67.5, 55, 72.5, col = hcl(260, 50, 70))
  rect(90, 47.5, 100, 52.5, col = hcl(260, 100, 50))
  lines(c(-80, 0, 80), c(30, 90, 30), lty = 2)
  rect(-85, 27.5, -75, 32.5, col = hcl(0, 80, 30))
  rect(-45, 57.5, -35, 62.5, col = hcl(0, 40, 60))
  rect(-5, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(35, 57.5, 45, 62.5, col = hcl(260, 40, 60))
  rect(75, 27.5, 85, 32.5, col = hcl(260, 80, 30))
  dev.off()
}
@
\includegraphics{pal-d-construct}
\caption{\label{fig:pal-d-construct} Constructing diverging palettes}
\end{center}
\end{figure}

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-d-examples, fig=TRUE, height=8, width=6>>=
par(mfrow = c(4, 1), mar = c(0, 0, 0, 0))
pal(diverge_hcl(7))
pal(diverge_hcl(7, c = 100, l = c(50, 90), power = 1))
pal(diverge_hcl(7, h = c(180, 330), c = 59, l = c(75, 95)))
pal(diverge_hcl(7, h = c(130, 43), c = 100, l = c(70, 90)))
@
\caption{\label{fig:pal-d-examples} Examples for diverging palettes}
\end{center}
\end{figure}

Given sequential palettes, deriving diverging palettes is easy: two different hues are
chosen for adding color to the same amount of `gray' at a given intensity $|i|$.
Figure~\ref{fig:pal-d-construct} shows the chroma/luminance plane back to back for the
hues $H = 0$ and 260. These particular hues were chosen because they have rather similar
chroma/luminance planes, allowing for different combinations of maximal chroma and luminance.

In \pkg{vcd}, the function
\begin{Soutput}
diverge_hcl(n, h = c(260, 0), c = 80, l = c(30, 90), power = 1.5, ...) 
\end{Soutput}
is provided.

Figure~\ref{fig:pal-d-examples} shows various examples of conceivable combinations of
hue, chroma and luminance. The first palette uses a broader range on the luminance axis
whereas the others use larger ranges on the chroma axis.

\section{Open questions}

From our experience, the paths through HCL space described in the previous sections
in principle work very well and can be used to effectively code qualitative and 
quantitative information. However, we are not sure how we can properly define trade
offs between increasing the range with respect to one dimension and decreasing the
range with respect to another. In other words, how much difference in chroma do I need
to make up for a certain difference in luminance etc.? A few more precise questions
are formulated below.

\begin{itemize}
  \item In sequential/diverging palettes, there is a trade-off between high chroma and high luminance.
        How should this be chosen in practice? Our impression is that when a small set of colors
        (such as 3 or 4) are used, large differnces in chroma work well and large differences
        in luminance are not necessary. However, when a larger set of colors is used (e.g., for 
        heatmaps where extreme values should be identifiable) it is much more important to have a big
        difference in luminance.

  \item How should the intensity $|i|$ be increased from 0 to 1?
        Our experience is that for a small set of colors linear increase is sufficient ($p = 1$)
        whereas in heatmaps where only very extreme regions are interesting a $p > 1$
        should be used.

  \item For diverging palettes, how should pairs of colors be chosen? The hues $H = 0$
        and $H = 260$ where chosen because they are on opposite sides of the color wheel
	but have a very similar chroma/luminance plane.
\end{itemize}

\bibliography{vcd}

\end{document}