### 2006-05-28 trunk/vcd/paper/Color/draft.Rnw
draft version of color paper for Ross & Paul
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@539 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{Z}
%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, echo=FALSE, results=hide}

\title{Choosing Color Palettes for Statistical Graphics}
\author{Achim Zeileis and Kurt Hornik\\Wirtschaftsuniversit\"at Wien}
\Plainauthor{Achim Zeileis, Kurt Hornik}
\Keywords{qualitative palette, sequential palette, diverging palette, HCL colors, HSV colors, perceptually-based color space}

\Abstract{
  Statistical graphics are often augmented by the use of color
  coding information contained in some variable. When this involves
  the shading of areas (and not only points or lines)---e.g., as in barplots, pie charts,
  mosaic displays or heatmaps---it is important that the colors are perceptually based
  and do not introduce optical illusions or systematic bias. Here, we discuss
  how the perceptually-based HCL color space can be used for deriving suitable
  color palettes for coding categorical data (qualitative palettes) and numerical
  variables (sequential and diverging palettes).
}

\begin{document}

\section{Introduction} \label{sec:intro}

<<preliminaries>>=
library("vcd")
library("e1071")
library("mvtnorm")
library("ellipse")
library("kernlab")
library("KernSmooth")

source("functions.R")
load("Bundestagswahl2005.rda")

generate_png <- FALSE
@

First, we motivate why HCL colors \citep{vcd:Ihaka:2003} are suitable for choosing color palettes
by contrasting them with the more commonly implemented HSV colors.

Then, we describe strategies how color palettes for categorical and numerical
data can be chosen in this space. Following \cite{vcd:Brewer:1999}, we distinguish
three types of palettes: qualitative, sequential and diverging.
All palettes described are 
available in \proglang{R} \citep{vcd:R:2006} in the package \pkg{vcd} \citep{vcd:Meyer+Zeileis+Hornik:2006}
using the HCL color implementation from \pkg{colorspace} \citep{vcd:Ihaka:2004}.
Technical documentation to the \proglang{R} implementations along with a large
collection of examples is available via \verb/help("rainbow_hcl")/ that provides 
more comparisons between existing \proglang{R} palettes (based on HSV colors) and
the HCL color palettes.

\section{Color spaces} \label{sec:spaces}

For choosing color palettes, it is imperative to have an understanding
how colors are perceived. For this it is helpful to have an idea how
human color vision evolved. It has been hypothesized that it developed in three
distinct stages: 1.~perception of \textit{light/dark} contrasts (monochrome only),
2.~\textit{yellow/blue} contrasts (usually associated with our notion of warm/cold colors),
3.~\textit{green/red} contrasts (helpful for assessing the ripeness of fruit). See 
\cite{vcd:Ihaka:2003} for more details and references.

Due to these three color axes, colors are typically described as
locations in a 3-dimensional spaces. However, human perception of color
does not correspond to the physiological axes above, but rather to polar
coordinates in the color plane (yellow/blue vs.~green/red) plus a third
light/dark axis. Thus, perceptually-based color spaces are defined by three dimensions
that try to capture
\begin{enumerate}
  \item \textbf{hue} (dominant wavelength)
  \item \textbf{chroma} (colorfulness)
  \item \textbf{luminance} (brightness, `amount of gray')
\end{enumerate}

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[p]
\begin{center}
<<hsv-hcl>>=
if(generate_png) {
  files <- c("hcl-dcone", "hsv-cylinder")
  PNG <- function(x) paste(x, ".png", sep = "")

  png(file = PNG(files[1]), height = 480, width = 720)
  hcl.wheel(luminance = seq(5, 95, length = 6))
  dev.off()
  png(file = PNG(files[2]), height = 480, width = 720)
  hsv.wheel(value = seq(0, 1, length = 6), cone = TRUE)
  dev.off()
}
@
\includegraphics{hsv-cylinder}

\vspace*{2.5cm}

\includegraphics{hcl-dcone}
\caption{\label{fig:hsv-hcl} HSV and HCL space}
\end{center}
\end{figure}

\begin{figure}[p]
\begin{center}
<<hsv-hcl-gray>>=
if(generate_png) {
  for(i in files) {
    file.copy(PNG(i), PNG(paste(i, "-gray", sep = "")), overwrite = TRUE)
    system(paste("mogrify -type Grayscale", PNG(paste(i, "-gray", sep = ""))))
  }
}
@
\includegraphics{hsv-cylinder-gray}

\vspace*{2.5cm}

\includegraphics{hcl-dcone-gray}
\caption{\label{fig:hsv-hcl-gray} HSV and HCL space (in gray levels)}
\end{center}
\end{figure}

A popular implementation of such a color space, available in many graphics and
statistics software packages, are \textit{HSV} (hue, saturation, value) colors.
They are a simple transformation of \textit{RGB} (red, green, blue) colors
and are defined by a triplet $(H, S, V)$ with $H \in [0, 360]$ and $S, V \in [0, 100]$.
HSV space has the shape of a single regular cone (often inflated to a regular
cylinder). Vertical sections through this space are shown in the upper panel
of Figure~\ref{fig:hsv-hcl}, depicting hue and saturation given value.
Although simple to specify and easily available in many computing environments,
HSV colors have a fundamental drawback: its three dimensions map to the three
dimensions of human color perception very poorly. The three dimensions are
confounded which is most easily seen when converting the vertical sections to 
gray scale images in Figure~\ref{fig:hsv-hcl-gray}. Clearly, the brightness of
colors is not uniform over hues and saturations (given value)---therefore,
HSV colors are often not considered to be perceptually based.

To overcome these drawbacks, various color spaces have been suggested
that properly map to the perception dimensions, the most prominent of which are
the CIELUV and CIELAB spaces developed by the \cite{vcd:CIE:2004}. \cite{vcd:Ihaka:2003}
argues that CIELUV colors are typically preferred for use with emissive technologies
such as computer screens which makes them an obvious candidate for implementation
in statistical software packages. By taking polar coordinates in the UV plane of
CIELUV, \textit{HCL} (hue, chroma, luminance) colors are obtained, defined by
a triplet $(H, C, L)$ with $H \in [0, 360]$ and $C, L \in [0, 100]$. HCL space
has the shape of a distorted double cone: the admissable chroma and luminance values
depend on the hue chosen. The lower panel of Figure~\ref{fig:hsv-hcl} shows
vertical sections through this space: each of the resulting hue/chroma planes (given
luminance) is now properly balanced towards the same gray (going from black to white
with increasing luminance) which becomes obvious when converting the colors to
a gray scale is in Figure~\ref{fig:hsv-hcl-gray}. This balancing of HCL colors
gives us the opportunity to conveniently choose color palettes which code categorical
and/or numerical information by translating it to the three perceptual dimensions.
However, some care is required for dealing with the irregular shape of the HCL space
which will be addressed in the following sections.

\section{Qualitative palettes} \label{sec:qualitative}

Qualitative palettes are sets of colors for depicting different categories,
i.e., for coding a categorical variable. Usually, these should give the same
perceptual weight to each category so that no group is perceived to be larger/more
important than any other one. Typical applications of qualitative palettes in
statistics would be bar plots, pie charts (see Figure~\ref{fig:seats}) 
or highlighted mosaic displays (see Figure~\ref{fig:votes}).

\cite{vcd:Ihaka:2003} describes a simple strategy for choosing such palettes: chroma
and luminance are kept fixed and only the hue is varied for obtaining different colors
which are consequently all balanced towards the same gray. If colors from the full
color wheel (i.e., $H \in [0, 360]$) should be used, not all combinations of chroma
and luminance are feasible. Figure~\ref{fig:pal-q-construct} depicts how three colors
are chosen, given $C = 50$ and $L = 70$.

Various strategies for choosing the hues in a certain palette are conceivable.
A simple and intuitive one is to use colors as metaphors for categories
(e.g., for political parties). Figure~\ref{fig:pal-q-examples} shows a few
further examples for generating qualitative sets of colors $(H, 50, 70)$.
In the upper left panel colors from the full spectrum are used ($H = 30, 120, 210, 300$)
creating a `dynamic' set of colors. The upper right panel shows a `harmonic'
set with $H = 60, 120, 180, 240$. Warm colors (from the blue/green part of the spectrum: 
$H = 270, 230, 190, 150$) and cold colors (from the yellow/red part of the spectrum:
$H = 90, 50, 10, 330$) are shown in the lower left and right panel, respectively.

In \pkg{vcd}, these palettes are available in the function
\begin{Soutput}
rainbow_hcl(n, c = 50, l = 70, start = 0, end = 360*(n-1)/n, ...)
\end{Soutput}

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-q-construct>>=
if(generate_png) {
  png(file = "pal-q-construct.png", height = 480, width = 480)
  hcl.wheel(luminance = 70, do.label = FALSE, inner.circle = TRUE)
  rect(0.45, -0.05, 0.55, 0.05, col = hcl(0, 50, 70))
  rect(-cos(pi/3)/2-0.05, sin(pi/3)/2-0.05, -cos(pi/3)/2+0.05, sin(pi/3)/2+0.05, col = hcl(120, 50, 70))
  rect(-cos(pi/3)/2-0.05, -sin(pi/3)/2-0.05, -cos(pi/3)/2+0.05, -sin(pi/3)/2+0.05, col = hcl(240, 50, 70))
  dev.off()
}
@
\includegraphics{pal-q-construct}
\vspace*{-2cm}
\caption{\label{fig:pal-q-construct} Constructing qualitative palettes}
\end{center}
\end{figure}

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-q-examples, fig=TRUE, height=4, width=6>>=
par(mfrow = c(2, 2), mar = c(0, 0, 3, 0))
pal(rainbow_hcl(4, start = 30, end = 300), main = "dynamic (30 - 300)")
pal(rainbow_hcl(4, start = 60, end = 240), main = "harmonic (60 - 240)")
pal(rainbow_hcl(4, start = 270, end = 150), main = "cold (270 - 150)")
pal(rainbow_hcl(4, start = 90, end = -30), main = "warm (90 - -30)")
@
\caption{\label{fig:pal-q-examples} Examples for qualitative palettes}
\end{center}
\end{figure}

\section{Sequential palettes} \label{sec:sequential}

Sequential palettes are used for coding numerical information that simply ranges
in a certain interval where low values are considered to be uninteresting and
high values are interesting. Without loss of generality, we assume that we want
to visualize an intensity or interestingness $i \in [0, 1]$. A typical application
in statistics are heatmaps (see Figure~\ref{fig:bkde}).

The simplest solution to this task is to employ light/dark contrasts, i.e., employ the
oldest and simples perceptual axis. The interestingness is thus coded
by an increasing amount of gray (i.e., decreasing luminance) 
  \[ (H, 0, 90 - i \cdot 60), \]
where the hue $H$ used does not matter, chroma is set to 0 (i.e., no color), and 
luminance ranges in $[30, 90]$ avoiding the extreme colors white ($L = 100$) and 
black ($L = 0$). Instead of going linearly from light to dark gray, luminance
could also be increased nonlinearly, e.g., by some function $f(i)$ that
controls wether intensity/luminance is increased quickly or not. We found $f(i) = i^p$ 
to be a convenient transformation where the power $p$ can be varied to achieve
different degrees of non-linearity.

Furthermore, the intensity $i$ could additionally be coded by colorfulness (chroma), e.g.,
  \[ (H, 0 + i^p \cdot C_{\max}, L_{\max} - i^p \cdot (L_{\max} - L_{\min})). \]
This strategy is depicted in the left panel of Figure~\ref{fig:pal-s-construct} for a blue
hue $H = 260$ and different combinations of maximal chroma (0, 80 and 100, respecitvely) and
minimal luminance (30, 30 and 50, respectively). The first two combinations are also
shown in the first two rows of Figure~\ref{fig:pal-s-examples}. The right panel of
Figure~\ref{fig:pal-s-construct} shows that the exact same strategy is not possible 
for the green hue $H = 120$. While the gray colors without chroma can be chosen in 
the same way, there is a stronger trade-off between using dark colors (with low luminance)
and colorful colors (with high chroma). Hence, the second path from light gray to 
full green ends at a much lighter color with $L = 75$.

\setkeys{Gin}{width=\textwidth}
\begin{figure}[b!]
\begin{center}
<<pal-s-construct>>=
if(generate_png) {
  png(file = "pal-s-construct.png", height = 420, width = 840)
  par(mfrow = c(1, 2))
  hue.slice(260, grid.n = 301)
  lines(c(0, 100), c(90, 50), lty = 2)
  lines(c(0, 80), c(90, 30), lty = 2)
  lines(c(2.5, 2.5), c(90, 30), lty = 2)
  rect(0, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(47.5, 67.5, 52.5, 72.5, col = hcl(260, 50, 70))
  rect(95, 47.5, 100, 52.5, col = hcl(260, 100, 50))
  rect(37.5, 57.5, 42.5, 62.5, col = hcl(260, 40, 60))
  rect(77.5, 27.5, 82.5, 32.5, col = hcl(260, 80, 30))
  rect(0, 57.5, 5, 62.5, col = hcl(260, 0, 60))
  rect(0, 27.5, 5, 32.5, col = hcl(260, 0, 30))
  hue.slice(120, grid.n = 301)
  lines(c(0, 100), c(90, 75), lty = 2)
  lines(c(2.5, 2.5), c(90, 30), lty = 2)
  rect(0, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(47.5, 80, 52.5, 85, col = hcl(120, 50, 82.5))
  rect(95, 72.5, 100, 77.5, col = hcl(120, 100, 75))
  rect(0, 57.5, 5, 62.5, col = hcl(260, 0, 60))
  rect(0, 27.5, 5, 32.5, col = hcl(260, 0, 30))
  dev.off()
}
@
\includegraphics{pal-s-construct}
\caption{\label{fig:pal-s-construct} Constructing sequential palettes}
\end{center}
\end{figure}

In \pkg{vcd}, this strategy is implemented in the function
\begin{Soutput}
sequential_hcl(n, h = 260, c = c(80, 0), l = c(30, 90), power = 1.5, ...) 
\end{Soutput}

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-s-examples, fig=TRUE, height=9, width=6>>=
par(mfrow = c(5, 1), mar = c(0, 0, 0, 0))
pal(sequential_hcl(12, c = 0, power = 2.2))
pal(sequential_hcl(12, power = 2.2))
pal(terrain_hcl(12, c = c(65, 0), l = c(45, 95), power = c(1/3, 1.5)))
pal(heat_hcl(12, c = c(80, 30), l = c(30, 90), power = c(1/5, 1.5)))
pal(rev(heat_hcl(12, h = c(0, -100), l = c(75, 40), c = c(40, 80), power = 1)))
@
\caption{\label{fig:pal-s-examples} Examples for sequential palettes}
\end{center}
\end{figure}

To increase the contrast between the colors in the palette even further, the ideas from
the previous sequential palettes can also be combined with qualitative palettes by
simultaneously varying the hue as well:
  \[ (H_2 - i \cdot (H_1 - H_2), C_{\max} - i^{p_1} \cdot (C_{\max} - C_{\min}), L_{\max} - i^{p_2} \cdot (L_{\max} - L_{\min})). \]
A typical application would be heat colors that increase from a light yellow (e.g., $(90, 30, 90)$)
to a full red (e.g., $(0, 100, 50)$). To make the change in hue visible, the chroma needs to
increase rather quickly for low values of $i$ and then only slowly for higher values of $i$.
This can be achieved by choosing a power $p_1 < 1$.

In \proglang{R}, these are available in the function
\begin{Soutput}
heat_hcl(n, h = c(0, 90), c = c(100, 30), l = c(50, 90), power = c(1/5, 1), ...)
\end{Soutput}
with which the lower three rows in Figure~\ref{fig:pal-s-examples} are produced.

\section{Diverging palettes} \label{sec:diverging}

Diverging palettes are also used for coding numerical information ranging in a certain
interval---however, this interval includes a neutral value. Examples for this include
residuals or correlations (both with the neutral value 0) or binary classification
probabilities (with neutral value 0.5) that could be visualized in mosaic plots (see Figure~\ref{fig:arthritis})
or classification maps (see Figure~\ref{fig:class}). Without loss of generality, we assume
that we want to visualize an intensity or interestingness $i \in [-1, 1]$.

\setkeys{Gin}{width=\textwidth}
\begin{figure}[b!]
\begin{center}
<<pal-d-construct>>=
if(generate_png) {
  png(file = "pal-d-construct.png", height = 480, width = 540)
  hue260 <- hue.slice(260, grid.n = 301, plot = FALSE)
  hue360 <- hue.slice(360, grid.n = 301, plot = FALSE)
  mychroma <- as.numeric(colnames(hue260))
  luminance <- as.numeric(rownames(hue260))
  nc <- ncol(hue260)
  nl <- nrow(hue260)
  plot(0.5, 0.5, xlim = c(-100, 100), ylim = c(0, 100), type = "n", axes = FALSE,
       xlab = "", ylab = "", yaxs = "i", xaxs = "i", main = "")
  for(i in 1:(nc-1)) {
    rect(mychroma[i], luminance[-nl], mychroma[i] + 100/(nc-1), luminance[-1], border = hue260[,i+1], col = hue260[,i+1])
  }
  for(i in 1:(nc-1)) {
    rect(-mychroma[i], luminance[-nl], -(mychroma[i] + 100/(nc-1)), luminance[-1], border = hue360[,i+1], col = hue360[,i+1])
  }
  axis(1, labels = FALSE, at = c(-100, 0, 100))
  axis(2, at = c(0, 30, 50, 60, 70, 90, 100), labels = c(0, 30, 50, 60, 70, 90, 100))
  axis(3, at = -4:4*25, labels=c(4:0*25, 1:4*25))
  axis(4, at = c(0, 30, 50, 60, 70, 90, 100), labels = c(0, 30, 50, 60, 70, 90, 100))
  mtext(c("hue = 0", "hue = 260"), side = 3, at = c(-50, 50), line = 3, cex = 1.2)
  mtext("chroma", side = 3, at = 0, line = 2)
  mtext("luminance", side = 2, at = 50, line = 2)
  mtext("luminance", side = 4, at = 50, line = 2)
  lines(c(-100, 0, 100), c(50, 90, 50), lty = 2)
  rect(-100, 47.5, -90, 52.5, col = hcl(0, 100, 50))
  rect(-55, 67.5, -45, 72.5, col = hcl(0, 50, 70))
  rect(-5, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(45, 67.5, 55, 72.5, col = hcl(260, 50, 70))
  rect(90, 47.5, 100, 52.5, col = hcl(260, 100, 50))
  lines(c(-80, 0, 80), c(30, 90, 30), lty = 2)
  rect(-85, 27.5, -75, 32.5, col = hcl(0, 80, 30))
  rect(-45, 57.5, -35, 62.5, col = hcl(0, 40, 60))
  rect(-5, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(35, 57.5, 45, 62.5, col = hcl(260, 40, 60))
  rect(75, 27.5, 85, 32.5, col = hcl(260, 80, 30))
  dev.off()
}
@
\includegraphics{pal-d-construct}
\caption{\label{fig:pal-d-construct} Constructing diverging palettes}
\end{center}
\end{figure}

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-d-examples, fig=TRUE, height=8, width=6>>=
par(mfrow = c(4, 1), mar = c(0, 0, 0, 0))
pal(diverge_hcl(7))
pal(diverge_hcl(7, c = 100, l = c(50, 90), power = 1))
pal(diverge_hcl(7, h = c(180, 330), c = 59, l = c(75, 95)))
pal(diverge_hcl(7, h = c(130, 43), c = 100, l = c(70, 90)))
@
\caption{\label{fig:pal-d-examples} Examples for diverging palettes}
\end{center}
\end{figure}

Given sequential palettes, deriving diverging palettes is easy: two different hues are
chosen for adding color to the same amount of `gray' at a given intensity $|i|$.
Figure~\ref{fig:pal-d-construct} shows the chroma/luminance plane back to back for the
hues $H = 0$ and 260 with two different paths---with slightly different emphasis on
luminance or chroma contrasts---from a full red over a neutral grey to a full blue.
These particular hues were chosen because they have rather similar
chroma/luminance planes, allowing for many different combinations of maximal chroma and luminance.
As Figure~\ref{fig:pal-s-construct} illustrates, a bit of care is needed to choose two colors
that are sufficiently similar in their chroma/luminance plane to allow for the same
luminance and/or chroma contrasts.

In \pkg{vcd}, the following function is provided:
\begin{Soutput}
diverge_hcl(n, h = c(260, 0), c = 80, l = c(30, 90), power = 1.5, ...) 
\end{Soutput}

Figure~\ref{fig:pal-d-examples} shows various examples of conceivable combinations of
hue, chroma and luminance. The first palette uses a broader range on the luminance axis
whereas the others use larger ranges on the chroma axis.

\section{Illustrations}

In this section, we show a few examples for the various types of palettes applied
to statistical graphics. The first illustration is visualization of a bivariate
density estimation for the Old Faithful geyser eruptions data. Figure\~ref{fig:bkde}
shows heatmaps of a bivariate kernel density estimate of waiting times between
and duration of geyser eruptions in Yellowstone National Park. Both use a sequential
palette as derived in Section~\ref{sec:sequential} balanced towards the same gray levels
with $L \in [30, 90]$ and $p_2 = 2$. The sequential palette in the left panel uses only gray
colors (i.e., $C_{\max} = 0$) and the palette in the right panel additionally employs
colors with $H \in [0, 90]$, $C \in [30, 80]$ and $p_1 = 1/5$.

\setkeys{Gin}{width=\textwidth}
\begin{figure}[b!]
\begin{center}
<<bkde, height=6, width=12>>=
if(generate_png) {
  png(file = "bkde.png", height = 420, width = 840)
  par(mfrow = c(1, 2))
  data("geyser", package = "MASS")
  dens <- bkde2D(cbind(geyser$duration, geyser$waiting), bandwidth = c(0.2, 3), gridsize = c(201, 201))
  image(dens$x1, dens$x2, sqrt(dens$fhat), xlab = "duration", ylab = "waiting", 
    col = rev(sequential_hcl(33, c = 0, power = 2)))
  box()
  image(dens$x1, dens$x2, sqrt(dens$fhat), xlab = "duration", ylab = "waiting", 
    col = rev(heat_hcl(33, c = c(80, 30), l = c(30, 90), power = c(1/5, 2))))
  box()
  dev.off()
}
@
\includegraphics{bkde}
\caption{\label{fig:bkde} Bivariate density estimation for Old Faithful geyser eruptions}
\end{center}
\end{figure}

\setkeys{Gin}{width=.6\textwidth}
\begin{figure}[p]
\begin{center}
<<seats, fig=TRUE, height=5, width=6>>=
par(mar = rep(0.8, 4))
pie(seats, clockwise = TRUE, col = hcl(c(240, 60, 300, 120, 0), 60, 75), radius = 1)
@
\caption{\label{fig:seats} Seats in the German parliament}
\end{center}
\end{figure}

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[p]
\begin{center}
<<votes, fig=TRUE, height=9, width=11>>=
votes <- votes[c(1, 3, 4, 5, 9, 11, 13, 14, 15, 16, 2, 6, 7, 8, 10, 12),]
votes <- votes[,c(2, 4, 1, 3, 5)]
votes <- as.table(as.matrix(votes))
names(dimnames(votes)) <- c("", "")

mosaic(votes, labeling_args = list(rot_labels = c(0, 90, 0, 0), pos_labels = "center",
  just_labels = c("center", "center", "center", "right")),
  gp = gpar(fill = hcl(c(240, 60, 0, 120, 300), 60, 75)),
  spacing = spacing_highlighting, labeling = labeling_left,
  margins = unit(c(2.5, 1, 1, 12), "lines"), keep = FALSE)
@
\caption{\label{fig:votes} Votes in the German election 2005}
\end{center}
\end{figure}

To illustrate qualitative palettes, data from the 2005 election for the German parliament `Bundestag'
are employed. In that election, five parties were able to obtain enough votes to enter the
Bundestag---the distribution of seats is depicted in a pie chart in Figure~\ref{fig:seats}.
The colors used are rough metaphors for the political parties, using a red hue $H = 0$ for the
social democrats SPD, a blue hue $H = 240$ for the conservative CDU, a yellow hue $H = 60$
for the liberal FDP, a green hue $H = 120$ for the green party `Die Gruenen' and a purple 
hue $H = 300$ for the leftist party `Die Linke'. All colors use the same chroma $C = 60$ and
luminance $L = 75$. The pie chart clearly shows that neither the governing coalition of SPD
and Gruene nor the opposition of CDU and FDP could assemble a majority. Given that no party
would enter a coalition with the leftists, this lead to a big coalition of CDU and SPD.
Figure~\ref{fig:votes} shows the distribution of votes in this election stratified by 
province (Bundesland) in a highlighted mosaic display. The order of provinces is from 
north to south, first for the 10 Western provinces, then for the 6 Eastern provinces. Clearly,
the SPD performed better in the North and the CDU better in the South; furthermore, the
Die Linke performed particularly well in the Eastern provinces and the Saarland.

As pointed out in Section~\ref{sec:diverging}, diverging palettes are particularly useful
when visualizing residuals or correlation (with natural neutral value 0) or probabilities 
in 2-class supervised learning (with neutral value 0.5). Examples for both situations are
provided here. Figure~\ref{fig:arthritis} visualizes the outcome of a double-blind clinical
trial investigating a new treatment for rheumatoid arthritis. The mosaic rectangles alone
signal that the treatment lead to higher improvement compared to the placebo group which are
shown to be significant by the shading based on the Pearson residuals. Positive residuals, corresponding
to more observations in the corresponding cell than expected under independence, are depicted
in blue, negative residuals in red. Light colors signal significance at 10\% level, full colors
significance at 1\% level. Hence, it can be concluded that there are significantly more
marked improvements in the treated group and significantly fewer in the placebo group than
would be expected under independence between treatment and improvement. More details can
be found in \citep{vcd:Zeileis+Meyer+Hornik:2005}.

Figure~\ref{fig:class} shows the fit of a support vector machine (SVM) to an artificial 
2-class supervised learning example: a mixture of two bivariate normal distributions.
The circles and triangles show the original observations, solid symbols correspond to
the support vectors found. The shading underlying the plot visualizes the fitted decision
values: values around 0 are on the decision boundary and are shaded in light gray, while
regions that are firmly classified to one or the other class are shaded in full blue and
red respectively.

\setkeys{Gin}{width=.7\textwidth}
\begin{figure}[p]
\begin{center}
<<arthritis, fig=TRUE, height=6, width=7>>=
art <- xtabs(~ Treatment + Improved, data = Arthritis, subset = Sex == "Female")
names(dimnames(art))[2] <- "Improvement"
set.seed(1234)
mosaic(art, gp = shading_max)
@
\caption{\label{fig:arthritis} Extended mosaic display for arthritis data}
\end{center}
\end{figure}

\setkeys{Gin}{width=.7\textwidth}
\begin{figure}[p]
\begin{center}
<<class, height=6, width=8>>=
if(generate_png) {
  png(file = "class.png", height = 480, width = 640)
  set.seed(123)
  x1 <- rmvnorm(75, mean = c(1.5, 1.5), sigma = matrix(c(1, 0.8, 0.8, 1), ncol = 2))
  x2 <- rmvnorm(75, mean = c(-1, -1), sigma = matrix(c(1, -0.3, -0.3, 1), ncol = 2))
  X <- rbind(x1, x2)
  ex1 <- data.frame(class = factor(c(rep("a", 75), rep("b", 75))), x1 = X[,1], x2 = X[,2])
  rm(x1, x2, X)
  fm <- ksvm(class ~ ., data = ex1, C = 0.5)
  plot(fm, data = ex1)
  dev.off()
}
@
\includegraphics{class}
\caption{\label{fig:class} SVM classification plot}
\end{center}
\end{figure}

\section{Open questions}

From our experience, the paths through HCL space described in the previous sections
in principle work very well and can be used to effectively code qualitative and 
quantitative information. However, there are some open questions, formulated below.
In particular, we are not sure how we can properly define trade-offs
between increasing the range with respect to one dimension and decreasing the
range with respect to another.

\begin{itemize}
  \item In sequential/diverging palettes, there is a trade-off between high chroma and high luminance.
        How should this be chosen in practice? Our impression is that when a small set of colors
        (such as 3 or 4) are used, large differnces in chroma work well and large differences
        in luminance are not necessary. However, when a larger set of colors is used (e.g., for 
        heatmaps where extreme values should be identifiable) it is much more important to have a big
        difference in luminance.

  \item How should the intensity $|i|$ be increased from 0 to 1?
        Our experience is that for a small set of colors linear increase is sufficient ($p = 1$)
        whereas in heatmaps where only very extreme regions are interesting a $p > 1$
        should be used.

  \item For diverging palettes, how should pairs of colors be chosen? The hues $H = 0$
        and $H = 260$ where chosen because they are on opposite sides of the color wheel
	but have a very similar chroma/luminance plane.

  \item To allow more convenient navigation in HCL space, some more \proglang{R} infrastructure
        could be helpful. We are not sure what would the best or most natural interface for this.
	A good start might be a function that computes the admissible value on one axis given 
	a particular value on the other two axes. Maybe you have already experimented with 
	something like this?
\end{itemize}

\bibliography{vcd}

\end{document}