### 2009-02-23 pkg/vcd/paper/Color/letter.tex
rearrange repository
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@690 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{wuletter}
\signature{Achim Zeileis, Kurt Hornik}
\usepackage[american]{babel}
\begin{document}
\begin{letter}{
Peter Westfall, Editor\\
MS 2101\\
Texas Tech University\\
Lubbock, TX  79409\\
806-742-2174}

\concerning{Manuscript submission}

\opening{Dear Prof.~Westfall,}

please find attached our manuscript
\begin{quote}
\textit{Choosing Color Palettes for Statistical Graphics}
\end{quote}
as a submission for the ``Statistical Computing and Graphics''
section of \emph{The American Statistician} (TAS). We feel that the
choice of suitable colors coding statistical information in
graphical displays is a topic of great interest to TAS
readers because it is a problem all of us face in every-day
practice. The paper discusses some simple algorithms how 
suitable perceptually-based colors can be chosen. \textsf{R}
software implementing these color selection strategies is
also freely available.

A potential problem might be the inevitable use of color 
within the paper because it is, to the best of our knowledge,
not possible to have color pages in the printed version of TAS.
If it is possible, we would be happy to contribute to the extra
costs---if not, we could easily make an electronic version of
the graphics available to the readers and, of course, all \textsf{R}
users can easily reproduce the graphics from the paper using our
software.

We consider TAS to be the ideal outlet for discussing a topic of
such broad interest and would be very pleased if it were considered
for review as a TAS submission.

\closing{Best regards,}
\end{letter}
\end{document}